---
title: "Quality control"
output:
  workflowr::wflow_html:
    code_folding: hide
    fig_caption: yes
    number_sections: yes
    pandoc_args: []
    self_contained: yes
    smart: yes
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: no
      smooth_scroll: yes
editor_options: 
  chunk_output_type: console
---

```{r libs-and-functions, echo = F, message = FALSE, warning=FALSE}
library("Cairo")
library("tidyverse")
library("Seurat")
library("stringr")
library("yaml")
library("ggplot2")
library("cowplot")
library("reshape2")
library("patchwork")
library("kableExtra")
library("scico")
library("semla")

# Michael color themes
source("code/helper/Rmarkdown-style.R")
theme_set(mytheme(base_size = 7))
```

```{r setup, message=FALSE, warning=FALSE, echo=FALSE, cache=F}
knitr::opts_chunk$set(
  echo = FALSE,
  error = FALSE,
  message = FALSE,
  cache = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.align = 'center',
  fig.show = 'hold',
  autodep = TRUE,
  dpi = 96,
  dev = c('CairoPNG') # alternative: dev = c('png') type = 'cairo'
)
```

```{r manifest-and-objects, echo=F}
manifest = yaml.load_file("./assets/manifest.yaml")
current_sample = scan(file.path(manifest$workdata_parent, ".tmp/active_sample"), what = "character", quiet = T)
sample_manifest = manifest[[current_sample]]

default_plot_focus = sample_manifest$crop_area
source("code/helper/FG_plotting_functions.R")
base_pt_size = estim_point_size(sample_manifest)

work.dir = file.path(manifest$workdata_parent, sample_manifest$workdata)
se = readRDS(file.path(work.dir, "01_seurat.rds"))
se.full = readRDS(file.path(work.dir, "01_seurat_full.rds"))

DefaultAssay(se) = "SCT"
spatgenes = CorSpatialFeatures(se)
DefaultAssay(se) = "Spatial"

source("assets/houseKeepingMarkers.R")

mt.not.present = length(rownames(se)[grepl("MT-", rownames(se))]) == FALSE
```


```{r sample-title, results='asis', echo = FALSE}
sample_full_name = sample_manifest$full_name
cat("Sample: ", sample_full_name, "\n")
```

Author: Michael Rade, Florian Große (Last updated: `r Sys.Date()`)


# HE image

```{r HE-image, fig.width=8.5, fig.asp=.7, fig.show = 'hold', echo = F, dpi = 120}
p = ImagePlot(object = se.full, image_use = "transformed") &
  mytheme_histoplot()
invisible(p)
```

# Overlay features on HE image

## Spots outside and inside the tissue region

During library preparation, transcripts can diffuse out into the solution and end
up anywhere outside the tissue. If you expect that you have over-permeabilized
your tissue it could be useful to look at the expression patterns outside the
tissue region as well. 

```{r featureOverlay-all, fig.width=8.5, fig.asp=.7, fig.show = 'hold', fig.cap=figRef("featureOverlay-all-ref", ""), echo=T, eval=T}
feature_overlay(se.full, c("nFeature_Spatial", "nCount_Spatial"), 
                titles = c("No. of genes per spot", "No. of counts per spot"),
                crop_area = NULL, pt.size = 1.1) &
  theme(
      legend.position = "top",
      legend.direction = "horizontal",
      legend.key.width = unit(2., "lines"),
      legend.key.height = unit(0.6, "lines"),
      legend.title = element_blank()
  )
```

## Spots inside the tissue region

**Mitochondrial enrichment**. The following is assumed for individual single cells: cells  with a low count depth,
few detected genes, and a high fraction of mitochondrial counts are indicative of
cells whose cytoplasmic mRNA has leaked out through a broken membrane, and thus,
only mRNA located in the mitochondria is still conserved. `r if(mt.not.present){colorize("No mitochondrial genes were found in the data set.", "red")}`

**Housekeeping genes** from Tirosh et al., Science 2016.: `r I(housekeeping)`

```{r featureOverlay-inside, fig.width=8.5, fig.asp=1.65, fig.show = 'hold', fig.cap=figRef("featureOverlay-inside-ref", "Black spots: Number of genes or counts < 1000"), echo=T, eval=T}
# drop anything below a set number of counts but don't change the scale when the minimum is already higher
generate_minlimit <- function(min_counts){
  fun = function(lims){
    if (is.null(lims)) {
      return(NA)
    } else {
      return(c(max(min_counts, min(lims)), max(lims)))
    }
  }
  return(fun)
}

# legend on top
topbar = theme(
      legend.position = "top",
      legend.direction = "horizontal",
      legend.key.width = unit(2., "lines"),
      legend.key.height = unit(0.6, "lines"),
      legend.title = element_blank()
    )

plot_grid(
  feature_overlay(se, "nFeature_Spatial", "Detected genes per spot", limits = generate_minlimit(1000), 
                  pt.size = base_pt_size) & topbar,
  feature_overlay(se, "nCount_Spatial", "Detected UMIs per spot", limits = generate_minlimit(1000),
                  pt.size = base_pt_size) & topbar,
  feature_overlay(se, "Perc_of_mito_spot", "% of mitochondrial genes", limits = c(0, NA),
                  pt.size = base_pt_size) & topbar,
  feature_overlay(se, "Perc_of_ribosomal_spot", "% of ribosomal genes", limits = c(0, NA),
                  pt.size = base_pt_size) & topbar,
  feature_overlay(se, "hk_exprs_ave", "Average expression of housekeeping genes",
                  pt.size = base_pt_size) & topbar,
  feature_overlay_discrete(se, "Phase", "Cell cycle phase",
                           pt.size = base_pt_size) & topbar,
  ncol = 2
)
```


# Spot stats

```{r spot-stats, fig.width=8.5, fig.asp=.4, fig.cap=figRef("spot-stats-ref", ""), echo=T, eval=T}
p1 = annotatedHist(se$nFeature_Spatial, mads = 1:3) +
  xlab("No. of genes") +
  ylab("No. of spots") +
  ggtitle("No. of genes per spot") + 
  mytheme_grid(aspect.ratio = 0.6180)
p2 = annotatedHist(se$nCount_Spatial, mads = 1:3) +
  xlab("No. of counts") +
  ylab("No. of spots") +
  ggtitle("Total counts per spot") +
  mytheme_grid(aspect.ratio = 0.6180)

gene_attr = data.frame(nUMI = Matrix::rowSums(se@assays$Spatial$counts), nSpots = Matrix::rowSums(se@assays$Spatial$counts > 0))
p3 = annotatedHist(gene_attr$nUMI, mads = 1) +
  scale_x_log10() +
  xlab("No. of counts") +
  ylab("No. of genes") +
  ggtitle("\n\n\nTotal counts per gene (log10 scale)") +
  mytheme_grid(aspect.ratio = 0.6180)

p4 =  se@meta.data %>%
  ggplot(aes(nCount_Spatial, nFeature_Spatial, colour = Perc_of_mito_spot)) +
  geom_point(stroke = .3) +
  scale_colour_scico(palette = "roma", end = .9, direction = -1) +
  guides(colour = guide_colorbar(title.position = "left", title = "% of mitochondrial genes", title.vjust = 1, barwidth = unit(8, "lines"), barheight = unit(.5, "lines"))) +
  xlab("No. of counts per spot") +
  ylab("No. of genes per spot") +
  mytheme_grid(aspect.ratio = 0.6180, legend.position = "top")

plot_grid(p1, p2, scale = 0.95)
plot_grid(p3, p4, scale = 0.95)

# Get raw count data 
umi_data <- GetAssayData(object = se, layer = "counts", assay = "Spatial")
gene_attr <- data.frame(mean = rowMeans(umi_data),
                        detection_rate = rowMeans(umi_data > 0),
                        var = apply(umi_data, 1, var), 
                        row.names = rownames(umi_data)) %>%
  mutate(log_mean = log10(mean), log_var = log10(var))

p5 <- ggplot(gene_attr, aes(log_mean, log_var)) + 
  geom_point(alpha = 0.3, size = .1) + 
  geom_abline(intercept = 0, slope = 1, color = 'red', linewidth = .1) +
  ggtitle("Mean-variance relationship") +
  mytheme_grid(aspect.ratio = 0.6180)

p6 <- ggplot(gene_attr, aes(log_mean, detection_rate)) + 
  geom_point(alpha = 0.3,  size = .1) + 
  ggtitle("Mean-detection-rate relationship") +
  mytheme_grid(aspect.ratio = 0.6180)

plot_grid(p5, p6, scale = 0.95)

p7 = annotatedHist(se$Perc_of_mito_spot, mads = 1:3) +
  xlab("% counts on mitochondrial genes") +
  ylab("No. of spots") +
  ggtitle("Percent of counts belonging to mitochondrial genes per spot") +
  mytheme_grid(aspect.ratio = 0.6180)

p8 = annotatedHist(se$Perc_of_ribosomal_spot, mads = 1:3) +
  xlab("% counts on ribosomal genes") +
  ylab("No. of spots") +
  ggtitle("Percent of counts belonging to ribosomal genes per spot") +
  mytheme_grid(aspect.ratio = 0.6180)

plot_grid(p7, p8, scale = 0.95)
```

# Top genes with highest average expression

```{r tops-ave, fig.width=8.5, fig.asp=1, fig.cap=figRef("tops-ave-ref", "Top 50 Genes with the highest average expression across all spots, along with their expression in each individual spots. The x-axis depicts the percentage of counts accounted for by the top 50 most highly expressed features across the dataset. Each row on the plot corresponds to a gene and is sorted by average expression (denoted by the point). The distribution of expression across all cells is shown as tick marks for each feature."), echo=T, eval=T}

## Find the most highly expressed features in this dataset
exprs_mat = se@assays$Spatial$counts
# exprs_mat = exprs_mat[!grepl("^RPL|^RPS", rownames(exprs_mat)), ]
ave_exprs = rowSums(exprs_mat)
oo = order(ave_exprs, decreasing=TRUE)

chosen = head(oo, 50)
sub_mat = exprs_mat[chosen,,drop=FALSE]
sub_ave = ave_exprs[chosen]


## Compute expression values and reshape them for ggplot.
total_exprs = sum(ave_exprs)
top_pctage = 100 * sum(sub_ave) / total_exprs
sub_mat = 100 * sweep(sub_mat, 2, colSums(exprs_mat), "/", check.margin=FALSE)

ordered_names = factor(names(sub_ave), rev(names(sub_ave))) # rev() so that most highly expressed is last (i.e., highest y-axis).
df_exprs_by_cell_long <- data.frame(
  Cell=rep(seq_len(ncol(sub_mat)), each=nrow(sub_mat)),
  Tag=rep(ordered_names, ncol(sub_mat)),
  value=as.numeric(sub_mat) # column major collapse.
)

## Adding median expression values for each gene.
df_to_plot <- data.frame(Feature=ordered_names)
df_to_plot$pct_total <- 100 * sub_ave / total_exprs

ggplot(df_exprs_by_cell_long, aes(y = Tag, x = value)) +
  geom_point(alpha = 0.6, shape = 124) +
  geom_point(data = df_to_plot, aes(x = pct_total, y = Feature), colour = "red", shape = 16) +
  xlab("% of total counts") + ylab("Genes") +
  mytheme_grid() +
  theme(panel.grid.minor = element_blank(),
        axis.text.x = element_text(colour = "gray35"),
        axis.text.y = element_text(colour = "gray35"),
        axis.title.x = element_text(colour = "gray35"),
        axis.title.y = element_text(colour = "gray35"),
        title = element_text(colour = "gray35"))
```

# Cell marker {.tabset}

<span style="color:red">**Raw counts**</span> are shown.

## Melanoma

```{r Melanoma, fig.width=8.5, eval=T, dpi = 60}
plot_markers <- function(genes){
  plot_grid(plotlist =  lapply(ftrs, function(x){
    feature_overlay(se, x, pt.size = base_pt_size/2, pt.stroke = 0.25, labels = ~ sprintf("%.1f", .x)) & 
      theme(plot.margin = margin(1,1,1,1),
            legend.justification = c(0,0.5),
            legend.margin = margin(1,4,1,1),
            legend.key.height = unit(1.5, "lines"))}), ncol = 3, nrow = 2, scale = 1.02)
}

ftrs = c("MLANA", "DCT", "PMEL", "TYRP1", "TYR", "SOX10")
ftrs = ftrs[ftrs %in% rownames(se)]
if (length(ftrs) > 0) {plot_markers(ftrs)} else {NULL}
```

## T-cells

```{r T-cells, fig.width=8.5, eval=T, dpi = 60}
ftrs = c("CD4", "CD8A", "CD8B", "CD3D", "CD3E", "CD2", "CD247", "TRAC")
ftrs = ftrs[ftrs %in% rownames(se)]
if (length(ftrs) > 0) {plot_markers(ftrs)} else {NULL}
```

## Naive T-cells

```{r Naive-T-cells, fig.width=8.5, eval=T, dpi = 60}
ftrs = c("CCR7", "LEF1", "SELL", "TCF7") # Naive T-cells
ftrs = ftrs[ftrs %in% rownames(se)]
if (length(ftrs) > 0) {plot_markers(ftrs)} else {NULL}
```

## Exhausted T-cells

```{r Exhausted-T-cells, fig.width=8.5, eval=T, dpi = 60}
ftrs = c("CTLA4", "HAVCR2", "LAG3", "PDCD1", "TIGIT") # Exhausted T-cells
ftrs = ftrs[ftrs %in% rownames(se)]
if (length(ftrs) > 0) {plot_markers(ftrs)} else {NULL}
```

## Activated T-cells

```{r Activated-T-cells, fig.width=8.5, eval=T, dpi = 60}
ftrs = c("CD28", "IL2", "CD69", "TNFRSF9", "CD27", "IFNG") # Activated T-cells
ftrs = ftrs[ftrs %in% rownames(se)]
if (length(ftrs) > 0) {plot_markers(ftrs)} else {NULL}
```

## Cytotoxic T-cells

```{r Cytotoxic-T-cells, fig.width=8.5, eval=T, dpi = 60}
ftrs = c("PRF1", "GZMB", "GZMA", "KLRK1", "NKG7", "GNLY") # Cytotoxic T-cells
ftrs = ftrs[ftrs %in% rownames(se)]
if (length(ftrs) > 0) {plot_markers(ftrs)} else {NULL}
```

## memory, activation and cell survival (T-cells)

```{r Mem-act-surv-T-cells, fig.width=8.5, eval=T, dpi = 60}
ftrs = c("IL7R", "TCF7", "REL", "FOXP1", "FOSL2", "STAT4") # memory, activation and cell survival (T-cells)
ftrs = ftrs[ftrs %in% rownames(se)]
if (length(ftrs) > 0) {plot_markers(ftrs)} else {NULL}
```

## NK cells

```{r NK-cells, fig.width=8.5, eval=T, dpi = 60}
ftrs = c("FCGR3A", "NCR1", "NCAM1") # NK cells
ftrs = ftrs[ftrs %in% rownames(se)]
if (length(ftrs) > 0) {plot_markers(ftrs)} else {NULL}
```

## B-cells

```{r B-cells, fig.width=8.5, eval=T, dpi = 60}
ftrs = c("CD19", "MS4A1") # B-cells
ftrs = ftrs[ftrs %in% rownames(se)]
if (length(ftrs) > 0) {plot_markers(ftrs)} else {NULL}
```

## Myeloid, Monocyte, Macrophage

```{r Myeloid-Monocyte-Macrophage, fig.width=8.5, eval=T, dpi = 60}
ftrs = c("CD33", "CD163", "ITGAM", "HLA-DRA", "CD14", "FCGR1A") # Myeloid, Monocyte, Macrophage
ftrs = ftrs[ftrs %in% rownames(se)]
if (length(ftrs) > 0) {plot_markers(ftrs)} else {NULL}
```

## DC

```{r DC, fig.width=8.5, eval=T, dpi = 60}
ftrs = c("IL3RA", "NRP1", "LILRA4","HLA-DRA", "ITGAX", "THBD") # DC
ftrs = ftrs[ftrs %in% rownames(se)]
if (length(ftrs) > 0) {plot_markers(ftrs)} else {NULL}
```

# Spatial autocorrelation

Spatial autocorrelation test was used to find genes with clear spatial patterns. Top 25 genes are presented and top 9 are shown (SCT-transformed expression).

```{r autcor-table, fig.cap=figRef("autcor-table-ref", ""), echo=F, results=T, eval=T}
head(spatgenes[[1]], 25) %>%
  kbl() %>%
  kable_styling(full_width = F)
```

```{r autcor-overlay, fig.width=8.5, fig.asp=1.1, echo=T, eval=T, dpi = 80}
# feature_overlay = function(.se = NULL, .t = NULL, .title = NULL) {
#   FeatureOverlay(
#     .se,
#     features = .t,
#     pt.size = 1,
#     pt.alpha = 1,
#     add.alpha = F,
#     show.sb = F,
#     min.cutoff = "q1",
#     max.cutoff = "q99",
#     pt.border = T) +
#   coord_equal() +
#   scale_fill_scico(palette = "romaO", end = 1, direction = -1) +
#   guides(fill = guide_colorbar(title.position = "top", title.hjust = 0, barwidth = unit(.4, "lines"), 
#                                title = "Exprs", barheight = unit(6, "lines"))) +
#   theme(
#     axis.title = element_blank(),
#     axis.text = element_blank(),
#     axis.ticks = element_blank()
#   )
# }

plot_grid(plotlist = lapply(spatgenes[[1]]$gene[1:9], function(x){
    feature_overlay(se, x, pt.size = base_pt_size/2, pt.stroke = 0.25, 
                    labels = ~ sprintf("%.1f", .x), max_cutoff = 0.99) & 
      theme(plot.margin = margin(1,1,1,1),
            legend.justification = c(0,0.5),
            legend.margin = margin(1,4,1,1),
            legend.key.height = unit(1.5, "lines"))}), ncol = 3)
```

