---
title: "Reference-based Deconvolution Results"
output:
  workflowr::wflow_html:
    code_folding: hide
    fig_caption: yes
    number_sections: yes
    pandoc_args: []
    self_contained: yes
    smart: yes
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: no
      smooth_scroll: yes
editor_options: 
  chunk_output_type: console
---

```{r libs-and-functions, echo = F, message = FALSE, warning=FALSE}
library("Cairo")
library("tidyverse")
library("Seurat")
library("yaml")
library("ggplot2")
library("cowplot")
library("stats")
library("naturalsort")
library("reshape2")
library("patchwork")
library("scales")
library("kableExtra")
library("scico")
library("semla")


# >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
# Functions
# >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
source("code/helper/Rmarkdown-style.R")
theme_set(mytheme(base_size = 7))

```

```{r setup, message=FALSE, warning=FALSE, echo=FALSE, cache=F}
knitr::opts_chunk$set(
  echo = FALSE,
  error = FALSE,
  message = FALSE,
  cache = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.align = 'center',
  fig.show = 'hold',
  autodep = TRUE,
  dpi = 80,
  dev = c('CairoPNG') # alternative: dev = c('png') type = 'cairo'
)
```

```{r manifest-and-objects, echo=F}
manifest = yaml.load_file("./assets/manifest.yaml")
current_sample = scan(file.path(manifest$workdata_parent, ".tmp/active_sample"), what = "character", quiet = T)
sample_manifest = manifest[[current_sample]]

default_plot_focus = sample_manifest$crop_area
source("code/helper/FG_plotting_functions.R")
base_pt_size = estim_point_size(sample_manifest)

work.dir = file.path(manifest$workdata_parent, sample_manifest$workdata)

obj = readRDS(file.path(work.dir, "05_seurat_deconv_results.rds"))
se = obj$se
CARD_obj_broad = obj$CARD_obj_broad
CARD_obj_medium = obj$CARD_obj_medium
CARD_obj_fine = obj$CARD_obj_fine
RCTD_obj_medium = obj$RCTD_obj_medium

ref_collapsing_file = "assets/single_cell_reference/cell_type_state_matching.tsv"
celltype_collapsing = read_tsv(ref_collapsing_file)
```

```{r sample-title, results='asis', echo = FALSE}
sample_full_name = sample_manifest$full_name
cat("Sample: ", sample_full_name, "\n")
```

Author: Florian Große, Michael Rade (Last updated: `r Sys.Date()`)

# CARD Deconvolution results {.tabset .tabset-pills}

These are the deconvolution results generated with the CARD package and a single cell reference consisting of either melanoma or nevi cells from the Stubenvoll et al. 2025
publication. (https://doi.org/10.1002/cac2.12651)

## Broad {.tabset}

Black spots have an estimated proportion <= 1%.
Missing spots were not deconvolved (mostly due to too few counts present). 

```{r CARD_broad, echo=FALSE, results='asis'}
# remove cells that were not deconvolved due to too little genes present, 
# because visualization functions complain otherwise
tmp_broad = se %>% 
  SubsetSTData(spots = rownames(CARD_obj_broad@Proportion_CARD)) %>%
  AddMetaData(CARD_obj_broad@Proportion_CARD)
  
res <- lapply(sort(colnames(CARD_obj_broad@Proportion_CARD)), function(celltype) {
  scalemax = max(0.3, max(CARD_obj_broad@Proportion_CARD[,celltype])+0.01)  
  knitr::knit_child(text = c(
    '',
    '### `r paste(celltype)`',
    '',
    '```{r, echo=F, results=T, eval=T}',
    '',
    'feature_overlay(tmp_broad, celltype, limits = c(0.01, scalemax), scalename = "Cell Type Proportion", pt.size = base_pt_size*1.5) + theme(legend.key.width = unit(0.8, "lines"), legend.key.height = unit(3., "lines"))',
    '',
    '```',
    ''
  ), envir = environment(), quiet = T)
})
cat(unlist(res), sep = '\n\n')
```

## Medium {.tabset}

Black spots have an estimated proportion <= 1%.
Missing spots were not deconvolved (mostly due to too few counts present). 

```{r CARD_medium, echo=FALSE, results='asis'}
# remove cells that were not deconvolved due to too little genes present, 
# because visualization functions complain otherwise
tmp_medium = se %>% 
  SubsetSTData(spots = rownames(CARD_obj_medium@Proportion_CARD)) %>%
  AddMetaData(CARD_obj_medium@Proportion_CARD)
  
res <- lapply(sort(colnames(CARD_obj_medium@Proportion_CARD)), function(celltype) {
  scalemax = max(0.3, max(CARD_obj_medium@Proportion_CARD[,celltype])+0.01)  
  knitr::knit_child(text = c(
    '',
    '### `r paste(celltype)`',
    '',
    '```{r, echo=F, results=T, eval=T}',
    '',
    'feature_overlay(tmp_medium, celltype, limits = c(0.01, scalemax), scalename = "Cell Type Proportion", pt.size = base_pt_size*1.5) + theme(legend.key.width = unit(0.8, "lines"), legend.key.height = unit(3., "lines"))',
    '',
    '```',
    ''
  ), envir = environment(), quiet = T)
})
cat(unlist(res), sep = '\n\n')
```

## Fine {.tabset}

Black spots have an estimated proportion <= 1%.
Missing spots were not deconvolved (mostly due to too few counts present). 

```{r CARD_fine, echo=FALSE, results='asis'}
# remove cells that were not deconvolved due to too little genes present, 
# because visualization functions complain otherwise
tmp_fine = se %>% 
  SubsetSTData(spots = rownames(CARD_obj_fine@Proportion_CARD)) %>%
  AddMetaData(CARD_obj_fine@Proportion_CARD)
  
res <- lapply(sort(colnames(CARD_obj_fine@Proportion_CARD)), function(celltype) {
  scalemax = max(0.3, max(CARD_obj_fine@Proportion_CARD[,celltype])+0.01)  
  knitr::knit_child(text = c(
    '',
    '### `r paste(celltype)`',
    '',
    '```{r, echo=F, results=T, eval=T}',
    '',
    'feature_overlay(tmp_fine, celltype, limits = c(0.01, scalemax), scalename = "Cell Type Proportion", pt.size = base_pt_size*1.5) + theme(legend.key.width = unit(0.8, "lines"), legend.key.height = unit(3., "lines"))',
    '',
    '```',
    ''
  ), envir = environment(), quiet = T)
})
cat(unlist(res), sep = '\n\n')
```










# CARD Deconvolution Summed {.tabset .tabset-pills}

Same as above, but cell type proportions were summed to a canonical set of broad classes and thresholded.

## Broad Summed {.tabset}

Black spots have an estimated proportion <= 1%.
Missing spots were not deconvolved (mostly due to too few counts present). 

Proportions of subtypes were summed together here.

```{r CARD_broad_summed, echo=FALSE, results='asis'}
sum_types_list = lapply(setNames(nm=unique(celltype_collapsing$sum_type)), function(ct){
  unique(celltype_collapsing[celltype_collapsing$sum_type==ct,]$celltype_broad)
})
prop_summed = tibble(!!!lapply(sum_types_list, function(cts){
  if (length(cts)>1) 
  {
    rowSums(CARD_obj_broad@Proportion_CARD[,cts])
  } else {
    CARD_obj_broad@Proportion_CARD[,cts]
  }
})) %>% as.data.frame()
rownames(prop_summed) = rownames(CARD_obj_broad@Proportion_CARD)

# remove cells that were not deconvolved due to too little genes present, 
# because visualization functions complain otherwise
tmp_broad = se %>% 
  SubsetSTData(spots = rownames(prop_summed)) %>%
  AddMetaData(prop_summed)
  
res <- lapply(sort(colnames(prop_summed)), function(celltype) {
  scalemax = max(0.3, max(prop_summed[,celltype])+0.01)  
  knitr::knit_child(text = c(
    '',
    '### `r paste(celltype)`',
    '',
    '```{r, echo=F, results=T, eval=T}',
    '',
    'feature_overlay(tmp_broad, celltype, limits = c(0.01, scalemax), scalename = "Cell Type Proportion", pt.size = base_pt_size*1.5) + theme(legend.key.width = unit(0.8, "lines"), legend.key.height = unit(3., "lines"))',
    '',
    '```',
    ''
  ), envir = environment(), quiet = T)
})
cat(unlist(res), sep = '\n\n')
```

### Melanoma/Immune  (>5%)

```{r, echo=F, results=T, eval=T}
# percentage deconvolution treshold for assigning spots as positive
pct_thresh = 0.05

tmp_broad[["tumor.immune.spots"]] =
  ifelse(tmp_broad$Melanocytic_cells>=pct_thresh,
         ifelse(tmp_broad$`T/NK_cells`>=pct_thresh,
                "Both",
                "Tumor"),
         ifelse(tmp_broad$`T/NK_cells`>=pct_thresh,
                "Immune",
                "None")
         )

print(table(tmp_broad$tumor.immune.spots))

feature_overlay_discrete(tmp_broad, "tumor.immune.spots", pt.size = base_pt_size*1.5) &
  theme(legend.key.width = unit(0.8, "lines"), 
        legend.key.height = unit(3., "lines")) &
  scale_fill_manual(values = c(Both = "yellow", None = "black", Immune = "green", Tumor = "red"), name = paste0("Types exceeding ", sprintf("%2.0d", pct_thresh*100), "%")) & ggtitle(NULL, subtitle = NULL)
```

### Melanoma/Immune  (>10%)

```{r, echo=F, results=T, eval=T}
# percentage deconvolution treshold for assigning spots as positive
pct_thresh = 0.1

tmp_broad[["tumor.immune.spots"]] =
  ifelse(tmp_broad$Melanocytic_cells>=pct_thresh,
         ifelse(tmp_broad$`T/NK_cells`>=pct_thresh,
                "Both",
                "Tumor"),
         ifelse(tmp_broad$`T/NK_cells`>=pct_thresh,
                "Immune",
                "None")
         )

print(table(tmp_broad$tumor.immune.spots))

feature_overlay_discrete(tmp_broad, "tumor.immune.spots", pt.size = base_pt_size*1.5) &
  theme(legend.key.width = unit(0.8, "lines"), 
        legend.key.height = unit(3., "lines")) &
  scale_fill_manual(values = c(Both = "yellow", None = "black", Immune = "green", Tumor = "red"), name = paste0("Types exceeding ", sprintf("%2.0d", pct_thresh*100), "%")) & ggtitle(NULL, subtitle = NULL)
```

### Melanoma/Immune  (>20%)

```{r, echo=F, results=T, eval=T}
# percentage deconvolution treshold for assigning spots as positive
pct_thresh = 0.2

tmp_broad[["tumor.immune.spots"]] =
  ifelse(tmp_broad$Melanocytic_cells>=pct_thresh,
         ifelse(tmp_broad$`T/NK_cells`>=pct_thresh,
                "Both",
                "Tumor"),
         ifelse(tmp_broad$`T/NK_cells`>=pct_thresh,
                "Immune",
                "None")
         )

print(table(tmp_broad$tumor.immune.spots))

feature_overlay_discrete(tmp_broad, "tumor.immune.spots", pt.size = base_pt_size*1.5) &
  theme(legend.key.width = unit(0.8, "lines"), 
        legend.key.height = unit(3., "lines")) &
  scale_fill_manual(values = c(Both = "yellow", None = "black", Immune = "green", Tumor = "red"), name = paste0("Types exceeding ", sprintf("%2.0d", pct_thresh*100), "%")) & ggtitle(NULL, subtitle = NULL)
```

## Medium Summed {.tabset}

Black spots have an estimated proportion <= 1%.
Missing spots were not deconvolved (mostly due to too few counts present). 

Proportions of subtypes were summed together here.

```{r CARD_medium_summed, echo=FALSE, results='asis'}
sum_types_list = lapply(setNames(nm=unique(celltype_collapsing$sum_type)), function(ct){
  unique(celltype_collapsing[celltype_collapsing$sum_type==ct,]$celltype_medium)
})
prop_summed = tibble(!!!lapply(sum_types_list, function(cts){
  if (length(cts)>1) 
  {
    rowSums(CARD_obj_medium@Proportion_CARD[,cts])
  } else {
    CARD_obj_medium@Proportion_CARD[,cts]
  }
})) %>% as.data.frame()
rownames(prop_summed) = rownames(CARD_obj_medium@Proportion_CARD)

# remove cells that were not deconvolved due to too little genes present, 
# because visualization functions complain otherwise
tmp_medium = se %>% 
  SubsetSTData(spots = rownames(prop_summed)) %>%
  AddMetaData(prop_summed)
  
res <- lapply(sort(colnames(prop_summed)), function(celltype) {
  scalemax = max(0.3, max(prop_summed[,celltype])+0.01)  
  knitr::knit_child(text = c(
    '',
    '### `r paste(celltype)`',
    '',
    '```{r, echo=F, results=T, eval=T}',
    '',
    'feature_overlay(tmp_medium, celltype, limits = c(0.01, scalemax), scalename = "Cell Type Proportion", pt.size = base_pt_size*1.5) + theme(legend.key.width = unit(0.8, "lines"), legend.key.height = unit(3., "lines"))',
    '',
    '```',
    ''
  ), envir = environment(), quiet = T)
})
cat(unlist(res), sep = '\n\n')
```

### Melanoma/Immune  (>5%)

```{r, echo=F, results=T, eval=T}
# percentage deconvolution treshold for assigning spots as positive
pct_thresh = 0.05

tmp_medium[["tumor.immune.spots"]] =
  ifelse(tmp_medium$Melanocytic_cells>=pct_thresh,
         ifelse(tmp_medium$`T/NK_cells`>=pct_thresh,
                "Both",
                "Tumor"),
         ifelse(tmp_medium$`T/NK_cells`>=pct_thresh,
                "Immune",
                "None")
         )

print(table(tmp_medium$tumor.immune.spots))

feature_overlay_discrete(tmp_medium, "tumor.immune.spots", pt.size = base_pt_size*1.5) &
  theme(legend.key.width = unit(0.8, "lines"), 
        legend.key.height = unit(3., "lines")) &
  scale_fill_manual(values = c(Both = "yellow", None = "black", Immune = "green", Tumor = "red"), name = paste0("Types exceeding ", sprintf("%2.0d", pct_thresh*100), "%")) & ggtitle(NULL, subtitle = NULL)
```

### Melanoma/Immune  (>10%)

```{r, echo=F, results=T, eval=T}
# percentage deconvolution treshold for assigning spots as positive
pct_thresh = 0.1

tmp_medium[["tumor.immune.spots"]] =
  ifelse(tmp_medium$Melanocytic_cells>=pct_thresh,
         ifelse(tmp_medium$`T/NK_cells`>=pct_thresh,
                "Both",
                "Tumor"),
         ifelse(tmp_medium$`T/NK_cells`>=pct_thresh,
                "Immune",
                "None")
         )

print(table(tmp_medium$tumor.immune.spots))

feature_overlay_discrete(tmp_medium, "tumor.immune.spots", pt.size = base_pt_size*1.5) &
  theme(legend.key.width = unit(0.8, "lines"), 
        legend.key.height = unit(3., "lines")) &
  scale_fill_manual(values = c(Both = "yellow", None = "black", Immune = "green", Tumor = "red"), name = paste0("Types exceeding ", sprintf("%2.0d", pct_thresh*100), "%")) & ggtitle(NULL, subtitle = NULL)
```

### Melanoma/Immune  (>20%)

```{r, echo=F, results=T, eval=T}
# percentage deconvolution treshold for assigning spots as positive
pct_thresh = 0.2

tmp_medium[["tumor.immune.spots"]] =
  ifelse(tmp_medium$Melanocytic_cells>=pct_thresh,
         ifelse(tmp_medium$`T/NK_cells`>=pct_thresh,
                "Both",
                "Tumor"),
         ifelse(tmp_medium$`T/NK_cells`>=pct_thresh,
                "Immune",
                "None")
         )

print(table(tmp_medium$tumor.immune.spots))

feature_overlay_discrete(tmp_medium, "tumor.immune.spots", pt.size = base_pt_size*1.5) &
  theme(legend.key.width = unit(0.8, "lines"), 
        legend.key.height = unit(3., "lines")) &
  scale_fill_manual(values = c(Both = "yellow", None = "black", Immune = "green", Tumor = "red"), name = paste0("Types exceeding ", sprintf("%2.0d", pct_thresh*100), "%")) & ggtitle(NULL, subtitle = NULL)
```

## Fine Summed {.tabset}

Black spots have an estimated proportion <= 1%.
Missing spots were not deconvolved (mostly due to too few counts present). 

Proportions of subtypes were summed together here.

```{r CARD_fine_summed, echo=FALSE, results='asis'}
sum_types_list = lapply(setNames(nm=unique(celltype_collapsing$sum_type)), function(ct){
  unique(celltype_collapsing[celltype_collapsing$sum_type==ct,]$celltype)
})
prop_summed = tibble(!!!lapply(sum_types_list, function(cts){
  if (length(cts)>1) 
  {
    rowSums(CARD_obj_fine@Proportion_CARD[,cts])
  } else {
    CARD_obj_fine@Proportion_CARD[,cts]
  }
})) %>% as.data.frame()
rownames(prop_summed) = rownames(CARD_obj_fine@Proportion_CARD)

# remove cells that were not deconvolved due to too little genes present, 
# because visualization functions complain otherwise
tmp_fine = se %>% 
  SubsetSTData(spots = rownames(prop_summed)) %>%
  AddMetaData(prop_summed)
  
res <- lapply(sort(colnames(prop_summed)), function(celltype) {
  scalemax = max(0.3, max(prop_summed[,celltype])+0.01)  
  knitr::knit_child(text = c(
    '',
    '### `r paste(celltype)`',
    '',
    '```{r, echo=F, results=T, eval=T}',
    '',
    'feature_overlay(tmp_fine, celltype, limits = c(0.01, scalemax), scalename = "Cell Type Proportion", pt.size = base_pt_size*1.5) + theme(legend.key.width = unit(0.8, "lines"), legend.key.height = unit(3., "lines"))',
    '',
    '```',
    ''
  ), envir = environment(), quiet = T)
})
cat(unlist(res), sep = '\n\n')
```

### Melanoma/Immune  (>5%)

```{r, echo=F, results=T, eval=T}
# percentage deconvolution treshold for assigning spots as positive
pct_thresh = 0.05

tmp_fine[["tumor.immune.spots"]] =
  ifelse(tmp_fine$Melanocytic_cells>=pct_thresh,
         ifelse(tmp_fine$`T/NK_cells`>=pct_thresh,
                "Both",
                "Tumor"),
         ifelse(tmp_fine$`T/NK_cells`>=pct_thresh,
                "Immune",
                "None")
         )

print(table(tmp_fine$tumor.immune.spots))

feature_overlay_discrete(tmp_fine, "tumor.immune.spots", pt.size = base_pt_size*1.5) &
  theme(legend.key.width = unit(0.8, "lines"), 
        legend.key.height = unit(3., "lines")) &
  scale_fill_manual(values = c(Both = "yellow", None = "black", Immune = "green", Tumor = "red"), name = paste0("Types exceeding ", sprintf("%2.0d", pct_thresh*100), "%")) & ggtitle(NULL, subtitle = NULL)
```

### Melanoma/Immune  (>10%)

```{r, echo=F, results=T, eval=T}
# percentage deconvolution treshold for assigning spots as positive
pct_thresh = 0.1

tmp_fine[["tumor.immune.spots"]] =
  ifelse(tmp_fine$Melanocytic_cells>=pct_thresh,
         ifelse(tmp_fine$`T/NK_cells`>=pct_thresh,
                "Both",
                "Tumor"),
         ifelse(tmp_fine$`T/NK_cells`>=pct_thresh,
                "Immune",
                "None")
         )

print(table(tmp_fine$tumor.immune.spots))

feature_overlay_discrete(tmp_fine, "tumor.immune.spots", pt.size = base_pt_size*1.5) &
  theme(legend.key.width = unit(0.8, "lines"), 
        legend.key.height = unit(3., "lines")) &
  scale_fill_manual(values = c(Both = "yellow", None = "black", Immune = "green", Tumor = "red"), name = paste0("Types exceeding ", sprintf("%2.0d", pct_thresh*100), "%")) & ggtitle(NULL, subtitle = NULL)
```

### Melanoma/Immune  (>20%)

```{r, echo=F, results=T, eval=T}
# percentage deconvolution treshold for assigning spots as positive
pct_thresh = 0.2

tmp_fine[["tumor.immune.spots"]] =
  ifelse(tmp_fine$Melanocytic_cells>=pct_thresh,
         ifelse(tmp_fine$`T/NK_cells`>=pct_thresh,
                "Both",
                "Tumor"),
         ifelse(tmp_fine$`T/NK_cells`>=pct_thresh,
                "Immune",
                "None")
         )

print(table(tmp_fine$tumor.immune.spots))

feature_overlay_discrete(tmp_fine, "tumor.immune.spots", pt.size = base_pt_size*1.5) &
  theme(legend.key.width = unit(0.8, "lines"), 
        legend.key.height = unit(3., "lines")) &
  scale_fill_manual(values = c(Both = "yellow", None = "black", Immune = "green", Tumor = "red"), name = paste0("Types exceeding ", sprintf("%2.0d", pct_thresh*100), "%")) & ggtitle(NULL, subtitle = NULL)
```









# RCTD Deconvolution {.tabset .tabset-pills}

The following results are generated with the same reference but using the RCTD package in multi-mode (maximum 4 celltypes per spot).

## Medium {.tabset}

Black spots have an estimated proportion <= 1%.
Missing spots were not deconvolved (mostly due to too few counts present). 

```{r RCTD_medium, echo=FALSE, results='asis'}
# sanity check
if (ncol(RCTD_obj_medium@spatialRNA@counts) != length(RCTD_obj_medium@results)) {stop("RCTD object is broken, mismatch between spot names and number of results.")}

RCTD_medium_res_df = do.call(rbind, lapply(RCTD_obj_medium@results, function(spotres){
  ret = setNames(rep(0., length(names(spotres$all_weights))), nm = names(spotres$all_weights))
  ret[spotres$cell_type_list] = spotres$sub_weights
  ret
}))
rownames(RCTD_medium_res_df) = colnames(RCTD_obj_medium@spatialRNA@counts)


# remove cells that were not deconvolved due to too little genes present, 
# because visualization functions complain otherwise

tmp_medium_RCTD = se %>% 
  SubsetSTData(spots = names(RCTD_obj_medium@spatialRNA@nUMI)) %>%
  AddMetaData(RCTD_medium_res_df)
  
res <- lapply(sort(colnames(RCTD_medium_res_df)), function(celltype) {
  scalemax = max(0.3, max(RCTD_medium_res_df[,celltype])+0.01)  
  knitr::knit_child(text = c(
    '',
    '### `r paste(celltype)`',
    '',
    '```{r, echo=F, results=T, eval=T}',
    '',
    'feature_overlay(tmp_medium_RCTD, celltype, limits = c(0.01, scalemax), scalename = "Cell Type Proportion", pt.size = base_pt_size*1.5) + theme(legend.key.width = unit(0.8, "lines"), legend.key.height = unit(3., "lines"))',
    '',
    '```',
    ''
  ), envir = environment(), quiet = T)
})
cat(unlist(res), sep = '\n\n')
```

## Medium Summed {.tabset}

Black spots have an estimated proportion <= 1%.
Missing spots were not deconvolved (mostly due to too few counts present). 

Proportions of subtypes were summed together here.

```{r RCTD_medium_summed, echo=FALSE, results='asis'}
sum_types_list = lapply(setNames(nm=unique(celltype_collapsing$sum_type)), function(ct){
  intersect(unique(celltype_collapsing[celltype_collapsing$sum_type==ct,]$celltype_medium), levels(RCTD_obj_medium@reference@cell_types))
})
prop_summed = tibble(!!!lapply(sum_types_list, function(cts){
  if (length(cts)>1) 
  {
    rowSums(RCTD_medium_res_df[,cts])
  } else {
    RCTD_medium_res_df[,cts]
  }
})) %>% as.data.frame()
rownames(prop_summed) = rownames(RCTD_medium_res_df)

# remove cells that were not deconvolved due to too little genes present, 
# because visualization functions complain otherwise
tmp_medium_summed = se %>% 
  SubsetSTData(spots = rownames(prop_summed)) %>%
  AddMetaData(prop_summed)
  
res <- lapply(sort(colnames(prop_summed)), function(celltype) {
  scalemax = max(0.3, max(prop_summed[,celltype])+0.01)  
  knitr::knit_child(text = c(
    '',
    '### `r paste(celltype)`',
    '',
    '```{r, echo=F, results=T, eval=T}',
    '',
    'feature_overlay(tmp_medium_summed, celltype, limits = c(0.01, scalemax), scalename = "Cell Type Proportion", pt.size = base_pt_size*1.5) + theme(legend.key.width = unit(0.8, "lines"), legend.key.height = unit(3., "lines"))',
    '',
    '```',
    ''
  ), envir = environment(), quiet = T)
})
cat(unlist(res), sep = '\n\n')
```

### Melanoma/Immune  (>5%)

```{r, echo=F, results=T, eval=T}
# percentage deconvolution treshold for assigning spots as positive
pct_thresh = 0.05

tmp_medium_summed[["tumor.immune.spots"]] =
  ifelse(tmp_medium_summed$Melanocytic_cells>=pct_thresh,
         ifelse(tmp_medium_summed$`T/NK_cells`>=pct_thresh,
                "Both",
                "Tumor"),
         ifelse(tmp_medium_summed$`T/NK_cells`>=pct_thresh,
                "Immune",
                "None")
         )

print(table(tmp_medium_summed$tumor.immune.spots))

feature_overlay_discrete(tmp_medium_summed, "tumor.immune.spots", pt.size = base_pt_size*1.5) &
  theme(legend.key.width = unit(0.8, "lines"), 
        legend.key.height = unit(3., "lines")) &
  scale_fill_manual(values = c(Both = "yellow", None = "black", Immune = "green", Tumor = "red"), name = paste0("Types exceeding ", sprintf("%2.0d", pct_thresh*100), "%")) & ggtitle(NULL, subtitle = NULL)
```

### Melanoma/Immune  (>10%)

```{r, echo=F, results=T, eval=T}
# percentage deconvolution treshold for assigning spots as positive
pct_thresh = 0.1

tmp_medium_summed[["tumor.immune.spots"]] =
  ifelse(tmp_medium_summed$Melanocytic_cells>=pct_thresh,
         ifelse(tmp_medium_summed$`T/NK_cells`>=pct_thresh,
                "Both",
                "Tumor"),
         ifelse(tmp_medium_summed$`T/NK_cells`>=pct_thresh,
                "Immune",
                "None")
         )

print(table(tmp_medium_summed$tumor.immune.spots))

feature_overlay_discrete(tmp_medium_summed, "tumor.immune.spots", pt.size = base_pt_size*1.5) &
  theme(legend.key.width = unit(0.8, "lines"), 
        legend.key.height = unit(3., "lines")) &
  scale_fill_manual(values = c(Both = "yellow", None = "black", Immune = "green", Tumor = "red"), name = paste0("Types exceeding ", sprintf("%2.0d", pct_thresh*100), "%")) & ggtitle(NULL, subtitle = NULL)
```

### Melanoma/Immune  (>20%)

```{r, echo=F, results=T, eval=T}
# percentage deconvolution treshold for assigning spots as positive
pct_thresh = 0.2

tmp_medium_summed[["tumor.immune.spots"]] =
  ifelse(tmp_medium_summed$Melanocytic_cells>=pct_thresh,
         ifelse(tmp_medium_summed$`T/NK_cells`>=pct_thresh,
                "Both",
                "Tumor"),
         ifelse(tmp_medium_summed$`T/NK_cells`>=pct_thresh,
                "Immune",
                "None")
         )

print(table(tmp_medium_summed$tumor.immune.spots))

feature_overlay_discrete(tmp_medium_summed, "tumor.immune.spots", pt.size = base_pt_size*1.5) &
  theme(legend.key.width = unit(0.8, "lines"), 
        legend.key.height = unit(3., "lines")) &
  scale_fill_manual(values = c(Both = "yellow", None = "black", Immune = "green", Tumor = "red"), name = paste0("Types exceeding ", sprintf("%2.0d", pct_thresh*100), "%")) & ggtitle(NULL, subtitle = NULL)
```
