---
title: "Clustering"
output:
  workflowr::wflow_html:
    code_folding: hide
    fig_caption: yes
    number_sections: yes
    pandoc_args: []
    self_contained: yes
    smart: yes
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: no
      smooth_scroll: yes
editor_options: 
  chunk_output_type: console
---

```{r libs-and-functions, echo = F, message = FALSE, warning=FALSE}
library("Cairo")
library("tidyverse")
library("Seurat")
library("yaml")
library("ggplot2")
library("cowplot")
library("clustree")
library("naturalsort")
library("reshape2")
library("patchwork")
library("ggthemes")
library("kableExtra")
library("scico")
library("semla")
library("scales")
library("ggrepel")
library("DT")

# bioc
library("org.Hs.eg.db")
library("clusterProfiler")

# >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
# Functions
# >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
source("code/helper/Rmarkdown-style.R")
theme_set(mytheme(base_size = 7))

```

```{r setup, message=FALSE, warning=FALSE, echo=FALSE, cache=F}
knitr::opts_chunk$set(
  echo = FALSE,
  error = FALSE,
  message = FALSE,
  cache = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.align = 'center',
  fig.show = 'hold',
  autodep = TRUE,
  dpi = 96,
  dev = c('CairoPNG') # alternative: dev = c('png') type = 'cairo'
)
```

```{r manifest-and-objects, echo=F}
manifest = yaml.load_file("./assets/manifest.yaml")
current_sample = scan(file.path(manifest$workdata_parent, ".tmp/active_sample"), what = "character", quiet = T)
sample_manifest = manifest[[current_sample]]

default_plot_focus = sample_manifest$crop_area
source("code/helper/FG_plotting_functions.R")
base_pt_size = estim_point_size(sample_manifest)

work.dir = file.path(manifest$workdata_parent, sample_manifest$workdata)
se = readRDS(file.path(work.dir, "06_seurat.rds"))

obj.mrkrs.ora = readRDS(file.path(work.dir, "07_cluster_markers.rds"))

# Cluster marker
markers = obj.mrkrs.ora$cluster.marker.wlcxn
dgea.pvalue = obj.mrkrs.ora$dgea.pvalue
markers.sign = subset(markers, p_val_adj < dgea.pvalue)

# ORA
res.cP.cc.go = obj.mrkrs.ora$res.cP.cc.go
res.cP.cc.go.sim = obj.mrkrs.ora$res.cP.cc.go.sim
ora.oe = obj.mrkrs.ora$tops.l.oe
ora.pvalue = obj.mrkrs.ora$ora.pvalue

DefaultAssay(se) = "SCT"
spatgenes = CorSpatialFeatures(se)
DefaultAssay(se) = "Spatial"

cl.res = obj.mrkrs.ora$cl.res
se$seurat_clusters = se[[cl.res]]
Idents(se) = "seurat_clusters"

# patched because ggthemes::ptol_pal only provides up to 12 colors, if more, add grey nd then some rainbow colors if necessary (although general visibility deteriorates with more colors in this case)
nclasses = length(levels(se$seurat_clusters))
if (nclasses <= 12) {
  cl.col = ggthemes::ptol_pal()(nclasses)
} else {
  cl.col = c(ggthemes::ptol_pal()(12), "grey20", rainbow(nclasses-13))
}

```


```{r sample-title, results='asis', echo = FALSE}
sample_full_name = sample_manifest$full_name
cat("Sample: ", sample_full_name, "\n")
```

Author: Michael Rade, Florian Große (Last updated: `r Sys.Date()`)

# Clustering tree

```{r clustree, fig.width=8.5, fig.asp=.75, fig.show = 'hold', fig.cap=figRef("clustree-ref", "Results from clustering using Seurat with resolution parameters from zero to 1. Nodes are colored according to the value of the resolution and sized according to the number of samples they represent. Edges are colored according to the number of samples (from blue representing few to yellow representing many). The transparency is adjusted according to the in-proportion, with stronger lines showing edges that are more important to the higher-resolution cluster. Cluster labels are randomly assigned by the clustering algorithm. [Paper](https://doi.org/10.1093/gigascience/giy083)"), echo=T, eval=T}
clustree(se, prefix = "SCT_snn_res.", node_size_range = c(4, 11), node_text_colour = "white") +
  mytheme() +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.border = element_rect(fill = NA, colour = NA)
  ) +
  scale_color_manual(values = tol11qualitative) +
  guides(color = guide_legend(ncol = 2))

```

```{r clustree-ftrs, fig.width=8.5, fig.asp=.5, fig.show = 'hold', fig.cap=figRef("clustree-ftrs-ref", "Top 10 genes from spatial autcorrelation"), echo=T, eval=T}
gene_list <- spatgenes[[1]]$gene[1:10]
plot_list <- lapply(gene_list, function(gene) {
  clustree(
    se, prefix = "SCT_snn_res.", node_colour = gene, node_colour_aggr = "mean", assay = "RNA", exprs = "data",
    node_size_range = c(1, 5), edge_width = 0.3, node_text_size = 0
  ) +
    scale_colour_scico(palette = "lajolla", begin = .2, direction = -1) +
    ggtitle(gene) +
    theme(
      legend.position = "none",
      plot.title = element_text(size = rel(1), hjust = 0.5, vjust = 1, margin = margin(5.5)
      )
    )
})

genes_plot <- plot_grid(plotlist = plot_list, nrow = 2)
genes_plot
```

<div class="alert alert-info">
  Following cluster resolution is used for downstream analysis.: `r I(gsub(".*res.", "", cl.res))`
</div>

# UMAP visualization of the clusters & HE Overlays

```{r umap, fig.width=5, fig.show = 'hold', fig.cap=figRef("umap-ref", ""), echo=T, eval=T}
DimPlot(se, group.by = "seurat_clusters", pt.size = .5, reduction = "umap") +
  mytheme() +
  theme(
    aspect.ratio = 1 ,
    legend.position = "right",
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = rel(1), hjust = .5)
  ) +
  ggtitle(NULL) +
  scale_colour_manual(values = cl.col) +
  guides(colour = guide_legend(title = "Cluster", ncol = 1, override.aes = list(size=8)))
```

```{r cluster-overlay, fig.width=8.5, fig.asp=1, fig.show = 'hold', fig.cap=figRef("cluster-overlay-ref", ""), echo=T, eval=T}
# TODO: currently missing the legend but this is the same as in the picture above, so no big deal
MapLabels(se, 
          column_name = "seurat_clusters", 
          section_number = 1,
          pt_size = 1., 
          pt_stroke = 0.2,
          image_use = "transformed", 
          crop_area = default_plot_focus,
          split_labels = T, 
          colors = cl.col, drop_na = T) & 
  theme(legend.position = "none") &
  ggtitle(NULL, subtitle = NULL)
```

# Marker genes for clusters

Genes that are significantly differentially expressed (FDR < `r I(dgea.pvalue)`)
in one cluster compared to all other spots. Here we use the Wilcoxon rank sum test
implemented in the Seurat R package. 

Function used: Seurat::FindMarkers(obj, clusterX, min.pct = 0.5)  This function was called for each of the clusters found

## Top 10 marker genes per cluster

Top 10 marker genes (FDR < `r I(dgea.pvalue)`) ordered by **positive** log2-fold change.

```{r marker-genes-tops, fig.width=8.5, fig.asp=1, fig.show = 'hold', fig.cap=figRef("marker-genes-tops-ref", ""), echo=T, eval=T}
top = subset(markers.sign, avg_log2FC > 0) %>% group_by(cluster) %>% top_n(10, avg_log2FC)
DoHeatmap(se, features  = top$gene, raster = F, angle = 0, size=3, hjust=0, draw.lines = F, assay = "Spatial", group.color = cl.col) +
  scale_fill_scico(palette = "vik", end = 1., direction = 1) +
  mytheme() +
  theme(axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks=element_blank(),
        axis.title=element_blank(),
        line = element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        plot.background=element_blank()) +
  labs(colour = "Cluster", fill = "Standardized\nepxression")
```

## All marker genes per cluster

Showing up- and down-regulated genes across all clusters. An adjusted p value < `r I(dgea.pvalue)` is indicated in red, while an adjusted p value ≥ `r I(dgea.pvalue)` is indicated in black.

```{r marker-genes, fig.width=8.5, fig.asp=.7, fig.show = 'hold', fig.cap=figRef("marker-genes-ref", "."), echo=T, eval=T}
cluster.de.summary = markers.sign %>%
    dplyr::mutate(IsUp = avg_log2FC > 0) %>% 
    dplyr::group_by(cluster) %>%  
    dplyr::summarise(Up = sum(IsUp), Down = sum(!IsUp)) %>% 
    dplyr::mutate(Down = -Down) %>%
    tidyr::gather(key = "Direction", value = "Count", -cluster)

cl.label = paste0(
  "Up: ",
  subset(cluster.de.summary, Direction == "Up")$Count,
  "\nDown: ",
  abs(subset(cluster.de.summary, Direction == "Down")$Count)
)

top5_up = markers.sign %>%
  dplyr::filter(avg_log2FC > 0) %>%
  dplyr::group_by(cluster) %>%
  dplyr::top_n(n = 5, wt = avg_log2FC)
top5_up$GENE_CL = paste0(top5_up$gene, "_", top5_up$cluster)

top5_down = markers.sign %>% 
  dplyr::filter(avg_log2FC < 0) %>%
  dplyr::group_by(cluster) %>%
  dplyr::top_n(n = -5, wt = avg_log2FC)
top5_down$GENE_CL = paste0(top5_down$gene, "_", top5_down$cluster)

markers$GENE_CL = paste0(markers$gene, "_", markers$cluster)
markers = markers %>% mutate(label_up = ifelse(GENE_CL %in% top5_up$GENE_CL, gene, ""))
markers = markers %>% mutate(label_down = ifelse(GENE_CL %in% top5_down$GENE_CL, gene, ""))
markers$SIGNIFICANT = markers$p_val_adj < dgea.pvalue
markers$SIGNIFICANT = factor(markers$SIGNIFICANT, levels = c(TRUE, FALSE))

axis.max = max(abs(markers$avg_log2FC)) + .1
no.cl = length(unique(markers$cluster))
top1.up= markers.sign %>% group_by(cluster) %>% slice_max(order_by = avg_log2FC, n = 1, with_ties = F)
top1.down= markers.sign %>% group_by(cluster) %>% slice_min(order_by = avg_log2FC, n = 1, with_ties = F)
pos = position_jitter(width = 0.35, seed = 1234)

ggplot() +
  # ylim(c(-axis.max, axis.max)) +
  geom_hline(yintercept = 0) +
  geom_rect(
    data = data.frame(xmin = seq(-0.4,no.cl-1), xmax = seq(-0.2,no.cl-1) + .6, ymin = top1.down$avg_log2FC - .1, ymax = top1.up$avg_log2FC + .1),  
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), fill = "#DDDDDD", alpha = .3) +
  geom_rect(
    data = data.frame(xmin = seq(-0.5,no.cl-1), xmax = seq(-0.5,no.cl-1) + 1, ymin = -.3, ymax = .3),
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), fill = cl.col, colour = "black", lwd = .2) +  
  geom_jitter(
    data = subset(markers, SIGNIFICANT == F),
    aes(x= as.numeric(cluster), y= avg_log2FC, color = SIGNIFICANT), position = pos, size = .3
  )  +
  geom_jitter(
    data = subset(markers, SIGNIFICANT == T),
    aes(x= as.numeric(cluster), y= avg_log2FC, color = SIGNIFICANT), position = pos, size = .3
  )  +
  geom_text_repel(
    data = subset(markers, SIGNIFICANT == T),
    mapping = aes(x = as.numeric(cluster), avg_log2FC, label = label_up),
    position = pos, size = 2, max.overlaps = 30
  ) +
  geom_text_repel(
    data = subset(markers, SIGNIFICANT == T),
    mapping = aes(x = as.numeric(cluster), avg_log2FC, label = label_down),
    position = pos, size = 2, max.overlaps = 30
  ) +
  mytheme(base_size = 8) +
  theme(
    aspect.ratio = 0.618,
    legend.position="bottom",
    axis.line.y = element_line(colour = "black"),
    panel.border = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  scale_color_manual(values = c("TRUE" = "#009988", "FALSE" = "black")) +
  labs(x = "Cluster", y = "average log2-fold-change", colour = paste0("FDR < ", 0.05))  +
  annotate("text", x = 0:(no.cl-1), y = 0, label = 0:(no.cl-1), colour = "white", size = 4) +
  annotate("text", x = 0:(no.cl-1), y = max(abs(markers$avg_log2FC)) + .5, label = cl.label, colour = "#009988", size = 2.5) +
  guides(colour = guide_legend(override.aes = list(size=3)))
```

## Enrichment test for GO biolocial process terms (FDR < `r I(ora.pvalue)`) 

Enrichment test based to cluster markers (FDR < `r I(dgea.pvalue)`) with **positive** log2-fold change.

## Top 10 by adjusted p-value

```{r ora_top_5_p, fig.width=8.5, fig.asp=1.2, fig.show = 'hold', fig.cap=figRef("ora_top_5_p-ref", "The dot size indicates the rich factor, which is the number of GEP genes in the GO term divided by the number of background genes in that term. Colours indicate the adjusted p-values of the significantly enriched GO term. The 10 most significant pathways (sorted by adjusted p-value) for each cluster are shown. How to read: For example, Pathway A is significantly enriched in Clusters 1 and 2. Pathway A is in the 5 most significant pathways in cluster 1, but not in cluster 2. To capture this information, Pathway A is included in Cluster 2, making the comparison in the dot plot more meaningful. The dot border indicates whether the pathways are among the 10 most significant."), echo=T, eval=T}

ora_dotplot = function(.tops = 10, .by = c("richFactor", "p.adjust"), cluster_order = NULL) {
  .by <- match.arg(.by)
  
  res = res.cP.cc.go.sim@compareClusterResult
  res = mutate(res, totalGenes = as.numeric(sub("\\d+/", "", GeneRatio)),
               richFactor = Count / as.numeric(sub("/\\d+", "", BgRatio)))
  # at least 5 genes or 10% (but never less than 2) of the positive marker genes of a cluster must 
  # belong to a term for it to be included
  res = res[(res$Count >= pmin(5, pmax(2, res$totalGenes*0.1))), ]
  res$geneID = NULL
  res$ID_CL = paste0(res$ID, "_", res$Cluster)
  
  # New line for long Description strings
  for (i in 1:nrow(res)) {
      res$Description[i] = paste(base::strwrap(res$Description[i], 70), collapse="\n")
  }
  
  res.cp = res
  
  tops.l = lapply(split(res, res$Cluster), function(x){
    if (.by == "richFactor") {
      x = x[order(x$richFactor, decreasing = T), ]  # sort by enrichment factor
    } else {
      x = x[order(x$p.adjust, decreasing = F), ]    # sort by adjusted p value
    }
    x = head(x, .tops)
    x
  })
  
  tops.l = tops.l[sapply(tops.l, function(x) dim(x)[1]) > 0]
  tops = as.data.frame(data.table::rbindlist(tops.l))
  tops$ID_CL = paste0(tops$ID, "_", tops$Cluster)
  res.cp = subset(res.cp, ID %in% tops$ID)
  
  lvls = tops.l[[1]]$Description
  lvls = c(lvls, setdiff(res.cp$Description, lvls))
  res.cp$Description = factor(res.cp$Description, levels = rev(lvls))
  res.cp$Cluster = factor(res.cp$Cluster, levels = unique(naturalsort::naturalsort(res.cp$Cluster)))
  
  res.cp$TOPS = res.cp$ID_CL %in% tops$ID_CL
  
  
  ora = ggplot(res.cp, aes(x = Cluster, y = Description)) +
    geom_point(aes(fill = p.adjust, size = richFactor), pch=21, stroke=.3) +
    geom_point(aes(colour = TOPS, size = richFactor), pch=21, stroke=.5) +
    scale_size(range = c(2,6)) +
    mytheme_grid(base_size = 8) +
    theme(
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.ticks = element_blank(),
      legend.position = "right"
    ) +
    guides(fill = guide_colorbar(
      title.position = 'top', title.hjust = 0, barwidth = unit(.5, 'lines'), barheight = unit(8, 'lines'),
      title = "Adjusted\np-value", order = 3)
    ) + 
    guides(size = guide_legend(
      ncol = 1, override.aes =, title.hjust = .5, title.position = "top",
      title = "Rich factor", order = 1)
    ) +
    guides(colour = guide_legend(
      ncol = 1, override.aes =, title.hjust = .5, title.position = "top",
      title = "Tops", order = 2)
    ) +
    # scale_fill_scico(palette = 'batlow', direction = -1, end = .7)  +
    scale_fill_scico(palette = 'bilbao', begin = .2, end = .8, direction = -1) +
    scale_colour_manual(values = c("#555555", "#44AA99")) +
    xlab("Cluster") + ylab(NULL)  

  if (!is.null(cluster_order)){
    ora = ora + lims(x = cluster_order)
  }
    
  ora
}

cl.labels = paste(sort(unique(se$seurat_clusters)))
plot(ora_dotplot(10, cluster_order = cl.labels, .by = "p"))
```

<br>

* Meaning of BgRatio, M/N and GeneRatio k/n
  * M = size of the gene-set (e.g. number of genes in the cell cycle term). More precisely, the total number of genes (ENTREZ Gene IDs) from the universe found in this gene-set.
  * N = size of all unique genes in the collection of gene-sets (e.g. the GO term collection for BPs). More precisely, the total number of genes (ENTREZ Gene IDs) from the universe found in this collection
  * k = size of overlap of genes with a specific gene-set (e.g. cell cycle term). Only unique ENTREZ Gene IDs were considered.
  * n = size of overlap of genes with all genes in the collection of gene-sets (e.g. the GO term collection for BPs). Only unique ENTREZ Gene IDs were considered.
  * The rich factor is k / M

<br>

```{r table_go, eval=T, results=T}
cp.res = res.cP.cc.go.sim@compareClusterResult
cp.res = cp.res[cp.res$Count >= 5, ]
cp.res = cp.res[naturalorder(cp.res$Cluster), ]

cp.res = cp.res %>%
  dplyr::rename(GO_ID = ID, Term = Description, Gene_Members = geneID) %>%
  dplyr::mutate(go_description_link = paste0('<a href=\'http://amigo.geneontology.org/amigo/term/', GO_ID,'\' target=\'_blank\'>',Term,'</a>')) %>%
  dplyr::mutate(q_value = as.numeric(qvalue)) %>%
  dplyr::mutate(p_value = as.numeric(pvalue)) %>%
  dplyr::mutate(Adjusted_p_value = as.numeric(p.adjust)) %>%
  dplyr::mutate(q_value = dplyr::if_else(!is.na(q_value), as.numeric(formatC(q_value, format = "e", digits = 2)), as.numeric(NA))) %>%
  dplyr::mutate(p_value = dplyr::if_else(!is.na(p_value), as.numeric(formatC(p_value, format = "e", digits = 2)), as.numeric(NA))) %>%
  dplyr::mutate(Adjusted_p_value = dplyr::if_else(!is.na(Adjusted_p_value), as.numeric(formatC(Adjusted_p_value, format = "e", digits = 2)), as.numeric(NA))) %>%
  tidyr::separate(BgRatio, c('num_background_hits','num_background_all'), sep='/',remove = F, convert = T) %>% 
  dplyr::mutate(Rich_factor = round(as.numeric(Count /num_background_hits), digits = 3))

cp.res$Gene_Members = gsub("\\/", ", ", cp.res$Gene_Members)

cp.res = 
  cp.res %>%
  dplyr::select(Cluster, Term, Rich_factor, Adjusted_p_value, BgRatio, GeneRatio, Gene_Members, Count, GO_ID, q_value, p_value) %>%
  dplyr::group_by(Cluster) %>% 
  dplyr::arrange(Cluster, desc(Rich_factor))


cp.res%>%
  DT::datatable(
    rownames = FALSE, escape = F, width = '100%', extensions = c('FixedColumns', 'Scroller', 'Buttons', 'KeyTable', "Responsive"),
    options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    deferRender = F,
    scroller = F
    )
)
```

# Top GO terms overlayed on the HE images

Top 6 GO Terms for all clusters.

Genes from GO Terms based on  marker genes (FDR < `r I(dgea.pvalue)`) with positive log2-fold change.

What is overall expression (OE): For each set of upregulated enrichment genes,
we calculated a signature score (i.e., a composite expression score of a set of genes)
using Seurat’s AddModuleScore function. The signature scores are standardized.

More precisely: The score represents an average expression level of these genes
on a spot level minus the aggregated expression of control gene sets.
All analyzed genes were binned based on averaged expression, and the control genes
were randomly selected from each bin.

```{r ora-overlay, fig.width=8.5, fig.asp=0.9, fig.cap=figRef("ora-overlay-ref", ""), echo=F, results=F, eval=F}
# feature_overlay = function(.se = NULL, .t = NULL, .title = NULL) {
#   FeatureOverlay(
#     .se,
#     features = .t,
#     pt.size = 1,
#     pt.alpha = 1,
#     add.alpha = F,
#     show.sb = F,
#     min.cutoff = "q1",
#     max.cutoff = "q99",
#     pt.border = T) +
#     coord_equal() +
#     scale_fill_scico(palette = "romaO", end = 0.8, direction = -1) +
#     guides(fill = guide_colorbar(title.position = "top", title.hjust = 0, barwidth = unit(.4, "lines"), title = "OE", barheight = unit(6, "lines"))) +
#     theme(
#       aspect.ratio = 1,
#       axis.title = element_blank(),
#       axis.text = element_blank(),
#       axis.ticks = element_blank()
#     ) 
# }

cp.res = res.cP.cc.go.sim@compareClusterResult %>% dplyr::select(Description, ID)
cp.res$ID = gsub(":", ".", cp.res$ID)
cp.res = cp.res[!duplicated(cp.res$ID), ]
rownames(cp.res) = cp.res$ID

maxplots.per.cluster = 6

ora.metric = ora.oe
o.l = list()
for (i in names(ora.metric)) {
#for (i in c("1","3","4","5")) {
  #print(i)
  d = ora.metric[[i]]
  
  for (j in 1:(min(maxplots.per.cluster, length(colnames(d))))) {
    j = colnames(d)[j]
    print(j)
    
    pl.title = cp.res[j, ]
    pl.title$Description = paste(base::strwrap(pl.title$Description, 40), collapse="\n")
    pl.title =  paste0("Cluster ", i, "; ", pl.title$ID,"\n", pl.title$Description)
    
    se.ora = Seurat::AddMetaData(se, d[, j, drop = F])
    o.l[[i]][[j]] = feature_overlay(se.ora, features = colnames(se.ora@meta.data)[grepl("GO.", colnames(se.ora@meta.data))]) +
      ggtitle(pl.title, subtitle = NULL)
  }
}

if (length(o.l)>0) {
  for (i in 1:length(o.l)) { #FG: fails if o.l==0, added quick and dirty catch above
    plot(plot_grid(plotlist = o.l[[i]], ncol = 3, nrow = 4), axis = "tb", align = "h")
  }
}
```

<!-- # Cluster Marker Enrichment Plots -->

<!-- ## Gene-Concept Network -->

<!-- cnetplot from enrichplot.R. See http://yulab-smu.top/biomedical-knowledge-mining-book/enrichplot.html#cnetplot  -->

<!-- Both the barplot() and dotplot() only displayed most significant or selected enriched terms, while users may want to know which genes are involved in these significant terms. In order to consider the potentially biological complexities in which a gene may belong to multiple annotation categories and provide information of numeric changes if available, we developed the cnetplot() function to extract the complex association. The cnetplot() depicts the linkages of genes and biological concepts (e.g. GO terms or KEGG pathways) as a network. GSEA result is also supported with only core enriched genes displayed. -->

```{r cnet, echo=F, fig.width=10., fig.asp=2., eval=F}
# last patch broke plotting functions in these packages so they look super ugly
# currently disabled

p1 = cnetplot(res.cP.cc.go.sim, showCategory = 2, cex_category = 1.2, node_label="category")
p2 = cnetplot(res.cP.cc.go.sim, showCategory = 2, cex_gene = 0.8, node_label="gene")

plot_grid(p1, p2, nrow = 2)
```

<!-- ## Enrichment Map -->

<!-- See http://yulab-smu.top/biomedical-knowledge-mining-book/enrichplot.html#biological-theme-comparison -->

<!-- Enrichment map organizes enriched terms into a network with edges connecting overlapping gene sets. In this way, mutually overlapping gene sets are tend to cluster together, making it easy to identify functional module. -->

```{r emap, echo=F, fig.width=10., fig.asp=1., eval=F}
d = GOSemSim::godata("org.Hs.eg.db", ont="BP")
tmp = enrichplot::pairwise_termsim(res.cP.cc.go, method = "Wang", semData = d)
emapplot(tmp, showCategory = 5, node_label = "none", size_category = 2, min_edge = 0.3, size_edge = 0.25, pie = "count") +
  geom_cnet_label(size = 4)
```
