---
title: "Aggregated Ligand-Receptor Analyses from marker-gene defined regions"
output:
  workflowr::wflow_html:
    code_folding: hide
    fig_caption: yes
    number_sections: yes
    pandoc_args: []
    self_contained: yes
    smart: yes
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: no
      smooth_scroll: yes
editor_options: 
  chunk_output_type: console
---

Author: Florian Große (Last updated: `r Sys.Date()`)

```{r libs-and-functions, echo = F, message = FALSE, warning=FALSE}
library("Cairo")
library("tidyverse")
library("Seurat")
library("yaml")
library("ggplot2")
library("cowplot")
library("rlang")
library("purrr")
library("naturalsort")
library("reshape2")
library("ggnewscale")
library("DT")
library("semla")
library("stringr")
library("scico")
library("circlize")

# bioc
library("SummarizedExperiment")
library("GenomicRanges")
library("recount")
library("biomaRt")

library("NICHES")

library("liana")
# >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
# Functions
# >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
source("code/helper/Rmarkdown-style.R")

theme_set(mytheme(base_size = 7))

render_table  <- function(tabl){
  numericcols = grep(colnames(tabl), pattern = "^zscore|^mean_diff|^log2FC", value = T)
  pvalcols = grep(colnames(tabl), pattern = "^pval", value = T)
  shortcols = setdiff(colnames(tabl), c(pvalcols, "mechanism"))
  
  tabl %>%
  dplyr::select(mechanism, count_signif, summed_zscores, starts_with("significant_in"), 
         starts_with("mean_diff"), starts_with("log2FC"), starts_with("zscore"), everything()) %>%
  arrange(desc(count_signif), desc(summed_zscores)) %>%
  DT::datatable(
    rownames = F, escape = F, width = '100%', extensions = c('FixedColumns', 'Scroller', 'Buttons', 'KeyTable'),
    fillContainer = T,
    options = list(
      buttons = c('copy', 'csv', 'excel'),
      deferRender = T,
      scrollY = 300,
      scroller = T,
      scrollX = T,
      autoWidth = T,
      columnDefs = list(list(width = "150px", targets = "mechanism"), list(width = '50px', targets = shortcols)),
      fixedColumns = list(leftColumns = 1)
    )
  ) %>%
  formatStyle('mechanism',  fontWeight = 'bold') %>%
  formatRound(numericcols, digits = 2)
}

save_table <- function(tabl, filename){
  tabl = tabl %>%
    dplyr::select(mechanism, count_signif, summed_zscores, starts_with("significant_in"), 
           starts_with("mean_diff"), starts_with("log2FC"), starts_with("zscore"), everything()) %>%
    arrange(desc(count_signif), desc(summed_zscores))
  readr::write_excel_csv(x = tabl, file = filename)
}

## Annotation and plotting functions in separate file
source("code/helper/aggregated_LRI_helper.R")

```

```{r setup, message=FALSE, warning=FALSE, echo=FALSE, cache=F}
knitr::opts_chunk$set(
  echo = FALSE,
  error = FALSE,
  message = FALSE,
  cache = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.align = 'center',
  fig.show = 'hold',
  autodep = TRUE,
  dpi = 96,
  dev = c('CairoPNG') # alternative: dev = c('png') type = 'cairo'
)
```

# Melanoma samples

```{r manifest-and-objects, echo=F}
manifest = yaml.load_file("assets/manifest.yaml")
all.samples = setdiff(c(manifest$samples, manifest$external_samples), manifest$rejected_samples)
# niches settings
Niches_result_used = manifest$Niches_result_used
which.result = list(CTN = "border.markers.CellToNeighborhood",
                    NTC = "border.markers.NeighborhoodToCell")[[Niches_result_used]]
Niches_normalization_used = manifest$Niches_normalization_used

# grouping of all samples (melanoma subtypes and so on)
sample_annos = sapply(all.samples, function(sample_id){
  sample_manifest = manifest[[sample_id]]
  return(sample_manifest$type)
})
sample_anno_colmap = setNames(c(tol11qualitative[1:length(unique(sample_annos))], "black"),
                              nm = c(unique(sample_annos), NA))


LRDEs = lapply(all.samples, function(sample_id){
  print(paste("Loading data for sample", sample_id))
  sample_manifest = manifest[[sample_id]]
  work.dir = file.path(manifest$workdata_parent, sample_manifest$workdata)
  resfile = file.path(work.dir, "09_NICHES_LRDE_res.rds")
  has.results = file.exists(resfile)
  if (has.results) {
    res = readRDS(resfile)
  } else {
    print("Sample has no results file, skipping.")
    res = NA
  }
  # contains border.markers.NeighborhoodToCell
  #          border.markers.CellToNeighborhood
  return(res)
})
names(LRDEs) = all.samples

results.are.available = sapply(LRDEs, function(x)all(!(is.na(x))))

all_naevus_samples = intersect(all.samples, manifest$naevus_samples)
all_naevus_samples = all_naevus_samples[results.are.available[all_naevus_samples]]
all_naevus_dfs = LRDEs[all_naevus_samples]
joint_naevi_df = make_consensus_df(all_naevus_dfs, results.name = which.result, results.type = Niches_normalization_used)
joint_naevi_df = annotate_df_with_significance_info(joint_naevi_df)

# POZNIAK data
all_external_samples = intersect(all.samples, manifest$external_samples)
all_external_samples = all_external_samples[results.are.available[all_external_samples]]
all_external_dfs = LRDEs[all_external_samples]
joint_externals_df = make_consensus_df(all_external_dfs, results.name = which.result, results.type = Niches_normalization_used)
joint_externals_df = annotate_df_with_significance_info(joint_externals_df)

# all other
all_mela_samples = setdiff(setdiff(all.samples, manifest$naevus_samples), manifest$external_samples)
all_mela_samples = all_mela_samples[results.are.available[all_mela_samples]]
all_mela_dfs = LRDEs[all_mela_samples]
joint_melas_df = make_consensus_df(all_mela_dfs, results.name = which.result, results.type = Niches_normalization_used)
joint_melas_df = annotate_df_with_significance_info(joint_melas_df)

```


```{r df-prep, echo=F}
TCGA_ref = make_TCGA.se(cohort = "TCGA-SKCM", sites = c("Primary Tumor"))
GTEx.expression.annos = readRDS(file.path(manifest$workdata_parent, "GTEx9_LRI_gene_expression_annos.Rds"))
JA.expression.annos = readRDS(file.path(manifest$workdata_parent, "jerby.se_LRI_gene_expression_annos.Rds"))
StubS.expression.annos = readRDS(file.path(manifest$workdata_parent, "stubs.se_LRI_gene_expression_annos.Rds"))

# add potential cell types of interest in the different annotation sources here
add.annos = function(df){
  # cell.type.filter usage:
  # `display name in results` = `name in data object` # repeat for each celltype of interest
  df = add.TCGA.annos(df, TCGA_reference = TCGA_ref)
  df = add.sc.annos(df, expression.annos = GTEx.expression.annos, annos.name = "GTEx",
                    cell.type.filter = c("DCs/Macrophages" = "Immune (DC/macrophage)", 
                                         "Langerhans Cells" = "Immune (Langerhans)",
                                         "T Cells" = "Immune (T cell)",
                                         "Pericytes/SMCs" = "Pericyte/SMC",
                                         "Lymphatic Endothelial Cells" = "Endothelial cell (lymphatic)", 
                                         "Vascular Endothelial Cells" = "Endothelial cell (vascular)", 
                                         "Fibroblasts" = "Fibroblast",
                                         "Basal Keratinocytes" = "Epithelial cell (basal keratinocyte)", 
                                         "Suprabasal Keratinocytes" = "Epithelial cell (suprabasal keratinocyte)", 
                                         "Melanocytes" = "Melanocyte"))
  df = add.sc.annos(df, expression.annos = JA.expression.annos, annos.name = "Jerby_Arnon",
                    cell.type.filter = c("CD4+ T Cells" = "T.CD4", 
                                         "CD8+ T Cells" = "T.CD8",
                                         "NK Cells" = "NK",
                                         "B Cells" = "B.cell",
                                         "Macrophages" = "Macrophage",
                                         "Endothelial Cells" = "Endo.", 
                                         "CAFs" = "CAF",
                                         "Malignant Cells" = "Mal"))
  df = add.sc.annos(df, expression.annos = StubS.expression.annos, annos.name = "Stubenvoll_Schmidt",
                    cell.type.filter = c("Cytotoxic T Cells" = "Tcyt-activated", 
                                         "Effector-memory T cells" = "Teff-mem",
                                         "Regulatory T cells" = "Treg",
                                         "NK Cells" = "NK",
                                         "B Cells" = "B-cells",
                                         "Dendritic-cells" = "Dendritic-cells",
                                         "Macrophages" = "Macrophages",
                                         "Endothelial Cells" = "Endothelial-cells",
                                         "Keratinocytes" = "Keratinocytes",
                                         "Fibroblasts" = "Fibroblasts",
                                         "Melanoma (melanocytic)" = "Mel-trans-melan",
                                         "Melanoma (transitory)" = "Mel-trans",
                                         "Melanoma (nc-like)" = "Mel-nc-like"))
  return(df)
}

# long dash: "  —  "


```

## Samples of all melanomas that will be used in the paper.

```{r all-mela-samples, results='asis', echo=T, eval=T}
for (sample in all_mela_samples){
  sample_full_name = manifest[[sample]]$full_name
  cat("* ", sample, " (", sample_full_name, ") \n", sep = "")
}
cat("\n")
```

```{r all-melas-samples-prep, echo=F, eval=T}
save_table(tabl = (joint_melas_df), filename = "data/LRI_results_melanomas_bare.csv")

# remove all samples not significant in any sample for the annotated output table, and fill missing
annotated_malanomas_table = add.annos(joint_melas_df) %>%
  mutate(across(contains("significant_in"), .fn = ~ replace_na(.x, F)),
         across(contains("pval_"), .fn = ~ replace_na(.x, 1)),
         across(contains("zscore_"), .fn = ~ replace_na(.x, 0)),
         across(contains("log2FC_"), .fn = ~ replace_na(.x, 0)),
         summed_zscores = rowSums(across(contains("zscore_")), na.rm = F)) %>%
  filter(count_signif > 0)
save_table(tabl = annotated_malanomas_table, filename = "data/LRI_results_melanomas_with_annos.csv.gz")

# subset to mechansisms significant in >50% of samples
joint_melas_df_50pct_signif_positive = joint_melas_df %>%
  filter(count_signif > (0.5*length(all_mela_dfs)))
joint_melas_df_all_signif_positive = joint_melas_df %>%
  filter(count_signif == length(all_mela_dfs))
```

```{r all-mela-samples-asps, results='asis', echo=T, eval=T}
pct50_asp = 0.5 + (nrow(joint_melas_df_50pct_signif_positive)/100)
pct100_asp = 0.5 + (nrow(joint_melas_df_all_signif_positive)/100)

cat("\n## ", Niches_normalization_used, ": ", Niches_result_used, "\n")
```

## Significant in all of our samples

Significant in all samples, annotated with Stubenvoll & Schmidt expression data:

```{r vis2-a-allsignificant-heatmap, fig.width=10.0, fig.asp=pct100_asp, echo=T, eval=nrow(joint_melas_df_all_signif_positive)>0}
pct100_data = add.annos(joint_melas_df_all_signif_positive)
plot_combined_df(pct100_data, plot_TCGA = "quant", plot_GTEx = "none", plot_JA = "none", plot_StubS = "quant",
                 legends.rel.heights = 0.1/pct100_data)
```

```{r vis2-b-allsignificant-circos, fig.width=10.0, fig.asp=1., echo=T, eval=nrow(joint_melas_df_all_signif_positive)>0}
circle.vis.LRI(pct100_data)
```

```{r vis2-c-allsignificant-table, eval=nrow(joint_melas_df_all_signif_positive)>0}
render_table(joint_melas_df_all_signif_positive)
```

## Filtered and annotated with external samples:

Excluded: HLA-_, CD4, CD8_, CD3_ containing mechanisms

```{r vis-filt-a-50pctsignificant-heatmap, fig.width=10.0, fig.asp=pct50_asp, echo=T, eval=nrow(joint_melas_df_50pct_signif_positive)>0}
pct50_data = add.annos(joint_melas_df_50pct_signif_positive)
exclude_regexes = c("^HLA-", "^CD4$", "^CD8", "^CD3") 
# exclude if any of the above occurs in either source or target column
pct50_data_excludes = apply(sapply(exclude_regexes, function(r)str_detect(pct50_data$target_genesymbol, pattern = r)), FUN = any, MARGIN = 1) | apply(sapply(exclude_regexes, function(r)str_detect(pct50_data$source_genesymbol, pattern = r)), FUN = any, MARGIN = 1)
pct50_data_df_wo_excluded = pct50_data[!pct50_data_excludes,]

plot_combined_df(pct50_data_df_wo_excluded, anno.signifi.legend = "plot", plot_StubS = "quant",
                 order.signif.overwrite = T, legends.rel.heights = 0.1/pct50_asp, 
                 plot_additional_dfs_list = list(joint_naevi_df, joint_externals_df))
```

```{r vis-filt-b-50pctsignificant-circos, fig.width=10.0, fig.asp=1., echo=T, eval=nrow(joint_melas_df_50pct_signif_positive)>0}
circle.vis.LRI(pct50_data_df_wo_excluded)
```

```{r vis-filt-c-50pctsignificant-table, eval=nrow(joint_melas_df_50pct_signif_positive)>0}
render_table(pct50_data_df_wo_excluded)
```

# External Samples

## Pozniak et al. 2024 samples

```{r external-samples, results='asis', echo=T, eval=T}
for (sample in all_external_samples){
  sample_full_name = manifest[[sample]]$full_name
  cat("* ", sample, " (", sample_full_name, ") \n", sep = "")
}
cat("\n")
```

```{r external-samples-prep, echo=F, eval=T}
samples_50pct = (0.5*length(all_external_samples))

save_table(tabl = (joint_externals_df), filename = "data/LRI_results_pozniak_bare.csv")
# remove all samples not significant in any sample for the annotated output table
joint_externals_df_anno = add.annos(joint_externals_df) %>%
  mutate(across(contains("significant_in"), .fn = ~ replace_na(.x, F)),
         across(contains("pval_"), .fn = ~ replace_na(.x, 1)),
         across(contains("zscore_"), .fn = ~ replace_na(.x, 0)),
         across(contains("log2FC_"), .fn = ~ replace_na(.x, 0)),
         summed_zscores = rowSums(across(contains("zscore_")), na.rm = F)) %>%
  filter(count_signif > 0)
save_table(tabl = joint_externals_df_anno, filename = "data/LRI_results_pozniak_with_annos.csv.gz")
# as before 
externals_50pct_signif_positive = joint_externals_df %>% filter(count_signif>samples_50pct)
```

```{r external-samples-asps, results='asis', echo=T, eval=T}
external_anno_df = add.annos(externals_50pct_signif_positive)

exclude_regexes = c("^HLA-", "^CD4$", "^CD8", "^CD3") 
external_df_excludes = apply(sapply(exclude_regexes, function(r)str_detect(external_anno_df$target_genesymbol, pattern = r)), FUN = any, MARGIN = 1) | apply(sapply(exclude_regexes, function(r)str_detect(external_anno_df$source_genesymbol, pattern = r)), FUN = any, MARGIN = 1)
external_anno_df_wo_excluded = external_anno_df[!external_df_excludes,]

pct50_asp = 0.5 + (nrow(external_anno_df_wo_excluded)/100)

cat("\n## ", Niches_normalization_used, ": ", Niches_result_used, "\n")
```

## Significant in > 50%

Excluded: HLA-_, CD4, CD8_, CD3_ containing mechanisms

```{r vis1-external-heatmap, fig.width=10.0, fig.asp=pct50_asp, echo=T, eval=nrow(externals_50pct_signif_positive)>0}
plot_combined_df(external_anno_df_wo_excluded, plot_TCGA = "quant", plot_GTEx = "none", plot_JA = "quant", 
                 plot_signif = F, legends.rel.heights = 0.1/pct50_asp)
```

```{r vis1-external-circos, fig.width=10.0, fig.asp=1., echo=T, eval=nrow(externals_50pct_signif_positive)>0}
circle.vis.LRI(external_anno_df_wo_excluded)
```

```{r vis1-external-table, eval=nrow(joint_melas_df_50pct_signif_positive)>0}
render_table(external_anno_df_wo_excluded)
```

# Naevi Samples

## Samples of compound naevi:

```{r naevi-samples, results='asis', echo=F, eval=T}
for (sample in all_naevus_samples){
  sample_full_name = manifest[[sample]]$full_name
  cat("* ", sample, " (", sample_full_name, ") \n", sep = "")
}
cat("\n")
naeviall_signif_positive=tibble()
pct100_asp=0
```

```{r naevi-samples-prep, echo=F, eval=length(all_naevus_dfs)>0}
save_table(tabl = (joint_naevi_df), filename = "data/LRI_results_nevi_bare.csv")
# remove all samples not significant in any sample for the annotated output table
annotated_nevi_table = add.annos(joint_naevi_df) %>%
  mutate(across(contains("significant_in"), .fn = ~ replace_na(.x, F)),
         across(contains("pval_"), .fn = ~ replace_na(.x, 1)),
         across(contains("zscore_"), .fn = ~ replace_na(.x, 0)),
         across(contains("log2FC_"), .fn = ~ replace_na(.x, 0)),
         summed_zscores = rowSums(across(contains("zscore_")), na.rm = F)) %>%
  filter(count_signif > 0)
save_table(tabl = annotated_nevi_table, filename = "data/LRI_results_nevi_with_annos.csv.gz")
# since we only have one naevus with any kind of signal, order by significance and take top50 only
order_by_col = which(startsWith(x = colnames(joint_naevi_df), prefix = "pval_adj"))
if (length(order_by_col)>1) stop("More than 1 naevi in analysis, remove Top50 filter again!")
naeviall_signif_positive = joint_naevi_df %>% 
  arrange(colnames(joint_naevi_df)[order_by_col]) %>%
  slice_head(n = 50)
  #filter(count_signif==length(joint_naevi_df)) %>%
```

```{r all-naevi-asps, results='asis', echo=T, eval=length(all_naevus_dfs)>0}
pct100_asp = 0.5 + (nrow(naeviall_signif_positive)/100)

cat("\n## ", Niches_normalization_used, ": ", Niches_result_used, "\n")
```

## Significant in all

```{r vis5-a, fig.width=10.0, fig.asp=pct100_asp, echo=T, eval=nrow(naeviall_signif_positive)>0}
full = add.annos(naeviall_signif_positive)
plot_combined_df(full, plot_TCGA = "quant", plot_GTEx = "none", plot_JA = "quant",
                 plot_signif = F, legends.rel.heights = 0.1/pct100_asp)
```

```{r vis5-b, fig.width=10.0, fig.asp=1., echo=T, eval=nrow(naeviall_signif_positive)>0}
circle.vis.LRI(full)
```

```{r vis5-c-allm1significant-table, eval=nrow(naeviall_signif_positive)>0}
render_table(naeviall_signif_positive)
```
