---
title: "Completed figure panels (High-res)"
output:
  workflowr::wflow_html:
    code_folding: hide
    fig_caption: yes
    number_sections: yes
    pandoc_args: []
    self_contained: yes
    smart: yes
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: no
      smooth_scroll: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, message=FALSE, warning=FALSE, echo=FALSE, cache=F}
knitr::opts_chunk$set(
  echo = FALSE,
  error = FALSE,
  message = FALSE,
  cache = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.align = 'center',
  fig.show = 'hold',
  autodep = TRUE,
  dpi = 450, 
  fig.retina = 1,
  size = 6.,
  dev = c('png')
)
```

This should contain all figure panels pieced together as they appear in the submission, such that font sizes and such are approximately constant. (At least all parts that can be programmatically generated)

```{r libs-and-functions, echo = F, message = FALSE, warning=FALSE}
library("Cairo")
library("tidyverse")
library("Seurat")
library("yaml")
library("ggplot2")
library("cowplot")
library("clustree")
library("naturalsort")
library("reshape2")
library("patchwork")
library("ggthemes")
library("ggnewscale")
library("kableExtra")
library("scico")
library("semla")
library("rlang")
library("scales")
library("ggrepel")
library("mixtools")
library("gridExtra")
library("circlize")
library("magick")
library("dplyr")

library("survival")
library("survminer")
library("OmnipathR")
library("car")
library("recount")
library("GenomicRanges")
library("SummarizedExperiment")
library("HDF5Array")


# bioc
library("SummarizedExperiment")
library("clusterProfiler")
library("NICHES")

# conflicts
library("conflicted")
conflicts_prefer(dplyr::select,
                 base::setdiff,
                 base::union,
                 dplyr::filter,
                 dplyr::desc,
                 base::intersect,
                 cowplot::align_plots)

# >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
# Functions
# >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
source("code/helper/Rmarkdown-style.R")
theme_set(mytheme(base_size = 5))

source("code/helper/FG_helper_NICHES.R")
source("code/helper/aggregated_LRI_helper.R")

standardize = function(z) {
  rowmean = apply(z, 1, mean, na.rm=TRUE)
  rowsd = apply(z, 1, sd, na.rm=TRUE)
  rv = sweep(z, 1, rowmean,"-")
  rv = sweep(rv, 1, rowsd, "/")
  return(rv)
}

# emulated ggplot2 default colors as explained here: https://stackoverflow.com/a/8197703
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

compute_tumor_bg_dens <- function(modelobj, valrange){
  model = modelobj$mixture.model
  t <- seq(min(valrange), max(valrange), len=300)
  # which component is the tumor density
  malmask = model$mu == max(model$mu)
  # all other are summed to a compound bg density
  bg_dens = dnormmixt(t, model$lambda[!malmask], model$mu[!malmask], model$sigma[!malmask])
  tumor_dens = dnormmixt(t, model$lambda[malmask], model$mu[malmask], model$sigma[malmask])
  return(data.frame(x = t, bg_dens = bg_dens, tumor_dens = tumor_dens))
}

# apply image magick functions to the images stored by semla.
# FUN needs to take a magick image and return a magick image of the same shape
# use e.g. magick::image_modulate(..., brightness = 120) for brightening or so
# this is just my internal shorthand, does not check for proper inputs
apply_magick_image_functions = function(se, FUN, image_slot = "transformed"){
  staff = se@tools$Staffli
  staff@rasterlists[[image_slot]] = lapply(staff@rasterlists[[image_slot]], function(rinp){
    im = magick::image_read(rinp)
    im_tr = FUN(im)
    return(grDevices::as.raster(im_tr))
  })
  se@tools$Staffli = staff
  return(se)
}

# load manifest
manifest = yaml.load_file("./assets/manifest.yaml")
figlab_hjust=1.
figlab_size=7
```

Author: Florian Große (Last updated: `r Sys.Date()`)

# Figure 2

```{r fig2-PM1-basics-prep, echo=F, eval=T}
# load PM1
current_sample = "PM1"
sample_manifest = manifest[[current_sample]]
default_plot_focus = sample_manifest$crop_area

source("code/helper/FG_plotting_functions.R")

p2scalebar = semla::scalebar(x = 1000, breaks = 3, highlight_breaks = c(1,3), breakheight = 2., 
                      title_position = "top", text_height = 25, linewidth = 0.2, padding.y = grid::unit(0, "mm"))
base_pt_size = estim_point_size(sample_manifest)*0.335
work.dir = file.path(manifest$workdata_parent, sample_manifest$workdata)
se.full = readRDS(file.path(work.dir, "01_seurat_full.rds"))
se.full = apply_magick_image_functions(se.full, function(x)image_average(c(image_contrast(x, sharpen = 1), image_normalize(x))))
obj1 = readRDS(file.path(work.dir, "09_seurat.rds"))
se = obj1$se
se = apply_magick_image_functions(se, function(x)image_average(c(image_contrast(x, sharpen = 1), image_normalize(x))))
obj.mrkrs.ora = readRDS(file.path(work.dir, "07_cluster_markers.rds"))
se$Cluster = se[[obj.mrkrs.ora$cl.res]]
Idents(se) = "Cluster"
cl.col = gg_color_hue(length(levels(se$Cluster)))
names(cl.col) = levels(se$Cluster)
```

```{r fig2-PM1-A, fig.width=2, fig.height=1.55, fig.show='hold', echo=F, eval=T}
# Histo anno overlay
p2A = feature_overlay(
    se.full,
    crop_area = default_plot_focus,
    pt.size = base_pt_size*1.5,
    features = "nFeature_Spatial",
    pt_alpha = 0, 
    add_scalebar = T,
    scalebar_position = c(0.05, 0.9),
    scalebar_height = 0.15,
    scalebar_gg = p2scalebar
) &
  ggtitle(NULL, subtitle = NULL) &
  theme(
    legend.position = "none",
    plot.margin = margin(1,1,1,1, "pt"),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )
# remove scalebar border
p2A[[1]][[2]] = p2A[[1]][[2]] + theme(panel.border = element_blank())
p2A
```

```{r fig2-PM1-B, fig.width=2, fig.height=1.55, fig.show='hold', echo=F, eval=T}
# sclusters on histo
p2B = feature_overlay_discrete(
  se, 
  features = "seurat_clusters",
  crop_area = default_plot_focus,
  pt.size = base_pt_size*1.5,
  pt.stroke = 0.1,
  add_scalebar = T,
  scalebar_position = c(0.05, 0.9),
  scalebar_height = 0.15,
  scalebar_gg = p2scalebar,
  colors = cl.col
  ) &
  theme(
    legend.text = element_text(size = unit(7, "pt")),
    legend.title = element_text(size = unit(6, "pt")),
    legend.key.height = unit(12, "pt"),
    legend.box.spacing = unit(4, "pt"),
    plot.margin = margin(1,1,1,1, "pt"),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "right"
  ) &
  guides(fill = guide_legend(title = "Cluster", ncol = 1, override.aes = list(size=2.5))) & 
  ggtitle(NULL, subtitle = NULL)
# remove scalebar border
p2B[[1]][[2]] = p2B[[1]][[2]] + theme(panel.border = element_blank())
p2B[[1]][[3]] = p2B[[1]][[3]] + theme(panel.border = element_blank()) # for whatever reason, with the current semla version this exists twice
p2B
```

```{r fig2-PM1-basics-mid-prep, echo=F, eval=T}
# prepare score plots for middle row
plotmax_T = max(quantile(unlist(se[["T.cells_logcounts"]]), probs = 0.995), 0.4)
plotmax_NK = max(quantile(unlist(se[["NK.cells_logcounts"]]), probs = 0.995), 0.4)
plotmax_Mel = max(quantile(unlist(se[["melanocytes.malignant.cells_logcounts"]]), probs = 0.995), 0.4)

p2_scoreplot <- function(se, scorename, title = "", limits, 
                        scalename, crop_area, pt.size, pt.stroke = 0.1){
  pmain = feature_overlay( se, scorename, titles = title, pt.stroke = pt.stroke,
                           limits = limits, crop_area = crop_area, pt.size = pt.size,
                           add_scalebar = T, scalebar_position = c(0.05, 0.85), 
                           scalebar_height = 0.15, scalebar_gg = p2scalebar,
                           oob = scales::squish, scalename = scalename) & 
    theme( legend.key.width = unit(0.25, "lines"),
           legend.key.height  = unit(0.5, "lines"),
           plot.margin = margin(0,0,0,0, "pt"), 
           legend.box.spacing = unit(2, "pt"),
           legend.box.margin = margin(0,0,0,0, "pt"), 
           legend.margin = margin(2,1,2,0, "pt"), 
           plot.title = element_text(size = unit(6, "pt"), 
                                     margin = margin(1,1,1,1,"pt"), vjust = 1),
           legend.text = element_text(size = unit(5, "pt")),
           legend.title = element_text(size = unit(5, "pt"), 
                                       margin = margin(1,1,2,1, unit = "pt")))
  pmain[[1]][[2]] = pmain[[1]][[2]] + theme(panel.border = element_blank())
  pmain[[1]][[3]] = pmain[[1]][[3]] + theme(panel.border = element_blank())
  return(pmain)
}

tscore_range = 1.1*range(se$T.cells_logcounts)
malscore_range = pmax(1.1*range(se$melanocytes.malignant.cells_logcounts), c(-0.55, 0.5)) # not lower than -0.5 plus padding
deconv_range_max = max(c(se$`CARD_decon_mean_T/NK_cells`, se$CARD_decon_mean_Melanocytic_cells))+0.01
tplots_dat = tibble(tnk_decon = se$`CARD_decon_mean_T/NK_cells`, 
              t_score = se$T.cells_logcounts, 
              act = se$T.positive)
nkplots_dat = tibble(tnk_decon = se$`CARD_decon_mean_T/NK_cells`, 
              nk_score = se$NK.cells_logcounts, 
              act = se$NK.positive)
malplots_dat = tibble(mal_decon = se$CARD_decon_mean_Melanocytic_cells, 
              mal_score = se$melanocytes.malignant.cells_logcounts, 
              act = se$TUMOR.positive)
malmodel_bg_dens = compute_tumor_bg_dens(Misc(se, "mal.model"), malscore_range)

p2CDs = align_plots(
  p2_scoreplot(se, "CARD_decon_mean_Melanocytic_cells", limits = c(0., NA), 
               crop_area = default_plot_focus, pt.size = base_pt_size,
               scalename = "Cell Type\nProportion", "Melanoma cells deconvolution"),
  p2_scoreplot(se, "CARD_decon_mean_T/NK_cells", limits = c(0., NA), 
               crop_area = default_plot_focus, pt.size = base_pt_size,
               scalename = "Cell Type\nProportion", title = "T/NK deconvolution"),
  p2_scoreplot(se, "melanocytes.malignant.cells_logcounts", limits = c(0., plotmax_Mel), 
               crop_area = default_plot_focus, pt.size = base_pt_size,
               scalename = "Module\nScore", title = "Melanoma cells score"),
  p2_scoreplot(se, "T.cells_logcounts", limits = c(0., plotmax_T), 
               crop_area = default_plot_focus, pt.size = base_pt_size,
               scalename = "Module\nScore", title = "T cells score"),
  p2_scoreplot(se, "NK.cells_logcounts", limits = c(0., plotmax_T), 
               crop_area = default_plot_focus, pt.size = base_pt_size,
               scalename = "Module\nScore", title = "NK cells score"))

histo_mal_score = ggplot(malplots_dat, aes(mal_score)) + 
    geom_histogram(mapping = aes(y = after_stat(density)), bins = 70, 
                   fill = "white", colour = "grey30", linewidth = 0.1) + 
    geom_vline(xintercept = Misc(se, "mal.model")$mixture.model.cutoff, 
               colour = "black", linewidth = 0.5) +
    geom_line(data = malmodel_bg_dens, mapping = aes(x,bg_dens), 
              color = "#708070", linewidth = 0.6) + 
    geom_line(data = malmodel_bg_dens, mapping = aes(x,tumor_dens), 
              color = "#AF4030", linewidth = 0.6) + 
    mytheme_grid(base_size = 5) + 
    theme(axis.title = element_text(size = unit(7, "pt"), margin = margin(2,2,2,2,"pt")),
          axis.text = element_text(size = unit(6, "pt"))) +
    coord_cartesian(xlim = malscore_range, 
                    expand = F) +
    xlab("Melanoma module score") + 
    ylab("Density")

histo_t_score = ggplot(tplots_dat, aes(t_score)) + 
    geom_histogram(mapping = aes(y = after_stat(density)), bins = 70, 
                   fill = "white", colour = "grey30", linewidth = 0.1) + 
    geom_vline(xintercept = Misc(se, "t.model")$fdr.corrected.background.cutoff, 
               colour = "black", linewidth = 0.5) +
    geom_line(data = tibble(x = seq(Misc(se, "t.model")$background.distr.model$location -
                                      8*Misc(se, "t.model")$background.distr.model$scale,
                                    Misc(se, "t.model")$background.distr.model$location +
                                      8*Misc(se, "t.model")$background.distr.model$scale, 
                                    len = 150),
                            dens = dnorm(
                              x=seq(Misc(se, "t.model")$background.distr.model$location -
                                      8*Misc(se, "t.model")$background.distr.model$scale,
                                    Misc(se, "t.model")$background.distr.model$location +
                                      8*Misc(se, "t.model")$background.distr.model$scale, 
                                    len = 150),
                              mean = Misc(se, "t.model")$background.distr.model$location,
                              sd = Misc(se, "t.model")$background.distr.model$scale)),
              aes(x,dens), color = "#40AF40", linewidth = 0.6) + # draw background distribution
    mytheme_grid(base_size = 5) + 
    theme(axis.title = element_text(size = unit(7, "pt"), margin = margin(2,2,2,2,"pt")),
          axis.text = element_text(size = unit(6, "pt"))) +
    coord_cartesian(xlim = tscore_range, expand = F) +
    xlab("T cell module score") + 
    ylab(" Density ")

histo_nk_score = ggplot(nkplots_dat, aes(nk_score)) + 
    geom_histogram(mapping = aes(y = after_stat(density)), bins = 70, 
                   fill = "white", colour = "grey30", linewidth = 0.1) + 
    geom_vline(xintercept = Misc(se, "nk.model")$fdr.corrected.background.cutoff, 
               colour = "black", linewidth = 0.5) +
    geom_line(data = tibble(x = seq(Misc(se, "nk.model")$background.distr.model$location -
                                      8*Misc(se, "nk.model")$background.distr.model$scale,
                                    Misc(se, "nk.model")$background.distr.model$location +
                                      8*Misc(se, "nk.model")$background.distr.model$scale, 
                                    len = 150),
                            dens = dnorm(
                              x=seq(Misc(se, "nk.model")$background.distr.model$location -
                                      8*Misc(se, "nk.model")$background.distr.model$scale,
                                    Misc(se, "nk.model")$background.distr.model$location +
                                      8*Misc(se, "nk.model")$background.distr.model$scale, 
                                    len = 150),
                              mean = Misc(se, "nk.model")$background.distr.model$location,
                              sd = Misc(se, "nk.model")$background.distr.model$scale)),
              aes(x,dens), color = "#40AF40", linewidth = 0.6) + # draw background distribution
    mytheme_grid(base_size = 5) + 
    theme(axis.title = element_text(size = unit(7, "pt"), margin = margin(2,2,2,2,"pt")),
          axis.text = element_text(size = unit(6, "pt"))) +
    coord_cartesian(xlim = tscore_range, expand = F) +
    xlab("NK cell module score") + 
    ylab(" Density ")

malmodel_scatter = ggplot(malplots_dat, aes(mal_score, mal_decon, color = act)) + 
    geom_hline(yintercept = 0.1, colour = "black", linewidth = 0.5) +
    geom_vline(xintercept = Misc(se, "mal.model")$mixture.model.cutoff, colour = "black", linewidth = 0.5) +
    geom_point(size = 0.2) + 
    scale_color_manual(values = c("False" = "grey", "True" = "#AF4030"), guide = "none") +
    mytheme_grid(base_size = 5) + 
    theme(axis.title = element_text(size = unit(7, "pt"), margin = margin(2,2,2,2,"pt")),
          axis.text = element_text(size = unit(6, "pt"))) +
    coord_cartesian(xlim = malscore_range, 
                    ylim = c(0., deconv_range_max), 
                    expand = F) +
    ylab("Deconv. melanoma content") +
    xlab("Melanoma module score")

tmodel_scatter = ggplot(tplots_dat, aes(t_score, tnk_decon, color = act)) + 
    geom_hline(yintercept = 0.1, colour = "black", linewidth = 0.5) +
    geom_vline(xintercept = Misc(se, "t.model")$fdr.corrected.background.cutoff, 
               colour = "black", linewidth = 0.5) +
    geom_point(size = 0.2) + 
    scale_color_manual(values = c("False" = "grey", "True" = "#40AF40"), guide = "none") +
    mytheme_grid(base_size = 5) + 
    theme(axis.title = element_text(size = unit(7, "pt"), margin = margin(2,2,2,2,"pt")),
          axis.text = element_text(size = unit(6, "pt"))) +
    coord_cartesian(xlim = tscore_range, ylim = c(0., deconv_range_max), expand = F) +
    ylab("Deconv. T/NK content") +
    xlab("T cell module score")

nkmodel_scatter = ggplot(tplots_dat, aes(t_score, tnk_decon, color = act)) + 
    geom_hline(yintercept = 0.1, colour = "black", linewidth = 0.5) +
    geom_vline(xintercept = Misc(se, "nk.model")$fdr.corrected.background.cutoff, 
               colour = "black", linewidth = 0.5) +
    geom_point(size = 0.2) + 
    scale_color_manual(values = c("False" = "grey", "True" = "#40AF40"), guide = "none") +
    mytheme_grid(base_size = 5) + 
    theme(axis.title = element_text(size = unit(7, "pt"), margin = margin(2,2,2,2,"pt")),
          axis.text = element_text(size = unit(6, "pt"))) +
    coord_cartesian(xlim = tscore_range, ylim = c(0., deconv_range_max), expand = F) +
    ylab("Deconv. T/NK content") +
    xlab("NK cell module score")

# nkmodel_scatter = ggplot(nkplots_dat, aes(nk_score, tnk_decon, color = act)) + 
#     geom_hline(yintercept = 0.1, colour = "black", linewidth = 1) +
#     geom_vline(xintercept = Misc(se, "nk.model")$fdr.corrected.background.cutoff, 
#                colour = "black", linewidth = 0.5) +
#     geom_point() + 
#     scale_color_manual(values = c("False" = "grey", "True" = "#40AF40"), guide = "none") +
#     mytheme_grid(base_size = 5) + 
#     theme(axis.title = element_text(size = unit(7, "pt"), margin = margin(2,2,2,2,"pt")),
#           axis.text = element_text(size = unit(6, "pt"))) +
#     coord_cartesian(xlim = 1.1*range(se$NK.cells_logcounts), 
#                     ylim = c(0., deconv_range_max), 
#                     expand = F) +
#     ylab(" Deconv. T/NK content ") +
#     xlab("NK cell module score") +
#     scale_y_continuous(sec.axis = dup_axis(labels = waiver())) +
#     theme(axis.title.y.right = element_blank(),
#           axis.ticks.y.left = element_blank(), 
#           axis.text.y.left = element_text(colour = "white", size = rel(0.5)))

p2EFs = align_patches(histo_mal_score, histo_t_score, histo_nk_score,
                      malmodel_scatter, tmodel_scatter, nkmodel_scatter)
```

```{r fig2-PM1-C, fig.width=1.41, fig.height=1.32, fig.show='hold', echo=F, eval=T}
plot(p2CDs[[1]])
plot(p2CDs[[2]])
```

```{r fig2-PM1-D, fig.width=1.41, fig.height=1.32, fig.show='hold', echo=F, eval=T}
plot(p2CDs[[3]])
plot(p2CDs[[4]])
plot(p2CDs[[5]])
```

```{r fig2-PM1-E, fig.width=1.41, fig.height=1.32, fig.show='hold', echo=F, eval=T}
p2EFs[[1]]
p2EFs[[2]]
p2EFs[[3]]
```

```{r fig2-PM1-F, fig.width=1.41, fig.height=1.32, fig.show='hold', echo=F, eval=T}
p2EFs[[4]]
p2EFs[[5]]
p2EFs[[6]]
```

```{r fig2-PM1-G, fig.width=2.8, fig.height=1.55, fig.show='hold', echo=F, eval=T}
p2G = feature_overlay_discrete(se, pt.size = base_pt_size, pt.stroke = 0.1,
                               crop_area = default_plot_focus,
                               features = "active.region.annotated",
                               titles = "",
                               colors = c("Double-negative spots" = "black",
                                          "Interfacing tumor-positive spots" = "red",
                                          "Interfacing immune-positive spots" = "green",
                                          "Non-interfacing tumor-positive spots" = "darkred",
                                          "Non-interfacing immune-positive spots" = "darkgreen",
                                          "Interfacing double-positive spots" = "yellow"),
                               pt_alpha = ifelse(se[["active.region"]]=="Background", 0., 1.),
                               add_scalebar = T, scalebar_position = c(0.05, 0.85),
                               scalebar_height = 0.18, scalebar_gg = p2scalebar) &
  theme(legend.title = element_blank(), 
        legend.position = "right",
        legend.key.height = unit(10, "pt"), 
        legend.box.spacing = unit(4, "pt"),
        legend.text = element_text(size = unit(5.5, "pt")), 
        plot.margin = margin(1,1,1,1, "pt"),
        plot.caption = element_blank()) &
  guides(fill=guide_legend(ncol = 1, override.aes = list(size = 2.5)))
p2G[[1]][[1]] = p2G[[1]][[1]] & labs(caption="Resulting spot choice for LR-analysis")
p2G[[1]][[2]] = p2G[[1]][[2]] + theme(panel.border = element_blank())
p2G[[1]][[3]] = p2G[[1]][[3]] + theme(panel.border = element_blank())

p2G
```

```{r fig2-PM1-H, fig.width=2.8, fig.height=1.55, fig.show='hold', echo=F, eval=T}
ggdraw() + draw_image("./assets/IF_pictures_for_merged_panels/Fig2_PM1_IF.png")
```


# Figure 3


```{r fig3-visium-combined-results-prep, echo=F, eval=T}
# prep LRI results
LRI_samples_used = setdiff(union(manifest$samples, manifest$external_samples), manifest$rejected_samples)
LRI_samples_used = LRI_samples_used[sapply(LRI_samples_used,
                                           function(sample_id) file.exists(file.path(manifest$workdata_parent, 
                                                                                     manifest[[sample_id]]$workdata, 
                                                                                     "09_NICHES_LRDE_res.rds")))]

# results table for LRI + circos plot
which.result = list(CTN = "border.markers.CellToNeighborhood",
                    NTC = "border.markers.NeighborhoodToCell")[[manifest$Niches_result_used]]
Niches_normalization_used = manifest$Niches_normalization_used
sample_annos = sapply(LRI_samples_used, function(sample_id){
  sample_manifest = manifest[[sample_id]]
  return(sample_manifest$type)
})
sample_anno_colmap = setNames(c(tol9qualitative[1:length(unique(sample_annos))], "black"),
                              nm = c(unique(sample_annos), NA))
LRDEs = lapply(setNames(nm=LRI_samples_used), function(sample_id){
  sample_manifest = manifest[[sample_id]]
  work.dir = file.path(manifest$workdata_parent, sample_manifest$workdata)
  resfile = file.path(work.dir, "09_NICHES_LRDE_res.rds")
  return(readRDS(resfile))
})
melanoma_samples = LRDEs[LRI_samples_used %in% manifest$primary_melanoma_samples]
melanoma_df = make_consensus_df(melanoma_samples, 
                                results.name = which.result, 
                                results.type = Niches_normalization_used)
melanoma_df = annotate_df_with_significance_info(melanoma_df)
pozniak_samples = LRDEs[LRI_samples_used %in% manifest$external_samples]
pozniak_df = make_consensus_df(pozniak_samples, 
                                results.name = which.result, 
                                results.type = Niches_normalization_used)
pozniak_df = annotate_df_with_significance_info(pozniak_df)

TCGA_ref = make_TCGA.se(cohort = "TCGA-SKCM", sites = c("Primary Tumor"))
StubS.expression.annos = readRDS(file.path(manifest$workdata_parent, "stubs.se_LRI_gene_expression_annos.Rds"))
# show only mechanisms significant in all samples
melanoma_df_filt = melanoma_df %>% filter(count_signif>2)
melanoma_df_filt = add.TCGA.annos(melanoma_df_filt, TCGA_reference = TCGA_ref)
melanoma_df_filt = add.sc.annos(melanoma_df_filt, expression.annos = StubS.expression.annos, annos.name = "Stubenvoll_Schmidt",
                    cell.type.filter = c("Cytotoxic T Cells" = "Tcyt-activated", 
                                         "Effector-memory T cells" = "Teff-mem",
                                         "Regulatory T cells" = "Treg",
                                         "NK Cells" = "NK",
                                         "B Cells" = "B-cells",
                                         "Dendritic-cells" = "Dendritic-cells",
                                         "Macrophages" = "Macrophages",
                                         "Endothelial Cells" = "Endothelial-cells",
                                         "Keratinocytes" = "Keratinocytes",
                                         "Fibroblasts" = "Fibroblasts",
                                         "Melanoma (melanocytic)" = "Mel-trans-melan",
                                         "Melanoma (transitory)" = "Mel-trans",
                                         "Melanoma (nc-like)" = "Mel-nc-like"))


# for already loaded PM1, add NICHES
se = add_NICHES_to_se(se, obj1$NICHES_result$logcounts)

# First row: border regions used for each sample
samples_to_show = setdiff(intersect(c(manifest$primary_melanoma_samples, manifest$external_samples), LRI_samples_used), "PM1")

p3_objs = lapply(setNames(nm=samples_to_show), function(sample){
  sample_manifest = manifest[[sample]]
  plt_crop = sample_manifest$crop_area
  base_pt_size = estim_point_size(sample_manifest)*0.33
  work.dir = file.path(manifest$workdata_parent, sample_manifest$workdata)
  obj = readRDS(file.path(work.dir, "09_seurat.rds"))
  se = obj$se
  se = apply_magick_image_functions(se, function(x)image_average(c(image_contrast(x, sharpen = 1), image_normalize(x))))
  se = add_NICHES_to_se(se, obj$NICHES_result$logcounts)
  return(list(plt_crop = plt_crop,
              base_pt_size = base_pt_size,
              se = se,
              NICHES_result = obj$NICHES_result))
})

pAE_widths = sapply(p3_objs, function(obj){
  crop = obj$plt_crop
  iminfo = obj$se@tools$Staffli@image_info
  w = (crop[3] * iminfo$width) - (crop[1] * iminfo$width)
  h = (crop[4] * iminfo$height) - (crop[2] * iminfo$height)
  return(w/h)
})

# change: for plotting, exchange _ in complexes for +, easier to understand 
melanoma_df_filt$mechanism = gsub(x = melanoma_df_filt$mechanism, "_", "+")
pozniak_df$mechanism = gsub(x = pozniak_df$mechanism, "_", "+")

```

```{r fig3-visium-combined-results-a, fig.width=7, fig.height=1.25, fig.show = 'hold', echo=F, eval=T}

p3As = lapply(names(p3_objs), function(sample_id){
  obj = p3_objs[[sample_id]]
  rel_width = pAE_widths[sample_id]
  se = obj$se
  p = feature_overlay_discrete(se, pt.size = rel_width*obj$base_pt_size,
                               pt.stroke = 0.05, crop_area = obj$plt_crop, titles = paste0("  ", sample_id),
                               features = "active.region.annotated",
                               colors = c("Double-negative spots" = "black",
                                  "Interfacing tumor-positive spots" = "red",
                                  "Interfacing immune-positive spots" = "green",
                                  "Non-interfacing tumor-positive spots" = "darkred",
                                  "Non-interfacing immune-positive spots" = "darkgreen",
                                  "Interfacing double-positive spots" = "yellow"),
                              pt_alpha = ifelse(se[["active.region"]]=="Background", 0., 1.),
                              add_scalebar = T, 
                              scalebar_position = c(0.03, 0.92), 
                 scalebar_height = 0.16, scalebar_gg = p2scalebar) &
    guides(fill = guide_legend(override.aes = list(size = 2), nrow = 1)) &
  theme(legend.title = element_blank(), 
        plot.title = element_text(size = 6, vjust = 0, 
                                  margin = margin(b=1.5, unit = "pt")),
        legend.position = "bottom", 
        legend.text = element_text(size = 5),
        legend.key.spacing = unit(1, "pt"),
        legend.key.size = unit(4.5, "pt"),
        plot.margin = margin(0,0,0,0, "pt"), 
        legend.margin = margin(0,0,0,0, "pt"), 
        legend.box.margin = margin(0,0,0,0, "pt"))
  
  
  p[[1]][[2]] = p[[1]][[2]] + theme(panel.border = element_blank())
  p[[1]][[3]] = p[[1]][[3]] + theme(panel.border = element_blank())
  return(p)
})
# cowplot struggles with this layered plot legend
p3Aleg_panelnames = cowplot::plot_component_names(p3As[[2]][[1]])
p3Aleg = cowplot::plot_components(p3As[[2]][[1]])[[which(str_starts(p3Aleg_panelnames, "guide-box-bottom"))]]

plot_grid(
  plot_grid(plotlist = lapply(p3As, function(x)x&theme(legend.position = "none")), 
          nrow = 1, rel_widths = pAE_widths),
  p3Aleg, ncol = 1, rel_heights = c(13,1)
)
```

```{r fig3-visium-combined-results-b, fig.width=7*0.98, fig.height=3.3, fig.show = 'hold', echo=F, eval=T}
# 3B: LRI consensus results table
p3B = plot_combined_df(melanoma_df_filt %>% filter(count_signif==4), 
                       anno.signifi.legend = "none", legend_rows = 1,
                       width_StubS = 12, plot_signif = F,
                       legend_barwidth = 4.5,
                       plot_StubS = "quant", legends.rel.heights = 0.025, 
                       plot_additional_dfs_list = list(pozniak_df)) & 
  theme(axis.text = element_text(size = unit(5, "pt")),
        legend.text = element_text(size = unit(5, "pt")),
        title = element_text(size = unit(6, "pt")),
        legend.title = element_text(size = unit(6, "pt"), hjust = 0.1),
        strip.text =  element_text(size = unit(6, "pt")), 
        legend.box.margin = margin(1,25,2,1,"pt"))

p3B
```

```{r fig3-visium-combined-results-c, fig.width=3.5, fig.height=3.5, fig.show = 'hold', echo=F, eval=T}
# p3C: circos
# for circleplot
# exclude if any of the above occurs in either source or target column
exclude_regexes = c("^HLA-", "^CD4$", "^CD8", "^CD3")
melanoma_df_filt_exludes = apply(sapply(exclude_regexes, function(r)str_detect(melanoma_df_filt$target_genesymbol, pattern = r)), FUN = any, MARGIN = 1) | apply(sapply(exclude_regexes, function(r)str_detect(melanoma_df_filt$source_genesymbol, pattern = r)), FUN = any, MARGIN = 1)
melanoma_df_filt_2 = melanoma_df_filt[!melanoma_df_filt_exludes,]

# 3-pc pca the distibution across celltypes and embed into rgb for colors
col.df.prep = rbind(melanoma_df_filt_2 %>%
              select(source_genesymbol, contains("source_Stubenvoll_Schmidt_norm_empirical_quantiles")) %>%
              rename_with(~gsub("source_", "", .x, ignore.case = T)),
            melanoma_df_filt_2 %>%
              select(target_genesymbol, contains("target_Stubenvoll_Schmidt_norm_empirical_quantiles")) %>%
              rename_with(~gsub("target_", "", .x, ignore.case = T))) %>%
  distinct()
minmax_std = function(xs, minval = 0.15, maxval = 0.75)(((xs-min(xs, na.rm = T))/(max(xs, na.rm = T)-min(xs, na.rm = T)))*(maxval-minval))+minval

set.seed(20122025) #fix umap
col.df = data.frame(genesymbol = col.df.prep$genesymbol,
                    as.data.frame(uwot::umap(col.df.prep %>% select(!genesymbol) %>% as.matrix(), n_components = 3, n_neighbors = round(nrow(col.df.prep)/3)))) %>%
  mutate(V1 = minmax_std(V1), V2 = minmax_std(V2), V3 = minmax_std(V3))
col.df = col.df %>% mutate(color = rgb(red = 1-V2, green = V1, blue = 1-V3)) 

#CairoPNG(filename = "public/circos_results.png", width = 3.5, height = 3.5, units = "in", dpi = 450)
circle.vis.LRI(melanoma_df_filt_2, fontsize = 0.33, text.offset = 0.05, text.height.mult = 0.1, small.gap = 1, big.gap = 4, colors = setNames(col.df$color, nm = col.df$genesymbol))
#dev.off()
```

```{r fig3-visium-combined-results-d, fig.width=3, fig.height=1.34, fig.show = 'hold', echo=F, eval=T}

p3_scoreplot <- function(se, feature, assay, title, scalename, crop, pt.size, 
                         leg_right = F, pt.stroke=0.06, colbar_width = 0.2, colbar_length = 0.6,
                         scalebar_position = c(0.05, 0.85), scalebar_height = 0.2){
  p = feature_overlay(se, feature, titles = title, assay = assay,
                 limits = c(0., NA), crop_area = crop, 
                 pt.size = pt.size, pt.stroke = pt.stroke,
                 add_scalebar = T, scalebar_position = scalebar_position, 
                 scalebar_height = scalebar_height, scalebar_gg = p2scalebar, 
                 colorbar_width = if(leg_right){unit(colbar_width, "lines")}else{unit(colbar_length, "lines")}, 
                 colorbar_height = if(leg_right){unit(colbar_length, "lines")}else{unit(colbar_width, "lines")},
                 gg_guide_colorbar = guide_colorbar(title.position = "top", title.vjust = 1., title = scalename),
                 oob = scales::squish) & 
  theme(legend.position = ifelse(leg_right, "right", "bottom"),
        plot.margin = margin(0,0,0,0, "pt"), 
        legend.margin = margin(1,1,1,1, "pt"),
        legend.box.spacing = unit(0, "pt"),
        legend.box.just = "top",
        legend.text = element_text(size = unit(5, "pt")),
        plot.title = element_text(size = unit(6, "pt"), margin = margin(2,1,1,1,"pt"), vjust = 1),
        legend.title = element_text(size = unit(6, "pt"), margin = margin(b = 2, unit = "pt")))
  p[[1]][[2]] = p[[1]][[2]] + theme(panel.border = element_blank())
  p[[1]][[3]] = p[[1]][[3]] + theme(panel.border = element_blank())
  return(p)
}

# 3D: LGALS1, PTPRC gene expression + combined NICHES score
p3D = plot_grid(nrow = 1, axis = "tb", align = "hv",
                p3_scoreplot(se, "LGALS1", "RNA_logcounts", "LGALS1", " \nLog(Counts+1)",
                             default_plot_focus, base_pt_size/1.2, leg_right=F),
                p3_scoreplot(se, "PTPRC", "RNA_logcounts", "PTPRC", " \nLog(Counts+1)",
                             default_plot_focus, base_pt_size/1.2, leg_right=F),
                p3_scoreplot(se, "LGALS1—PTPRC", "CellToNeighborhood", "LGALS1—PTPRC", "NICHES\ninteraction score", 
                             default_plot_focus, base_pt_size/1.1, leg_right=F)
  )

p3D
```

```{r fig3-visium-combined-results-e, fig.width=3.5, fig.height=2.5, fig.show = 'hold', echo=F, eval=T}
p3Es = lapply(names(p3_objs), function(sample_id){
  obj = p3_objs[[sample_id]]
  rel_width = pAE_widths[sample_id]
  se = obj$se
  p = p3_scoreplot(se, "LGALS1—PTPRC", "CellToNeighborhood", sample_id, NULL, 
                   scalebar_position = c(0.05, 0.83), scalebar_height = 0.13,
                   obj$plt_crop, rel_width*obj$base_pt_size/1.1)
  return(p)
})

plot_grid(plot_grid(plotlist = p3Es[1:3], nrow = 1, 
                    rel_widths = pAE_widths[1:3]),
          plot_grid(plotlist = p3Es[4:6], nrow = 1, 
                    rel_widths = pAE_widths[4:6]),
          ncol = 1, rel_heights = c(0.9, 1.0))

```

# Figure 6

```{r fig6-survival-prep, fig.show = 'hold', echo=F, eval=T}
TCGA.se = fetch_TCGA()
TCGA.se = TCGA.se[, TCGA.se$gdc_cases.project.project_id == "TCGA-SKCM"]

# removes "additional metastatic" normal tissue samples
sites = c("Primary Tumor", "Metastatic")
TCGA.se = TCGA.se[, TCGA.se$gdc_cases.samples.sample_type %in% sites]
colData(TCGA.se) = droplevels(colData(TCGA.se))

# removes trailing version numbers from ENSEMBL identifiers
rownames(TCGA.se) = gsub("\\..+", "", rownames(TCGA.se))

# load TCIA data 
TCIA.fractions= read_tsv("./assets/TCIA-CellTypeFractionsData_SKCM_only_march2024.tsv", col_names = c("Patient", "Disease", "Celltype", "quanTIseq_lsei_TIL10", "Cibersort_LM22"))

# should be good to go, but as sanity check retain only SKCM data
TCIA.fractions = TCIA.fractions %>%
  filter(Disease == "SKCM") %>%
  distinct()

TCIA.quanTIseq = transmute(TCIA.fractions, 
                           Patient = Patient, 
                           Celltype = Celltype,
                           Fraction = quanTIseq_lsei_TIL10) %>%
  pivot_wider(names_from = Celltype, values_from = Fraction) 

pat = TCIA.quanTIseq$Patient
TCIA.quanTIseq = TCIA.quanTIseq %>%
  select(!Patient) %>%
  rename_with(~ paste0("quanTIseq_", make.names(.x))) %>%
  as.data.frame()
rownames(TCIA.quanTIseq) = pat

TCIA.Cibersort = transmute(TCIA.fractions, 
                           Patient = Patient, 
                           Celltype = Celltype,
                           Fraction = Cibersort_LM22) %>%
  pivot_wider(names_from = Celltype, values_from = Fraction)

pat = TCIA.Cibersort$Patient
TCIA.Cibersort = TCIA.Cibersort %>%
  select(!Patient) %>%
  rename_with(~ paste0("cibersort_", make.names(.x))) %>%
  as.data.frame()
rownames(TCIA.Cibersort) = pat

TCIA.Cibersort = TCIA.Cibersort[,!(apply(is.na(TCIA.Cibersort), FUN = all, MARGIN = 2))]

# format and filter TCGA data, add TCIA data to TCGA pheno data
pheno = data.frame(colData(TCGA.se))
organ = unique(TCGA.se$gdc_cases.project.primary_site)

# remove columns containing only NAs
pheno =  pheno[, which(apply(pheno, 2, function(x) !all(is.na(x))))]

exprs.tpm = data.frame(assays(TCGA.se)$tpm, check.names = F)
exprs.mat = log1p(exprs.tpm)
var.nonzero = apply(exprs.mat, 1, function(x) { stats::var(x) > 0.0001 })
exprs.mat = exprs.mat[var.nonzero,]
exprs.mat.std = standardize(exprs.mat)

sampleIDs = rownames(pheno)
exprs.mat = exprs.mat[, colnames(exprs.mat) %in% sampleIDs]
exprs.mat.std = exprs.mat.std[, colnames(exprs.mat.std) %in% sampleIDs]
exprs.tpm = exprs.tpm[, colnames(exprs.tpm) %in% sampleIDs]
# Consistency check: Expression colnames are in the same order as pheno rownames
stopifnot(identical(rownames(pheno), colnames(exprs.mat)))
stopifnot(identical(rownames(pheno), colnames(exprs.tpm)))

# Combining time to event variables for surv_fu_days:
# - Time to death is in $gdc_cases.diagnoses.days_to_death
# - Death is indicated by non NA in $gdc_cases.diagnoses.days_to_death
#   (over $gdc_cases.diagnoses.vital_status, which contains 2 errors)
# - Follow up time is in $gdc_cases.diagnoses.days_to_last_follow_up
# - Negative entries in $gdc_cases.diagnoses.days_to_last_follow_up have been 
#   explained by the GDC Helpdesk in this issue: https://github.com/BioinformaticsFMRP/TCGAbiolinks/issues/324 

# surv_fu_days captures the time to event, therefore
# - DoD is indicated by non-NA value of $gdc_cases.diagnoses.days_to_death
# - time to event or censoring is the time to DoD if available and time to
#   last follow-up else. Negative or zero entries can then be discarded, since
#   they don't give any information about the survival/death of the patient 

pheno = cbind(pheno, 
              surv_fu_days = coalesce(pheno$gdc_cases.diagnoses.days_to_death,
                                      pheno$gdc_cases.diagnoses.days_to_last_follow_up),
              surv_deceased = !is.na(pheno$gdc_cases.diagnoses.days_to_death))
pheno$surv_fu_days[pheno$surv_fu_days <= 0] = NA # Changing uninformative time to event to "NA"

# sanity check 
stopifnot(identical(rownames(pheno), colnames(exprs.mat)))

which_not_na = which(!is.na(pheno$surv_fu_days))
pheno = pheno[which_not_na, ]
exprs.mat = exprs.mat[ ,which_not_na]
exprs.mat.std = exprs.mat.std[ ,which_not_na]
exprs.tpm = exprs.tpm[, which_not_na]

#sanity check
stopifnot(identical(rownames(pheno), colnames(exprs.mat)))
stopifnot(identical(rownames(pheno), colnames(exprs.tpm)))

# add TCIA Cybersort data
pat = pheno$cgc_case_id
pheno = cbind(pheno, TCIA.Cibersort[pat,])

# add TCIA quanTIseq data
pheno = cbind(pheno, TCIA.quanTIseq[pat,])

tcga_plt_dat = pheno %>%
  mutate(surv_fu_years = surv_fu_days/365,
         Sample_type = gdc_cases.samples.sample_type,
         ptprc = unlist(exprs.mat["ENSG00000081237",]),
         lgals1 = unlist(exprs.mat["ENSG00000100097",]),
         cd8t_median = quanTIseq_CD8.T.cells > median(quanTIseq_CD8.T.cells, na.rm = T))

## Gide et al. data
gide_tpm = readr::read_tsv("../../datasets/SRA_Downloads/gide_et_al_2019_melanoma/merged_output/quanTIseq_gene_tpm.txt")
gide_celltypes = readr::read_tsv("../../datasets/SRA_Downloads/gide_et_al_2019_melanoma/merged_output/quanTIseq_cell_fractions.txt")
Gide_meta_final = readr::read_tsv("./data/survival_data/Gide_et_al_metadata.tsv")

gide_pheno = Gide_meta_final %>% 
  dplyr::rename(Sample = Run) %>%
  left_join(gide_celltypes, by = "Sample") %>%
  mutate(surv_os = `Overall Survival (Days)`, 
         surv_deceased = `Last Followup Status` != "Alive",
         pretreatment_sample = str_detect(Sample_name, pattern = "PRE"))

gide_expr = data.frame(gide_tpm %>% select(!GENE))
rownames(gide_expr) = gide_tpm$GENE

gide_pheno$lgals1 = unlist(gide_expr["LGALS1", gide_pheno$Sample])
gide_pheno$ptprc = unlist(gide_expr["PTPRC", gide_pheno$Sample])

gide_plt_dat = gide_pheno %>%
  mutate(surv_fu_years = surv_os/365,
         PD1_Inhibitor = factor(ifelse(str_detect(Treatment, "Pembrolizumab"), 
                                       "Pembrolizumab", "Nivolumab"), 
                                   levels = c("Pembrolizumab", "Nivolumab")),
         CTLA4_Inhibitor = factor(ifelse(str_detect(Treatment, "Ipilimumab"), 
                                         "Ipilimumab", "None"), 
                                     levels = c("None", "Ipilimumab")),
         cd8t_median = T.cells.CD8 > median(T.cells.CD8, na.rm = T)) %>%
  as.data.frame()

# ggtheme similar to mytheme but for survplots
mysurvtheme = function(base_size = 12){
  half_line <- base_size/2
  theme_classic(base_size = base_size) &
    theme(
      axis.ticks.length = unit(half_line / 2.2, "pt"),
      strip.text.x = element_text(size = rel(1), colour = "black"),
      strip.text.y = element_text(size = rel(1), colour = "black"),
      strip.text = element_text(size = rel(1), colour = "black"),
      axis.text = element_text(size = rel(1), colour = "black"),
      axis.title = element_text(size = rel(1), colour = "black"),
      legend.title = element_text(colour = "black", size = rel(1)),
      legend.key.size = unit(1, "lines"),
      legend.text = element_text(size = rel(1), colour = "black"),
      legend.key = element_rect(colour = NA, fill = NA),
      legend.background = element_rect(colour = NA, fill = NA),
      plot.title = element_text(hjust = 0, face = "plain", colour = "black", size = rel(1)),
      plot.subtitle = element_text(colour = "black", size = rel(.85))
    )
}

# basic plots both

survival_extra_p1 = ggsurvplot_facet(survfit(Surv(surv_fu_years, surv_deceased) ~ cd8t_median, 
                                    data = tcga_plt_dat), ggtheme = mysurvtheme(),
                           conf.int = T, pval = T, pval.coord = c(0, 0.02), 
                           pval.method.coord =  c(0, 0.13), pval.method = T, 
                           data = tcga_plt_dat, facet.by = "Sample_type", ncol = 1) & 
            labs(x = "Time (years)", fill = "CD8 T cell content < median", 
                 color = "CD8 T cell content < median") & 
  theme(legend.position = "left", )

survival_extra_p2 = ggsurvplot_facet(survfit(Surv(surv_fu_years, surv_deceased) ~ cd8t_median, 
                                    data = gide_plt_dat), ggtheme = mysurvtheme(),
                           conf.int = T, pval = T, pval.coord = c(0, 0.02), 
                           pval.method.coord =  c(0, 0.13), pval.method = T, 
                           data = gide_plt_dat, facet.by = "PD1_Inhibitor", ncol = 1) & 
            labs(x = "Time (years)", fill = "CD8 T cell content < median", 
                 color = "CD8 T cell content < median") & 
  theme(legend.position = "left", )

# plot_grid(survival_extra_p1, NULL, survival_extra_p2, nrow = 1, labels = c("      TCGA-SKCM","", "      Gide et al."), 
#           rel_widths = c(1, 0.1, 1), label_fontface = "plain", label_size = 14, hjust = 0, vjust = 1.6)
#           

tcga_metas = tcga_plt_dat %>%
  filter(gdc_cases.samples.sample_type == "Metastatic" & quanTIseq_CD8.T.cells > 0) %>%
  mutate(CD8_T_cells = ifelse(quanTIseq_CD8.T.cells>median(quanTIseq_CD8.T.cells), "high", "low")) %>%
  mutate(PTPRC = ifelse(ptprc>median(ptprc), "high", "low"),
         LGALS1 = ifelse(lgals1>median(lgals1), "high", "low")) %>%
  ungroup() %>%
  as.data.frame()

gide_plt_subset = gide_plt_dat %>%
  filter(PD1_Inhibitor == "Pembrolizumab") %>%
  mutate(LGALS1 = ifelse(lgals1>median(lgals1, na.rm = T), "high", "low"),
         PTPRC = ifelse(ptprc>median(ptprc, na.rm = T), "high", "low")) %>%
  as.data.frame()

tcga_metas_cd8_split = tcga_plt_dat %>%
  filter(gdc_cases.samples.sample_type == "Metastatic" & quanTIseq_CD8.T.cells > 0) %>%
  mutate(CD8_T_cells = ifelse(quanTIseq_CD8.T.cells>median(quanTIseq_CD8.T.cells), "high", "low")) %>%
  group_by(CD8_T_cells) %>%
  mutate(PTPRC = ifelse(ptprc>median(ptprc), "high", "low"),
         LGALS1 = ifelse(lgals1>median(lgals1), "high", "low")) %>%
  ungroup() %>%
  as.data.frame()

```


```{r fig6-survival-a, fig.width=3.45*2, fig.height=1.75*2, fig.show = 'hold', out.width='50%', dpi=450/2, echo=F, eval=T}
# rendered in double size half resolution, because this survminer version has a problem with 
# changing teh text size of the p-value field -> quick fix: reduce size by half when inserting 

plot_grid(NULL, ggsurvplot_facet(survfit(Surv(surv_fu_years, surv_deceased) ~ PTPRC, 
                                         data = tcga_metas), pval.coord = c(0, 0.03), 
                                 pval.method.coord =  c(0, 0.15), pval = T, 
                                 size = 0.7, pval.method = T, data = tcga_metas, ggtheme = mysurvtheme(base_size = 12),
                                 facet.by = c("LGALS1"), nrow = 1, conf.int = T) & 
                    labs(x = "Time (years)", fill = "PTPRC", color = "PTPRC") & 
                    theme(legend.position = "left"), 
          ncol = 1, rel_heights = c(0.15, 2), label_size = 10, label_fontface = "plain", 
          hjust = 0, vjust = 1.6, labels = c("      TCGA-SKCM: filtered metastases, non-zero infiltration", ""))
```

```{r fig6-survival-b, fig.width=3.45*2, fig.height=1.75*2, fig.show = 'hold', out.width='50%', dpi=450/2, echo=F, eval=T}
plot_grid(NULL, ggsurvplot_facet(survfit(Surv(surv_fu_years, surv_deceased) ~ PTPRC, 
                                         data = gide_plt_subset), pval.coord = c(0, 0.03), 
                                 pval.method.coord =  c(0, 0.15), pval = T, 
                                 palette = RColorBrewer::brewer.pal(8, "Set2")[2:1],
                                 size = 0.7, pval.method = T, data = gide_plt_subset, 
                                 ggtheme = mysurvtheme(base_size = 12),
                                 facet.by = c("LGALS1"), nrow = 1, conf.int = T) & 
                    labs(x = "Time (years)", fill = "PTPRC", color = "PTPRC") & 
                    theme(legend.position = "left"), 
          ncol = 1, rel_heights = c(0.15, 2), label_size = 10, label_fontface = "plain", 
          hjust = 0, vjust = 1.6, labels = c("      Gide et al.: Pembrolizumab patients", ""))
```
```{r fig6-survival-c, fig.width=3.2*2, fig.height=2.6*2, fig.show = 'hold', out.width='50%', dpi=450/2, echo=F, eval=T}
plot_grid(NULL, ggsurvplot_facet(survfit(Surv(surv_fu_years, surv_deceased) ~ LGALS1, 
                              data = tcga_metas_cd8_split), pval.coord = c(0, 0.03), 
                           pval.method.coord =  c(0, 0.15), pval = T, size = 0.7, 
                      pval.method = T, data = tcga_metas_cd8_split, ggtheme = mysurvtheme(base_size = 12),
                      facet.by = c("CD8_T_cells", "PTPRC"), nrow = 1, conf.int = T) & 
            labs(x = "Time (years)", fill = "LGALS1", 
                 color = "LGALS1") & theme(legend.position = "left")
, ncol = 1, rel_heights = c(0.2, 1.6), 
            label_size = 10, label_fontface = "plain", hjust = 0, vjust = 1.6, 
            labels = c("      TCGA-SKCM:\tfiltered metastases, non-zero infiltration,\n\t\t\t\tLGALS1/PTPRC splits calculated within CD8 T cell subgroups", ""))
```

# Extended Figures

```{r supp-heatmap-gsea-prep}
other_samples_markers = lapply(setNames(nm=names(p3_objs)), 
                               function(sample){
  sample_manifest = manifest[[sample]]
  work.dir = file.path(manifest$workdata_parent, sample_manifest$workdata)
  obj.mrkrs.ora = readRDS(file.path(work.dir, "07_cluster_markers.rds"))
  se = p3_objs[[sample]]$se
  se.full = readRDS(file.path(work.dir, "01_seurat_full.rds"))
  se.full = apply_magick_image_functions(se.full,
                                         function(x)image_average(
                                           c(image_contrast(x, sharpen = 1),
                                             image_normalize(x))))

  se$Cluster = se[[obj.mrkrs.ora$cl.res]]
  Idents(se) = "Cluster"
  cl.col = gg_color_hue(length(levels(se$Cluster)))
  names(cl.col) = levels(se$Cluster)
  
  return(list(se = se,
              se.full = se.full,
              cl.col = cl.col,
              obj.mrkrs.ora = obj.mrkrs.ora,
              base_pt_size = p3_objs[[sample]]$base_pt_size,
              plt_crop = p3_objs[[sample]]$plt_crop))
})
```

## PM1

```{r supp-pm1-a, fig.width=3.5, fig.height=3.9, fig.show='hold', echo=F, eval=T, dpi=300}
# obj.mrkrs.ora
# cl.col

# Cluster marker
# markers = obj.mrkrs.ora$cluster.marker.wlcxn
# dgea.pvalue = obj.mrkrs.ora$dgea.pvalue
# 
# # ORA
# res.cP.cc.go = obj.mrkrs.ora$res.cP.cc.go
# res.cP.cc.go.sim = obj.mrkrs.ora$res.cP.cc.go.sim
# ora.oe = obj.mrkrs.ora$tops.l.oe
# ora.pvalue = obj.mrkrs.ora$ora.pvalue

markers.sign = subset(obj.mrkrs.ora$cluster.marker.wlcxn, p_val_adj < obj.mrkrs.ora$dgea.pvalue)
top = subset(markers.sign, (avg_log2FC > 0)) %>% group_by(cluster) %>% 
  mutate(logrank = dense_rank(desc(avg_log2FC)), pvalrank = dense_rank(p_val_adj)) %>%
  slice_min(logrank, n = 10)

DoHeatmap(se, features  = top$gene, raster = F, angle = 0, size=2.4, hjust=0.5, 
          draw.lines = F, assay = "Spatial", group.color = cl.col) +
  guides(color = "none") +
  scale_fill_scico(palette = "vik", end = 1, direction = 1, midpoint = 0.) +
  #mytheme() +
  theme(axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks=element_blank(),
        axis.title=element_blank(),
        line = element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        plot.background=element_blank(),
        legend.key.width = unit(0.5, "lines"),
        plot.margin = margin(1,1,-12,1, "pt"),
        axis.text.y.left = element_text(size = 4.9)) +
  labs(colour = "Cluster", fill = "Standardized\nexpression")
```

## PM2

```{r supp-pm2-a, fig.width=2, fig.height=1.4, fig.show='hold', echo=F, eval=T, dpi=300}
# Histo anno overlay
supp_histoplot = function(obj){
    p = feature_overlay(
      obj$se.full,
      crop_area = obj$plt_crop,
      pt.size = obj$base_pt_size*1.5,
      features = "nFeature_Spatial",
      pt_alpha = 0, 
      add_scalebar = T,
      scalebar_position = c(0.05, 0.9),
      scalebar_height = 0.15,
      scalebar_gg = p2scalebar
  ) &
    ggtitle(NULL, subtitle = NULL) &
    theme(
      legend.position = "none",
      plot.margin = margin(1,1,1,1, "pt"),
      axis.title = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank()
    )
  # remove scalebar border
  p[[1]][[2]] = p[[1]][[2]] + theme(panel.border = element_blank())
  p
}

supp_histoplot(other_samples_markers$PM2)
```

```{r supp-pm2-b, fig.width=2, fig.height=1.55, fig.show='hold', echo=F, eval=T, dpi=300}
# sclusters on histo
supp_clusterplot = function(obj){
  cl.col = gg_color_hue(length(levels(obj$se$seurat_clusters)))
  p = feature_overlay_discrete(
    obj$se, 
    features = "seurat_clusters",
    crop_area = obj$plt_crop,
    pt.size = obj$base_pt_size*1.5,
    pt.stroke = 0.1,
    add_scalebar = T,
    scalebar_position = c(0.05, 0.9),
    scalebar_height = 0.15,
    scalebar_gg = p2scalebar,
    colors = cl.col
    ) &
    theme(
      legend.text = element_text(size = unit(7, "pt")),
      legend.title = element_text(size = unit(6, "pt")),
      legend.key.height = unit(10, "pt"),
      legend.box.spacing = unit(4, "pt"),
      plot.margin = margin(1,1,1,1, "pt"),
      axis.title = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      legend.position = "right"
    ) &
    guides(fill = guide_legend(title = "Cluster", ncol = 1, override.aes = list(size=2.5))) & 
    ggtitle(NULL, subtitle = NULL)
  # remove scalebar border
  p[[1]][[2]] = p[[1]][[2]] + theme(panel.border = element_blank())
  p[[1]][[3]] = p[[1]][[3]] + theme(panel.border = element_blank())
  p
}

supp_clusterplot(p3_objs$PM2)
```

```{r supp-pm2-c, fig.width=3.5, fig.height=4.7, fig.show='hold', echo=F, eval=T, dpi=300}
supp_heatmap = function(obj, genesize = 4.8){
  markers.sign = subset(obj$obj.mrkrs.ora$cluster.marker.wlcxn, p_val_adj < obj$obj.mrkrs.ora$dgea.pvalue)
  top = subset(markers.sign, (avg_log2FC > 0)) %>% group_by(cluster) %>% 
    mutate(logrank = dense_rank(desc(avg_log2FC)), pvalrank = dense_rank(p_val_adj)) %>%
    slice_min(logrank, n = 10)
  
  p = DoHeatmap(obj$se, features  = top$gene, raster = F, angle = 0, size=2., 
                hjust=0.5, group.bar.height = 0.0175, 
                draw.lines = F, assay = "Spatial", group.color = obj$cl.col) +
    guides(color = "none") +
    scale_fill_scico(palette = "vik", end = 1, direction = 1, midpoint = 0.) +
    #mytheme() +
    theme(axis.line=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks=element_blank(),
          axis.title=element_blank(),
          line = element_blank(),
          panel.background=element_blank(),
          panel.border=element_blank(),
          plot.background=element_blank(),
          legend.key.width = unit(0.5, "lines"),
          plot.margin = margin(3,1,-12,1, "pt"),
          axis.text.y.left = element_text(size = genesize)) +
    labs(colour = "Cluster", fill = "Standardized\nexpression")
  p
}

supp_heatmap(other_samples_markers$PM2, genesize = 4.2)
```

```{r supp-pm2-h, fig.width=3., fig.height=2.6, fig.show = 'hold', echo=F, eval=T, dpi=300}
plt_deconv_and_scores_supp <- function(obj){
  se = obj$se
  plt_crop = obj$plt_crop
  base_pt_size = obj$base_pt_size/1.2
  se = add_NICHES_to_se(se, obj$NICHES_result$logcounts)

  plotmax_T = max(quantile(unlist(se[["T.cells_logcounts"]]), probs = 0.995), 0.4)
  plotmax_NK = max(quantile(unlist(se[["NK.cells_logcounts"]]), probs = 0.995), 0.4)
  plotmax_Mel = max(quantile(unlist(se[["melanocytes.malignant.cells_logcounts"]]), probs = 0.995), 0.4)
  
  return(wrap_plots(ncol = 2, byrow = F,
    p2_scoreplot(se, "CARD_decon_mean_T/NK_cells", limits = c(0., NA), 
                 crop_area = plt_crop, pt.size = base_pt_size, pt.stroke = 0.06,
                 scalename = "Cell Type\nProportion", "T/NK deconvolution"),
    p2_scoreplot(se, "T.cells_logcounts", limits = c(0., plotmax_T), 
                 crop_area = plt_crop, pt.size = base_pt_size, pt.stroke = 0.06,
                 scalename = "Module\nScore", title = "T cells score"),
    p2_scoreplot(se, "NK.cells_logcounts", limits = c(0., plotmax_NK), 
                 crop_area = plt_crop, pt.size = base_pt_size, pt.stroke = 0.06,
                 scalename = "Module\nScore", "NK cells score"),
    p2_scoreplot(se, "CARD_decon_mean_Melanocytic_cells", limits = c(0., NA), 
                 crop_area = plt_crop, pt.size = base_pt_size, pt.stroke = 0.06,
                 scalename = "Cell Type\nProportion", "Melanoma cells deconvolution"),
    p2_scoreplot(se, "melanocytes.malignant.cells_logcounts", limits = c(0., plotmax_Mel), 
                 crop_area = plt_crop, pt.size = base_pt_size, pt.stroke = 0.06,
                 scalename = "Module\nScore", title = "Melanoma cells score")))
}

plt_deconv_and_scores_supp(p3_objs$PM2)
```


## PM3

```{r supp-pm3-a, fig.width=2, fig.height=1.4, fig.show='hold', echo=F, eval=T, dpi=300}
supp_histoplot(other_samples_markers$PM3)
```

```{r supp-pm3-b, fig.width=2, fig.height=1.4, fig.show='hold', echo=F, eval=T, dpi=300}
supp_clusterplot(p3_objs$PM3)
```

```{r supp-pm3-c, fig.width=3.5, fig.height=4.4, fig.show='hold', echo=F, eval=T, dpi=300}
supp_heatmap(other_samples_markers$PM3, genesize = 4.6)
```

```{r supp-pm3-g, fig.width=3., fig.height=2.5, fig.show = 'hold', echo=F, eval=T, dpi=300}
plt_deconv_and_scores_supp(p3_objs$PM3)
```

## PM4

```{r supp-pm4-a, fig.width=2, fig.height=1.4, fig.show='hold', echo=F, eval=T, dpi=300}
supp_histoplot(other_samples_markers$PM4)
```

```{r supp-pm4-b, fig.width=2, fig.height=1.4, fig.show='hold', echo=F, eval=T, dpi=300}
supp_clusterplot(p3_objs$PM4)
```

```{r supp-pm4-c, fig.width=3.5, fig.height=4., fig.show='hold', echo=F, eval=T, dpi=300}
supp_heatmap(other_samples_markers$PM4, genesize = 4.8)
```

```{r supp-pm4-g, fig.width=3., fig.height=2.6, fig.show = 'hold', echo=F, eval=T, dpi=300}
plt_deconv_and_scores_supp(p3_objs$PM4)
```

## Extended 6: unfiltered results table (signif in min 3/4 samples)

```{r ext-6-results-long, fig.width=7., fig.height=9.5, fig.show = 'hold', echo=F, eval=T}
print(paste(melanoma_df_filt %>% filter(count_signif>=3) %>% nrow(), "interactions in at least 3/4 samples"))

plot_combined_df(melanoma_df_filt%>% filter(count_signif>=3), 
                 order.signif.overwrite = T,
                 anno.signifi.legend = "plot", legend_rows = 1,
                 signif_text_size = 1., width_StubS = 11,
                 legend_barwidth = 4.3,
                 plot_StubS = "quant", legends.rel.heights = 0.02, 
                 plot_additional_dfs_list = list(pozniak_df)) & 
  theme(axis.text = element_text(size = unit(4.5, "pt")),
        legend.text = element_text(size = unit(5, "pt")),
        legend.title = element_text(size = unit(6., "pt"), hjust = 0.1),
        title = element_text(size = unit(6.5, "pt")),
        strip.text =  element_text(size = unit(6.5, "pt")), #8.5
        legend.box.margin = margin(1,25,2,1,"pt"))
```

# Supplementary Figures

Less effort :)

## Survival Supplement

```{r survival-supp, fig.width=6.5*2, fig.height=2.5*2, fig.show = 'hold', echo=F, eval=T, dpi=300/2}
plot_grid(survival_extra_p1, NULL, survival_extra_p2, nrow = 1, 
          labels = c("      TCGA-SKCM","", "      Gide et al."), label_size = 14, 
          rel_widths = c(1, 0.1, 1), label_fontface = "plain", hjust = 0, vjust = 1.6)

gide_plt_subset_cd8_split = gide_plt_dat %>%
  filter(PD1_Inhibitor == "Pembrolizumab") %>%
  mutate(CD8_T_cells = ifelse(T.cells.CD8>median(T.cells.CD8), "high", "low")) %>%
  group_by(CD8_T_cells) %>%
  mutate(LGALS1 = ifelse(lgals1>median(lgals1, na.rm = T), "high", "low"),
         PTPRC = ifelse(ptprc>median(ptprc, na.rm = T), "high", "low")) %>%
  as.data.frame()

plot_grid(NULL, ggsurvplot_facet(survfit(Surv(surv_fu_years, surv_deceased) ~ LGALS1, 
                              data = gide_plt_subset_cd8_split), pval.coord = c(0, 0.03), 
                           pval.method.coord =  c(0, 0.15), pval = T, 
                      pval.method = T, data = gide_plt_subset_cd8_split, ggtheme = mysurvtheme(),
                      palette = RColorBrewer::brewer.pal(8, "Set2")[2:1],
                      facet.by = c("CD8_T_cells", "PTPRC"), nrow = 1, conf.int = T) & 
            labs(x = "Time (years)", fill = "LGALS1", 
                 color = "LGALS1") & theme(legend.position = "left"), ncol = 1, rel_heights = c(0.15, 1.6), 
            label_size = 14, label_fontface = "plain", hjust = 0, vjust = 1.6, 
            labels = c("      Gide et al.: Pembrolizumab patients, LGALS1/PTPRC calculated splits within CD8 T cell subgroups", ""))
```

## Supplement 6, Interaction Plots Visium

```{r supp-6-interaction-plots, fig.width=7., fig.height=3.2, fig.show = 'hold', echo=F, eval=T, dpi=300}
supp6_labels = paste0("PM", 1:4)
supp6_objs = setNames(c(list(PM1 = list(se = se, plt_crop = default_plot_focus, base_pt_size = base_pt_size)), 
                        p3_objs[paste0("PM", 2:4)]), nm = supp6_labels)
supp6_widths = sapply(supp6_objs, function(obj){
  crop = obj$plt_crop
  iminfo = obj$se@tools$Staffli@image_info
  w = (crop[3] * iminfo$width) - (crop[1] * iminfo$width)
  h = (crop[4] * iminfo$height) - (crop[2] * iminfo$height)
  return(w/h)
})

plt_genes_w_interaction_2_cols <- function(objs, lig, rec){
  ligrec = paste0(lig, "—", rec)
  n_objs = length(objs)
  n_rows = 1:ceiling(n_objs/2)
  
  plts = lapply(n_rows, function(i){
    obj1 = objs[[(2*i)-1]]
    se1 = obj1$se
    plt_crop1 = obj1$plt_crop
    base_pt_size1 = obj1$base_pt_size
    width1 = supp6_widths[(2*i)-1]
    name1 = supp6_labels[(2*i)-1]
    
    p1 = plot_grid(nrow = 1, axis = "tb", align = "hv", 
            p3_scoreplot(se1, lig, "RNA_logcounts", lig, "Log(Counts+1)",
                         plt_crop1, base_pt_size1, leg_right=F) &
              theme(plot.title = element_text(hjust = 0.5)),
            p3_scoreplot(se1, rec, "RNA_logcounts", rec, "Log(Counts+1)",
                         plt_crop1, base_pt_size1, leg_right=F) &
              theme(plot.title = element_text(hjust = 0.5)),
            p3_scoreplot(se1, ligrec, "CellToNeighborhood", ligrec, 
                         "NICHES\ninteraction score", plt_crop1, base_pt_size1, leg_right=F) &
              theme(plot.title = element_text(hjust = 0.5)))
    if ((2*i) > n_objs){
      p2 = NULL
      width2 = 1
      name2 = ""
    } else {
      obj2 = objs[[(2*i)]]
      se2 = obj2$se
      plt_crop2 = obj2$plt_crop
      base_pt_size2 = obj2$base_pt_size
      width2 = supp6_widths[(2*i)]
      name2 = supp6_labels[(2*i)]
      
      p2 = plot_grid(nrow = 1, axis = "tb", align = "hv", 
              p3_scoreplot(se2, lig, "RNA_logcounts", lig, "Log(Counts+1)",
                           plt_crop2, base_pt_size2, leg_right=F) &
                theme(plot.title = element_text(hjust = 0.5)),
              p3_scoreplot(se2, rec, "RNA_logcounts", rec, "Log(Counts+1)",
                           plt_crop2, base_pt_size2, leg_right=F) &
                theme(plot.title = element_text(hjust = 0.5)),
              p3_scoreplot(se2, ligrec, "CellToNeighborhood", ligrec, 
                           "NICHES\ninteraction score", plt_crop2, base_pt_size2, leg_right=F) &
                theme(plot.title = element_text(hjust = 0.5)))
    }
    return(plot_grid(p1, p2, nrow = 1, rel_widths = c(width1, width2), 
                     labels = c(name1, name2), label_size = 7))
  })
  return(plot_grid(plotlist = plts, ncol = 1))
}

supp6_plts_1 = plt_genes_w_interaction_2_cols(supp6_objs, "LGALS1", "PTPRC")
supp6_plts_2 = plt_genes_w_interaction_2_cols(supp6_objs, "MIF", "CD74")
supp6_plts_3 = plt_genes_w_interaction_2_cols(supp6_objs, "ITGB2", "ITGAL")
supp6_plts_4 = plt_genes_w_interaction_2_cols(supp6_objs, "CXCL9", "P2RY14")

plot(supp6_plts_1)
plot(supp6_plts_2)
plot(supp6_plts_3)
plot(supp6_plts_4)
```

## Supplement 8? Pozniak Data

```{r pozniak-plots-prep}

gen_pozniak_plots <- function(samplename){
  current_sample = samplename
  sample_manifest = manifest[[current_sample]]
  default_plot_focus = sample_manifest$crop_area
  base_pt_size = estim_point_size(sample_manifest)*0.335
  work.dir = file.path(manifest$workdata_parent, sample_manifest$workdata)
  obj1 = readRDS(file.path(work.dir, "09_seurat.rds"))
  se = obj1$se
  se = apply_magick_image_functions(se, function(x)image_average(c(image_contrast(x, sharpen = 1), image_normalize(x))))
  
  plotmax_T = max(quantile(unlist(se[["T.cells_logcounts"]]), probs = 0.995), 0.4)
  plotmax_NK = max(quantile(unlist(se[["NK.cells_logcounts"]]), probs = 0.995), 0.4)
  plotmax_Mel = max(quantile(unlist(se[["melanocytes.malignant.cells_logcounts"]]), probs = 0.995), 0.4)
  
  tscore_range = 1.1*range(se$T.cells_logcounts)
  malscore_range = pmax(1.1*range(se$melanocytes.malignant.cells_logcounts), c(-0.55, 0.5)) # not lower than -0.5 plus padding
  deconv_range_max = max(c(se$`CARD_decon_mean_T/NK_cells`, se$CARD_decon_mean_Melanocytic_cells))+0.01
  tplots_dat = tibble(tnk_decon = se$`CARD_decon_mean_T/NK_cells`, 
                t_score = se$T.cells_logcounts, 
                act = se$T.positive)
  nkplots_dat = tibble(tnk_decon = se$`CARD_decon_mean_T/NK_cells`, 
                nk_score = se$NK.cells_logcounts, 
                act = se$NK.positive)
  malplots_dat = tibble(mal_decon = se$CARD_decon_mean_Melanocytic_cells, 
                mal_score = se$melanocytes.malignant.cells_logcounts, 
                act = se$TUMOR.positive)
  scoreplots = align_plots(
    p2_scoreplot(se, "CARD_decon_mean_Melanocytic_cells", limits = c(0., NA), 
                 crop_area = default_plot_focus, pt.size = base_pt_size,
                 scalename = "Cell Type\nProportion", "Melanoma cells deconvolution"),
    p2_scoreplot(se, "CARD_decon_mean_T/NK_cells", limits = c(0., NA), 
                 crop_area = default_plot_focus, pt.size = base_pt_size,
                 scalename = "Cell Type\nProportion", title = "T/NK deconvolution"),
    p2_scoreplot(se, "melanocytes.malignant.cells_logcounts", limits = c(0., plotmax_Mel), 
                 crop_area = default_plot_focus, pt.size = base_pt_size,
                 scalename = "Module\nScore", title = "Melanoma cells score"),
    p2_scoreplot(se, "T.cells_logcounts", limits = c(0., plotmax_T), 
                 crop_area = default_plot_focus, pt.size = base_pt_size,
                 scalename = "Module\nScore", title = "T cells score"),
    p2_scoreplot(se, "NK.cells_logcounts", limits = c(0., plotmax_T), 
                 crop_area = default_plot_focus, pt.size = base_pt_size,
                 scalename = "Module\nScore", title = "NK cells score")
    )

  scatterplots = align_patches(
    ggplot(malplots_dat, aes(mal_score, mal_decon, color = act)) + 
        geom_hline(yintercept = 0.1, colour = "black", linewidth = 0.5) +
        geom_vline(xintercept = Misc(se, "mal.model")$mixture.model.cutoff, colour = "black", linewidth = 0.5) +
        geom_point(size = 0.2) + 
        scale_color_manual(values = c("False" = "grey", "True" = "#AF4030"), guide = "none") +
        mytheme_grid(base_size = 5) + 
        theme(axis.title = element_text(size = unit(7, "pt"), margin = margin(2,2,2,2,"pt")),
              axis.text = element_text(size = unit(6, "pt"))) +
        coord_cartesian(xlim = malscore_range, 
                        ylim = c(0., deconv_range_max), 
                        expand = F) +
        ylab("Deconv. melanoma content") +
        xlab("Melanoma module score"),
    ggplot(tplots_dat, aes(t_score, tnk_decon, color = act)) + 
        geom_hline(yintercept = 0.1, colour = "black", linewidth = 0.5) +
        geom_vline(xintercept = Misc(se, "t.model")$fdr.corrected.background.cutoff, 
                   colour = "black", linewidth = 0.5) +
        geom_point(size = 0.2) + 
        scale_color_manual(values = c("False" = "grey", "True" = "#40AF40"), guide = "none") +
        mytheme_grid(base_size = 5) + 
        theme(axis.title = element_text(size = unit(7, "pt"), margin = margin(2,2,2,2,"pt")),
              axis.text = element_text(size = unit(6, "pt"))) +
        coord_cartesian(xlim = tscore_range, ylim = c(0., deconv_range_max), expand = F) +
        ylab("Deconv. T/NK content") +
        xlab("T cell module score"),
    ggplot(tplots_dat, aes(t_score, tnk_decon, color = act)) + 
        geom_hline(yintercept = 0.1, colour = "black", linewidth = 0.5) +
        geom_vline(xintercept = Misc(se, "nk.model")$fdr.corrected.background.cutoff, 
                   colour = "black", linewidth = 0.5) +
        geom_point(size = 0.2) + 
        scale_color_manual(values = c("False" = "grey", "True" = "#40AF40"), guide = "none") +
        mytheme_grid(base_size = 5) + 
        theme(axis.title = element_text(size = unit(7, "pt"), margin = margin(2,2,2,2,"pt")),
              axis.text = element_text(size = unit(6, "pt"))) +
        coord_cartesian(xlim = tscore_range, ylim = c(0., deconv_range_max), expand = F) +
        ylab("Deconv. T/NK content") +
        xlab("NK cell module score")
  )
  
  borderregionplot = feature_overlay_discrete(se, pt.size = base_pt_size, pt.stroke = 0.1,
                                 crop_area = default_plot_focus,
                                 features = "active.region.annotated",
                                 titles = "",
                                 colors = c("Double-negative spots" = "black",
                                            "Interfacing tumor-positive spots" = "red",
                                            "Interfacing immune-positive spots" = "green",
                                            "Non-interfacing tumor-positive spots" = "darkred",
                                            "Non-interfacing immune-positive spots" = "darkgreen",
                                            "Interfacing double-positive spots" = "yellow"),
                                 pt_alpha = ifelse(se[["active.region"]]=="Background", 0., 1.),
                                 add_scalebar = T, scalebar_position = c(0.05, 0.85),
                                 scalebar_height = 0.18, scalebar_gg = p2scalebar) &
    theme(legend.title = element_blank(), 
          legend.position = "right",
          legend.key.height = unit(10, "pt"), 
          legend.box.spacing = unit(4, "pt"),
          legend.text = element_text(size = unit(5.5, "pt")), 
          plot.margin = margin(1,1,1,1, "pt"),
          plot.caption = element_blank()) &
    guides(fill=guide_legend(ncol = 1, override.aes = list(size = 2.5)))
  borderregionplot[[1]][[1]] = borderregionplot[[1]][[1]] & labs(caption="Resulting spot choice for LR-analysis")
  borderregionplot[[1]][[2]] = borderregionplot[[1]][[2]] + theme(panel.border = element_blank())
  borderregionplot[[1]][[3]] = borderregionplot[[1]][[3]] + theme(panel.border = element_blank())
  
  return(list(
    scoreplots = scoreplots,
    scatterplots = scatterplots,
    borderregionplot = borderregionplot
  ))
}
```

### POZNIAK 1

```{r poz1-scores, fig.width=1.41, fig.height=1.32, fig.show='hold', echo=F, eval=T, dpi=300}
poz1plots = gen_pozniak_plots("POZNIAK1")
for (p in poz1plots$scoreplots) {plot(p)}
```

```{r poz1-scatters, fig.width=1.86, fig.height=1.64, fig.show='hold', echo=F, eval=T, dpi=300}
for (p in poz1plots$scatterplots) {plot(p)}
```

```{r poz1-border, fig.width=2.8, fig.height=1.55, fig.show='hold', echo=F, eval=T, dpi=300}
plot(poz1plots$borderregionplot)
```

### POZNIAK 2

```{r poz2-scores, fig.width=1.41, fig.height=1.32, fig.show='hold', echo=F, eval=T, dpi=300}
poz2plots = gen_pozniak_plots("POZNIAK2")
for (p in poz2plots$scoreplots) {plot(p)}
```

```{r poz2-scatters, fig.width=1.86, fig.height=1.64, fig.show='hold', echo=F, eval=T, dpi=300}
for (p in poz2plots$scatterplots) {plot(p)}
```

```{r poz2-border, fig.width=2.8, fig.height=1.55, fig.show='hold', echo=F, eval=T, dpi=300}
plot(poz2plots$borderregionplot)
```

### POZNIAK 3

```{r poz3-scores, fig.width=1.41, fig.height=1.32, fig.show='hold', echo=F, eval=T, dpi=300}
poz3plots = gen_pozniak_plots("POZNIAK3")
for (p in poz3plots$scoreplots) {plot(p)}
```

```{r poz3-scatters, fig.width=1.86, fig.height=1.64, fig.show='hold', echo=F, eval=T, dpi=300}
for (p in poz3plots$scatterplots) {plot(p)}
```

```{r poz3-border, fig.width=2.8, fig.height=1.55, fig.show='hold', echo=F, eval=T, dpi=300}
plot(poz3plots$borderregionplot)
```

### POZNIAK 4

```{r poz4-scores, fig.width=1.41, fig.height=1.32, fig.show='hold', echo=F, eval=T, dpi=300}
poz4plots = gen_pozniak_plots("POZNIAK4")
for (p in poz4plots$scoreplots) {plot(p)}
```

```{r poz4-scatters, fig.width=1.86, fig.height=1.64, fig.show='hold', echo=F, eval=T, dpi=300}
for (p in poz4plots$scatterplots) {plot(p)}
```

```{r poz4-border, fig.width=2.8, fig.height=1.55, fig.show='hold', echo=F, eval=T, dpi=300}
plot(poz4plots$borderregionplot)
```

### POZNIAK 5

```{r poz5-scores, fig.width=1.41, fig.height=1.32, fig.show='hold', echo=F, eval=T, dpi=300}
poz5plots = gen_pozniak_plots("POZNIAK5")
for (p in poz5plots$scoreplots) {plot(p)}
```

```{r poz5-scatters, fig.width=1.86, fig.height=1.64, fig.show='hold', echo=F, eval=T, dpi=300}
for (p in poz5plots$scatterplots) {plot(p)}
```

```{r poz5-border, fig.width=2.8, fig.height=1.55, fig.show='hold', echo=F, eval=T, dpi=300}
plot(poz5plots$borderregionplot)
```

### POZNIAK 6

```{r poz6-scores, fig.width=1.41, fig.height=1.32, fig.show='hold', echo=F, eval=T, dpi=300}
poz6plots = gen_pozniak_plots("POZNIAK6")
for (p in poz6plots$scoreplots) {plot(p)}
```

```{r poz6-scatters, fig.width=1.86, fig.height=1.64, fig.show='hold', echo=F, eval=T, dpi=300}
for (p in poz6plots$scatterplots) {plot(p)}
```

```{r poz6-border, fig.width=2.8, fig.height=1.55, fig.show='hold', echo=F, eval=T, dpi=300}
plot(poz6plots$borderregionplot)
```

```{r dummy, eval=F}


# right side, PM2
# tscore2 = se2$T.cells_logcounts
# nkscore2 = se2$NK.cells_logcounts
# tnkdecon2 = se2$`CARD_decon_mean_T/NK_cells`
# malscore2 = se2$melanocytes.malignant.cells_logcounts
# maldecon2 = se2$CARD_decon_mean_Melanocytic_cells
# tpos2 = se2$T.positive
# nkpos2 = se2$NK.positive
# malpos2 = se2$TUMOR.positive
# 
# tmodel2 = Misc(se2, "t.model")
# #tmodel2 = list(fdr.corrected.background.cutoff = 0.05, background.distr.model = list(location = 0., scale = 0.02))
# nkmodel2 = Misc(se2, "nk.model")
# malmodel2 = Misc(se2, "mal.model")
# tscore_range2 = 1.1*range(tscore2)
# malscore_range2 = pmax(1.1*range(malscore2), c(-0.5, 0.5)) # not lower than -0.5 
# deconv_range_max = max(c(tnkdecon2, maldecon2))+0.01
# tplots_dat2 = tibble(tnk_decon = tnkdecon2, 
#               t_score = tscore2, 
#               act = tpos2)
# nkplots_dat2 = tibble(tnk_decon = tnkdecon2, 
#               nk_score = nkscore2, 
#               act = nkpos2)
# malplots_dat2 = tibble(mal_decon = maldecon2, 
#               mal_score = malscore2, 
#               act = malpos2)
# 
# malmodel_bg_dens = compute_tumor_bg_dens(malmodel2, malscore_range2)
# 
# p3A2rtop = ggplot(malplots_dat2, aes(mal_score, mal_decon, color = act)) + 
#   geom_hline(yintercept = 0.2, colour = "black", linewidth = 1) +
#   geom_vline(xintercept = malmodel2$mixture.model.cutoff, colour = "black", linewidth = 1) +
#   geom_point() + 
#   scale_color_manual(values = c("False" = "grey", "True" = "#AF4030"), guide = "none") +
#   mytheme_grid(base_size = 5) + 
#   coord_cartesian(xlim = malscore_range2, 
#                   ylim = c(0., deconv_range_max), 
#                   expand = F) +
#   ylab("Deconvolution melanoma cell content") +
#   scale_y_continuous(position = "right") +
#   xlab("Melanoma module score")
#   
# p3A2rbot = ggplot(malplots_dat2, aes(mal_score)) + 
#   geom_histogram(mapping = aes(y = after_stat(density)),bins = 70, fill = "white", colour = "grey30") + 
#   geom_vline(xintercept = malmodel2$mixture.model.cutoff, colour = "black", linewidth = 1) +
#   geom_line(data = malmodel_bg_dens, mapping = aes(x,bg_dens), color = "#708070", linewidth = 1.2) + 
#   geom_line(data = malmodel_bg_dens, mapping = aes(x,tumor_dens), color = "#AF4030", linewidth = 1.2) + 
#   mytheme_grid(base_size = 5) + 
#   coord_cartesian(xlim = malscore_range2, 
#                   expand = F) +
#   xlab("Melanoma module score") + 
#   ylab("Density")
# 
# p3A2mtop = ggplot(tplots_dat2, aes(t_score, tnk_decon, color = act)) + 
#   geom_hline(yintercept = 0.1, colour = "black", linewidth = 1) +
#   geom_vline(xintercept = tmodel2$fdr.corrected.background.cutoff, colour = "black", linewidth = 1) +
#   geom_point() + 
#   scale_color_manual(values = c("False" = "grey", "True" = "#40AF40"), guide = "none") +
#   mytheme_grid(base_size = 5) + 
#   coord_cartesian(xlim = tscore_range2, 
#                   ylim = c(0., deconv_range_max), 
#                   expand = F) +
#   ylab("Deconvolution T/NK cell content") +
#   scale_y_continuous(position = "right") +
#   xlab("T cell module score")
# 
# p3A2mbot = ggplot(tplots_dat2, aes(t_score)) + 
#   geom_histogram(mapping = aes(y = after_stat(density)),bins = 70, fill = "white", colour = "grey30") + 
#   geom_vline(xintercept = tmodel2$fdr.corrected.background.cutoff, colour = "black", linewidth = 1) +
#   geom_line(data = tibble(x = seq(tmodel2$background.distr.model$location-8*tmodel2$background.distr.model$scale,
#                                   tmodel2$background.distr.model$location+8*tmodel2$background.distr.model$scale, 
#                                   len = 150),
#                           dens = dnorm(
#                             x=seq(tmodel2$background.distr.model$location-8*tmodel2$background.distr.model$scale,
#                                   tmodel2$background.distr.model$location+8*tmodel2$background.distr.model$scale, 
#                                   len = 150),
#                             mean = tmodel2$background.distr.model$location,
#                             sd = tmodel2$background.distr.model$scale)),
#             aes(x,dens), color = "#40AF40", linewidth = 1.2) + # draw background distribution
#   mytheme_grid(base_size = 5) + 
#   coord_cartesian(xlim = tscore_range2, 
#                   expand = F) +
#   xlab("T cell module score") + 
#   ylab(" Density ")
# 
# p3A2ltop = ggplot(nkplots_dat2, aes(nk_score, tnk_decon, color = act)) + 
#   geom_hline(yintercept = 0.1, colour = "black", linewidth = 1) +
#   geom_vline(xintercept = nkmodel2$fdr.corrected.background.cutoff, colour = "black", linewidth = 1) +
#   geom_point() + 
#   scale_color_manual(values = c("False" = "grey", "True" = "#40AF40"), guide = "none") +
#   mytheme_grid(base_size = 5) + 
#   coord_cartesian(xlim = 1.1*range(nkscore2), 
#                   ylim = c(0., deconv_range_max), 
#                   expand = F) +
#   ylab(" Deconvolution T/NK cell content ") +
#   xlab("NK cell module score")
# 
# p3A2lbot = feature_overlay_discrete(se2, pt.size = base_pt_size2*0.5, crop_area = default_plot_focus2,
#                               features = "active.region.annotated", 
#                               titles = "", 
#                               colors = c("Double-negative spots" = "black",
#                                   "Interfacing tumor-positive spots" = "red",
#                                   "Interfacing immune-positive spots" = "green",
#                                   "Non-interfacing tumor-positive spots" = "darkred",
#                                   "Non-interfacing immune-positive spots" = "darkgreen",
#                                   "Interfacing double-positive spots" = "yellow"), 
#                               pt_alpha = ifelse(se2[["active.region"]]=="Background", 0., 1.),
#                               add_scalebar = T, scalebar_position = c(0.05, 0.82), 
#                               scalebar_height = 0.15, scalebar_gg = p2scalebar) &
#   theme(legend.title = element_blank(), legend.position = "none",
#         plot.title = element_blank(), plot.caption = element_text(hjust = 0.5))
# p3A2lbot[[1]][[1]] = p3A2lbot[[1]][[1]] & labs(caption="Resulting spot choice for LR-analysis")
# p3A2lbot[[1]][[2]] = p3A2lbot[[1]][[2]] + theme(panel.border = element_blank())
# p3A2lbot[[1]][[3]] = p3A2lbot[[1]][[3]] + theme(panel.border = element_blank())
# 
# p3A2 = wrap_elements(wrap_plots(A=p3A2ltop, B=p3A2mtop, C=p3A2rtop,
#            D=p3A2mbot, E=p3A2rbot, 
#            axes = "collect", design = "ABC
#                                        #DE")) +
#   inset_element(p3A2lbot, align_to = "full",
#                 left = 0.01, top = 0.48, right = 0.35, bottom = 0.1)
# 
# p3A = plot_grid(p3A1, NULL, 
#           plot_grid(NULL, p3Alegend, NULL, ncol = 1, rel_heights = c(1,1, 0.1)), p3A2, 
#           nrow = 1, rel_widths = c(9,0.13,2,9))


# 
# ora_dotplot = function(.tops = 10, .by = c("richFactor", "p.adjust"), cluster_order = NULL, base_size = 8, sizerange = c(2,6), show.tops.legend = F) {
#   .by <- match.arg(.by)
#   
#   res = res.cP.cc.go.sim@compareClusterResult
#   res = mutate(res, totalGenes = as.numeric(sub("\\d+/", "", GeneRatio)),
#                richFactor = Count / as.numeric(sub("/\\d+", "", BgRatio)))
#   # at least 5 genes or 10% (but never less than 2) of the positive marker genes of a cluster must 
#   # belong to a term for it to be included
#   res = res[(res$Count >= pmin(5, pmax(2, res$totalGenes*0.1))), ]
#   res$geneID = NULL
#   res$ID_CL = paste0(res$ID, "_", res$Cluster)
#   
#   # New line for long Description strings
#   # for (i in 1:nrow(res)) {
#   #     res$Description[i] = paste(base::strwrap(res$Description[i], 75), collapse="\n")
#   # }
#   
#   res.cp = res
#   
#   tops.l = lapply(split(res, res$Cluster), function(x){
#     if (.by == "richFactor") {
#       x = x[order(x$richFactor, decreasing = T), ]  # sort by enrichment factor
#     } else {
#       x = x[order(x$p.adjust, decreasing = F), ]    # sort by adjusted p value
#     }
#     x = head(x, .tops)
#     x
#   })
#   
#   tops.l = tops.l[sapply(tops.l, function(x) dim(x)[1]) > 0]
#   tops = as.data.frame(data.table::rbindlist(tops.l))
#   tops$ID_CL = paste0(tops$ID, "_", tops$Cluster)
#   res.cp = subset(res.cp, ID %in% tops$ID)
#   
#   lvls = tops.l[[1]]$Description
#   lvls = c(lvls, setdiff(res.cp$Description, lvls))
#   res.cp$Description = factor(res.cp$Description, levels = rev(lvls))
#   res.cp$Cluster = factor(res.cp$Cluster, levels = unique(naturalsort::naturalsort(res.cp$Cluster)))
#   
#   res.cp$TOPS = res.cp$ID_CL %in% tops$ID_CL
#   
#   
#   ora = ggplot(res.cp, aes(x = Cluster, y = Description)) +
#     geom_point(aes(fill = p.adjust, size = richFactor), pch=21, stroke=.3) +
#     geom_point(aes(colour = TOPS, size = richFactor), pch=21, stroke=.5, show.legend = show.tops.legend) +
#     scale_size(range = sizerange) +
#     mytheme_grid(base_size = base_size) +
#     theme(
#       panel.grid.minor = element_blank(),
#       panel.grid.major.x = element_blank(),
#       axis.ticks = element_blank(),
#       legend.position = "right"
#     ) +
#     guides(fill = guide_colorbar(
#       title.position = 'top', title.hjust = 0, barwidth = unit(.5, 'lines'), barheight = unit(6, 'lines'),
#       title = "Adjusted\np-value", order = 3)
#     ) + 
#     guides(size = guide_legend(
#       ncol = 1, override.aes =, title.hjust = .5, title.position = "top",
#       title = "Rich factor", order = 1)
#     ) +
#     guides(colour = guide_legend(
#       ncol = 1, override.aes =, title.hjust = .5, title.position = "top",
#       title = "Tops", order = 2)
#     ) +
#     # scale_fill_scico(palette = 'batlow', direction = -1, end = .7)  +
#     scale_fill_scico(palette = 'bilbao', begin = .2, end = .8, direction = -1) +
#     scale_colour_manual(values = c("#555555", "#44AA99")) +
#     xlab("Cluster") + ylab(NULL)  
# 
#   if (!is.null(cluster_order)){
#     ora = ora + lims(x = cluster_order)
#   }
#     
#   ora
# }


```
