---
title: "Plots for HD Data with LIANA Results"
output:
  workflowr::wflow_html:
    code_folding: hide
    fig_caption: yes
    number_sections: yes
    pandoc_args: []
    self_contained: yes
    smart: yes
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: no
      smooth_scroll: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, message=FALSE, warning=FALSE, echo=FALSE, cache=F}
knitr::opts_chunk$set(
  echo = FALSE,
  error = FALSE,
  message = FALSE,
  cache = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.align = 'center',
  fig.show = 'hold',
  autodep = TRUE,
  dpi = 300,
  dev = c('png') # alternative: dev = c('png') type = 'cairo'
)
```

```{r libs-and-functions, echo = F, message = FALSE, warning=FALSE, fig.show='hide'}
library("conflicted")
library("Cairo")
library("tidyverse")
library("yaml")
library("Seurat")
library("semla")

library("ggplot2")
library("ggnewscale")
library("naturalsort")
library("ggalluvial")
library("patchwork")
library("cowplot")
library("ggraph")
library("ggtext")
library("grid")
library("gridpattern")
library("tidygraph")
library("ggraph")

library("liana")

conflicts_prefer(
  dplyr::filter
)

# >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
# Functions
# >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
source("code/helper/Rmarkdown-style.R")
theme_set(mytheme(base_size = 6))

# color celltype labels by colmap in markdown, if not in colmap leave it as is
md_encode_colors <- function(data, colmap, bolden = T){
  ifelse(is.na(colmap[data]),
         paste0(data),
         paste0("<span style='color:",colmap[data],";'>",ifelse(bolden,"**",""),data,ifelse(bolden,"**",""),"</span>"))
}

# adapted from code from the liana package
bysample_dotplot <- function(liana_res,
                             source_groups = NULL,
                             target_groups = NULL,
                             ntop = NULL,
                             sample_column = "sample",
                             wrap_target_xaxis_sample = F,
                             Sample_order = NULL,
                             specificity = "natmi.edge_specificity",
                             magnitude = "sca.LRscore",
                             y.label = "Interactions (Ligand -> Receptor)",
                             size.label = "Interaction\nSpecificity",
                             colour.label = "Expression\nMagnitude",
                             show_complex = TRUE,
                             size_range = c(0.15, 2.),
                             invert_specificity = FALSE,
                             invert_magnitude = FALSE,
                             invert_function = function(x) -log10(x + 1e-10),
                             highlight = c(""),
                             title.margins = margin(2,8,2,8,"pt"),
                             base_size = 4.8,
                             cellcolors = c("#E69F00", "#56B4E9", # colour blind palette from http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
                                            "#009E73", "#F0E442", "#0072B2",
                                            "#D55E00", "#CC79A7", "#DF69A7"),
                             ...
){

  if(show_complex){
    entities <- c("ligand.complex", "receptor.complex")
  } else{
    entities <- c("ligand", "receptor")
  }

  # Modify for the plot
  liana_mod <- liana_res %>%
    # Filter to only the cells of interest
    `if`(!is.null(source_groups),
         filter(., source %in% source_groups),
         .) %>%
    `if`(!is.null(target_groups),
         filter(., target %in% target_groups),
         .)


  if(!is.null(ntop)){
    # Subset to the X top interactions
    top_int <- liana_mod %>% distinct_at(entities) %>% head(ntop)
    liana_mod = liana_mod %>% inner_join(top_int, by=entities)
  }

  if(invert_magnitude){
    liana_mod %<>% mutate(!!magnitude := invert_function(.data[[magnitude]]))
  }
  if(invert_specificity){
    liana_mod %<>% mutate(!!specificity := invert_function(.data[[specificity]]))
  }

  liana_mod %<>%
    dplyr::rename(magnitude = !!magnitude) %>%
    dplyr::rename(specificity = !!specificity) %>%
    dplyr::rename(Sample = !!sample_column) %>%
    mutate(across(all_of(entities), function(x)gsub(x = x, "_", "+"))) %>%
    unite(entities, col = "interaction", sep = "→") %>%
    unite(c("source", "target"), col = "source_target", remove = FALSE)

  colmap = if(is.null(names(cellcolors))) {
    setNames(cellcolors, nm = unique(c(liana_mod$source, liana_mod$target))[1:length(cellcolors)])
  } else {cellcolors}

  # ensure levels & order is kept the plot
  Sample_order <- if (is.null(Sample_order)) {liana_mod %>% pull("Sample") %>% unique()} else {Sample_order}
  liana_mod %<>%
    mutate(Sample = factor(Sample, levels=Sample_order)) %>%
    mutate(source = md_encode_colors(as.character(source), colmap), # encode colors
           target = md_encode_colors(as.character(target), colmap),
           interaction = md_encode_colors(as.character(interaction), highlight)) %>%
    mutate(interaction = factor(interaction, levels=rev(unique(interaction)))) %>%
    mutate(across(where(is.character), as.factor))

  if (length(unique(liana_mod$Sample))==1) {
    if (wrap_target_xaxis_sample) {
      if (length(unique(liana_mod$source))==1) {
        wrap_vars = vars(source)
        wrap_title = paste0("Target = ", unique(liana_mod$target), ", Source:")
      } else {
        wrap_vars = vars(source)
        wrap_title = paste0("Target = ", unique(liana_mod$target), ", Source:")
      }
    } else {
      if (length(unique(liana_mod$source))==1) {
        wrap_vars = vars(source)
        wrap_title = paste0("Sample = ", unique(liana_mod$Sample), ", Source:")
      } else {
        wrap_vars = vars(source)
        wrap_title = paste0("Sample = ", unique(liana_mod$Sample), ", Source:")
      }
    }
  } else {
    if (wrap_target_xaxis_sample) {
      if (length(unique(liana_mod$source))==1) {
        wrap_vars = vars(target)
        wrap_title = paste0("Source = ", unique(liana_mod$source), ", Target:")
      } else {
        wrap_vars = vars(source, target)
        wrap_title = "Source / Target"
      }
    } else {
      if (length(unique(liana_mod$source))==1) {
        wrap_vars = vars(Sample)
        wrap_title = paste0("Source = ", unique(liana_mod$source), ", Sample:")
      } else {
        wrap_vars = vars(source, Sample)
        wrap_title = "Source / Sample"
      }
    }
  }

  xaxisvar = if (wrap_target_xaxis_sample) {quo(Sample)} else {quo(target)}
  
  # smaller text size
  text_size = 0.8*base_size 
  # plot
  suppressWarnings(
    ggdraw(
      ggplot(liana_mod,
           aes(x = !!xaxisvar,
               y = interaction,
               colour = magnitude,
               size = specificity,
               group = target
           )) +
      geom_point() +
      scale_color_gradientn(colours = viridis::viridis(20)) +
      scale_size_continuous(range = size_range) +
      facet_grid(cols = wrap_vars,
                 space = "free",
                 scales ="free",
                 switch = "y")  +
      # scale_x_discrete(position = "right") +
      labs(y = y.label,
           colour = colour.label,
           size = size.label,
           x = ifelse(wrap_target_xaxis_sample, "Sample", "Target"),
           title= wrap_title
      ) +
      theme_bw(base_size = base_size) +
      theme(
        legend.text = element_text(size = text_size),
        axis.text.x = element_markdown(size = text_size,
                                   hjust = 1., face = "bold",
                                   angle = 40),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = text_size,
                                   margin = margin(r = 4, unit = "pt")),
        axis.text.y.left = element_markdown(),
        legend.title = element_text(size = text_size, colour = "gray6"),
        legend.box.spacing = unit(2, "pt"),
        panel.spacing = unit(0.1, "lines"),
        plot.title = element_markdown(vjust = 0, hjust=0.5, margin = title.margins),
        strip.background = element_rect(fill = NA, color = NA), 
        strip.clip = "off", # in case the text is bigger than the boxes, but it's not nice
        strip.text.x.top = element_markdown(size = base_size, margin = title.margins),
        ...
      )
    ) + draw_label("Sample", 0.05, 0.02, 0., 0., fontface = "bold", color = "gray6", size = text_size*1.2)
  )
}

# inspired by the HumanColonCancer_VisiumHD repo celltype connection plot,
# which in turn is mostly taken from CellChat [netVisual_circle]
# but written from scratch with ggraph
celltypes_connection_plot <- function( graph_df,
                                       celltype_colors_table,
                                       source_column = source,
                                       target_column = target,
                                       weight_column = n,
                                       layout = "dh",
                                       vertex.size = 40,
                                       text.size = 6,
                                       legend.font.size = 20,
                                       lighten.weak.edges = T,
                                       edge_size_range = c(1.5,5.5),
                                       arrows = arrow(type = "closed", angle = 25),
                                       show_size_legend = T,
                                       expand_mult = c(0.01, 0.01),
                                       expand_abs = c(0.2, 0.25),
                                       vertex_connections_cap = 16,
                                       size_legend_name = "Number of\nsamples with\nsignificant\nsignal"
){
  edges = graph_df %>% transmute(from = {{source_column}}, to = {{target_column}}, w={{weight_column}},
                                                   source_col = cellcolors[{{source_column}}], target_col = cellcolors[{{target_column}}], 
                                                   source_col_mod = colorspace::lighten(source_col, (max(w)-w)/(max(w)-min(w)+1), 
                                                                                        space = "combined"))
  gr_df = tbl_graph(nodes = tibble(name = unique(c(pull(graph_df, {{source_column}}), pull(graph_df, {{target_column}}))),
                                   cellcol = celltype_colors_table[name],
                                   textcol = ifelse(farver::decode_colour(cellcol, "rgb", "hcl")[,"l"]>55, "black", "white")),
                    edges = edges, directed = T)

  use.aes = if (lighten.weak.edges) aes(colour = source_col_mod, width = w) else aes(colour = source_col, width = w)
  w_range = min(edges$w):max(edges$w)
  modified_blacks = colorspace::lighten(rep("black", length(w_range)), space = "combined",
                                        (max(w_range)-w_range)/(max(w_range)-min(w_range)+1))
  p = ggraph(gr_df, layout = layout) +
    geom_edge_link(use.aes, lineend = "butt", linejoin = "mitre", linemitre = 4, show.legend = show_size_legend,
                   arrow = arrows, end_cap = circle(vertex_connections_cap*1.5, "mm"), start_cap = circle(vertex_connections_cap, "mm"), key_glyph = "blank") +
    geom_edge_loop(aes(colour = source_col, width = w, end_cap = circle(vertex_connections_cap, "mm"), start_cap = circle(vertex_connections_cap, "mm")), 
                   show.legend = show_size_legend) +
    scale_edge_width(size_legend_name, range = edge_size_range, breaks = function(x)min(x):max(x), 
                     guide = guide_legend(override.aes = list(edge_alpha = 1., edge_colour = modified_blacks)))+#ifelse(lighten.weak.edges, list(edge_color = "red", edge_alpha = 0.5), list()))) +
    scale_edge_color_identity() +
    geom_node_point(aes(color = cellcol), size = vertex.size) +
    scale_color_identity(aesthetics = c("color", "fill")) +
    geom_node_label(aes(label = name, fill = cellcol, color = textcol),  label.size = NA, size = text.size) +
    scale_x_continuous(expand = expansion(add = expand_abs[1], mult = expand_mult[1])) +
    scale_y_continuous(expand = expansion(add = expand_abs[2], mult = expand_mult[2])) +
    theme_void(base_size = legend.font.size)

  return(p)
}

rescale_saturating = function(range = c(0., 1.), from = c(0., 1.), rel_saturating = 0.5){
  force(range)
  function(x) {
    saturation_point = rel_saturating*diff(from) + from[1]
    pmin((x - from[1]) / (saturation_point - from[1]) * diff(range) + range[1], range[2])
  }
}

plot_scalebars = function(xstart, y, length = 200, label = paste0(length, "µm"),
                          ticks_major = c(0., 1.), ticks_minor = c(0.25, .5, 0.75),
                          tickheight_major = 10, tickheight_minor = 5,
                          background_color = "white", background_alpha = 0.8, 
                          background_height_adjust = 20, background_expansion = 5
                          ){
  return(list(
    annotate("rect", fill=alpha(background_color, background_alpha), ymin = y-tickheight_major-background_expansion,
             ymax = y+tickheight_major+background_expansion+background_height_adjust,
             xmin = xstart-background_expansion, xmax = xstart+length+background_expansion),
    annotate("segment", x = xstart, xend = xstart+length, y = y),
    annotate("segment", x = xstart+(length*ticks_major), y = y-tickheight_major, yend = y+tickheight_major),
    annotate("segment", x = xstart+(length*ticks_minor), y = y-tickheight_minor, yend = y+tickheight_minor),
    annotate("text", x = xstart+(length/2), y = y+tickheight_major, hjust = 0.5, vjust = 0., label = label)
  ))
}

cellcolors = c("Plasma_cells" = "#114477", #
  "B_cells" = "#4477AA", #
  "pDC" = "#77AADD", #
  "cDC" =  "#228895", #
  "cDC_1" =  "#228895",
  "cDC_2" = "#55BBBB",
  "Mac" = "#33D9B1", #
  "NK" = "#B9E532", #
  "T" = "#EE6310", #
  "Tcyt_activated" = "#FF5013",
  "Teff_mem" = "#EE6310",
  "Treg" = "#F1831D",
  "Tdp" = "#FEA514",
  "VE" = "#AA4455", #
  "VE_1" = "#661122",
  "VE_2" = "#AA4455",
  "LE" = "#EE7788", #
  "Mast_cells" = "#FEEB88", #
  "Pericyte" = "#BABABA",  #
  "Pericyte_1" = "#BABABA",
  "Pericyte_2" = "#808080",
  "Fb" = "#AAAA44", #
  "Fb_1" = "#777711",
  "Fb_2" = "#AAAA44",
  "Fb_3" = "#DDDD77",
  "Kc" = "#9A2E71", #
  "Kc_postmit" = "#9A2E71",
  "Kc_premit" = "#FDA7DF",
  "Eccrine_sweat_gland" = "#CC99BB", #
  "Sebaceous_gland" = "#BB1199",
  "Mel" = "#773311", #
  "Mel_nc-like" = "#331100", #
  "Mel_trans" = "#AA7744", #
  "Mel_trans_melan" = "#773311",
  "Other" = "#BABABA") 

celltype_spiderplots_collapsing = c("Plasma_cells" = "Plasma_cells", 
  "B_cells" = "B_cells", 
  "pDC" = "pDC", 
  "cDC" =  "cDC", 
  "cDC_1" =  "cDC",
  "cDC_2" = "cDC",
  "Mac" = "Mac", 
  "NK" = "NK", 
  "T" = "T", 
  "Tcyt_activated" = "Tcyt_activated",
  "Teff_mem" = "Teff_mem",
  "Treg" = "Treg",
  "Tdp" = "Tdp",
  "VE" = "Other", 
  "VE_1" = "Other",
  "VE_2" = "Other",
  "LE" = "Other", 
  "Mast_cells" = "Other", 
  "Mel" = "Mel", 
  "Mel_nc-like" = "Mel_nc-like", 
  "Mel_trans" = "Mel_trans", 
  "Mel_trans_melan" = "Mel_trans_melan",
  "Pericyte" = "Other",  
  "Pericyte_1" = "Other",
  "Pericyte_2" = "Other",
  "Fb" = "Fb", 
  "Fb_1" = "Fb",
  "Fb_2" = "Fb",
  "Fb_3" = "Fb",
  "Kc" = "Kc", 
  "Kc_postmit" = "Kc",
  "Kc_premit" = "Kc",
  "Eccrine_sweat_gland" = "Other", 
  "Sebaceous_gland" = "Other") 

grid.newpage()
pattern = grid::pattern(grid.pattern_stripe(density = 0.3, angle = 45, fill = "gray70", 
                                            linewidth = 0., spacing = 0.25), 
                        extend = "repeat", group = F)

```

```{r manifest-and-objects, echo=F}
manifest = yaml.load_file("./assets/manifest.yaml")
hd_samples = sapply(setNames(nm=manifest$HD_data_ids_used), function(s)paste0(
  manifest$workdata_parent, manifest$HD_intermediates, "HD_preprocessed_", s, ".rds"))

sample_data_existing = hd_samples[sapply(hd_samples, file.exists)]
hd_objs_processed = lapply(sample_data_existing, readRDS)
```

Author: Florian Große (Last updated: `r Sys.Date()`)

```{r samples, results='asis', echo = F}
cat("HD Samples with processed data: ", paste(names(sample_data_existing), collapse = ", "), "\n")
```

# HD Plots

## Basic Plots

```{r hd-res-1, fig.width=8., fig.asp=0.75, fig.show = 'hold', echo=T, eval=T}
obj_plt_data = lapply(hd_objs_processed, function(x)x$obs_for_alluvials)

for (s in names(obj_plt_data)) {
  curr_obj = obj_plt_data[[s]]
  curr_obj$pass_QC = factor(as.character(curr_obj$pass_QC), 
                            levels = c("TRUE", "FALSE"))
  p = ggplot(curr_obj, aes(x_tr, y_tr,
                         color = tacco_OT_celltype_broad  )) +
  geom_point(size = 0.125) +
  coord_equal(expand = F) + 
  # scale_y_reverse() + # not consistent
  ggtitle(paste0("Sample: ", s), "QC passed") +
  mytheme() +
  theme(legend.position = "bottom", 
        panel.grid.major = element_line("grey60", linewidth = 0.15),
        panel.grid.minor = element_line("grey80", linewidth = 0.1)) +
  scale_color_manual(values = cellcolors) +
  guides(color = guide_legend(title = "Celltype (broad)", 
                              override.aes = list(size = 4))) +
  facet_wrap(~pass_QC+labels_joint_source) # labels_joint_source, pass_QC
  plot(p)
}
```

## Liana

```{r hd-liana-1, fig.width=8., fig.asp=ceiling(length(hd_objs_processed)/4), fig.show = 'hold', echo=T, eval=T}

agg_rank_cutoff = 0.01

liana_results = lapply(hd_objs_processed, function(x)x$liana_res)
liana_aggs = lapply(liana_results, liana_aggregate)

# patch custom celltype colors into the liana heatmap
colored_liana_heatmap <- function(mat,
                          font_size = 6,
                          grid_text = FALSE,
                          name = 'Frequency',
                          pallette = c("white", "violetred2"),
                          row_title = "Sender (Cell types)",
                          column_title = "Receiver (Cell types)",
                          cell_anno = NULL,
                          ...){

    if(grid_text){
        grid_text <- function(j, i, x, y, width, height, fill) {
            grid_text <- grid.text(sprintf("%d", mat[i, j]),
                                   x, y, gp = gpar(fontsize = font_size*0.83))
        }
    } else{
        grid_text <- NULL
    }

    # define Annotations and Barplots
    if (is.null(cell_anno)) {
      cell_anno <- unique(rownames(mat))
      cell_anno <- grDevices::colorRampPalette(
        (RColorBrewer::brewer.pal(n = 8, name = 'Dark2'))
        )(length(cell_anno)) %>%
        setNames(cell_anno)
    } else {
      cell_anno = cell_anno[unique(rownames(mat))]
    }
    
    ## Annotations
    ha_opts <- list(show_legend = FALSE,
                    show_annotation_name = FALSE,
                    col = list("anno"=cell_anno),
                    simple_anno_size = grid::unit(0.25, "cm"))
    column_ha <- exec("HeatmapAnnotation", anno = names(cell_anno), !!!ha_opts)
    row_ha <- exec("rowAnnotation", anno = names(cell_anno), !!!ha_opts)

    # Barplots
    column_bar <- ComplexHeatmap::HeatmapAnnotation(
        bar = .anno_barplot(colSums(mat),
                            cell_anno,
                            axis.font.size = font_size*0.5
                            ),
        annotation_name_gp = gpar(fontsize = font_size*0.5),
        show_legend = FALSE,
        show_annotation_name = FALSE)

    row_bar <- ComplexHeatmap::rowAnnotation(
        bar2 = .anno_barplot(rowSums(mat),
                             cell_anno,
                             font_size*0.5
                             ),
        gp = gpar(fill = cell_anno,
                  col = cell_anno),
        show_legend = FALSE,
        show_annotation_name = FALSE)

    # Heatmap
    ComplexHeatmap::Heatmap(mat,
                            col=colorRampPalette(pallette)(10),
                            cluster_rows = FALSE,
                            cluster_columns = FALSE,
                            row_names_side = "left",
                            top_annotation = column_bar,
                            bottom_annotation = column_ha,
                            right_annotation = row_bar,
                            left_annotation = row_ha,
                            row_title = row_title,
                            row_names_gp = gpar(fontsize = font_size),
                            row_title_gp = gpar(fontsize = font_size*1.2),
                            column_names_gp = gpar(fontsize = font_size),
                            column_title = column_title,
                            column_title_gp = gpar(fontsize = font_size*1.2),
                            column_title_side = "bottom",
                            heatmap_legend_param = list(title_gp = gpar(fontsize = font_size*0.9,
                                                                        fontface = 'bold'),
                                                        border = NA,
                                                        labels_gp = gpar(fontsize = font_size*0.9),
                                                        grid_width = unit(2, "mm")),
                            name = name,
                            cell_fun = grid_text,
                            ...
    )

}
environment(colored_liana_heatmap) <- asNamespace("liana")
assignInNamespace("liana_heatmap", colored_liana_heatmap, asNamespace("liana"))

freqs = lapply(liana_aggs, function(x)grid::grid.grabExpr(ComplexHeatmap::draw(heat_freq(filter(x, aggregate_rank <= agg_rank_cutoff), cell_anno = cellcolors, font_size = 6))))

plot_grid(plotlist = freqs, ncol = 2, labels = names(liana_aggs), 
          hjust = 0., vjust = 1.1, label_size = 5)
```

LGALS1 -> PTPRC celltype connections across samples, min 2, 3 or 4 samples support

```{r hd-liana-tables-prep, fig.width=3.54, fig.asp=0.7, echo=T, eval=T}
# filter tables and arrange nicely
filter_by_supporting_samples_and_reorder = function(df, nmin = 2){
    # 2 stages because we don't want to discard any dots for interactions that
    # have supporting data for any source-target combination and would appear in the plot
    supported_interactions = df %>%
      group_by(source, target, ligand.complex, receptor.complex) %>%
      mutate(n = n(), LR = paste0(ligand.complex, "→", receptor.complex)) %>%
      ungroup() %>%
      filter(n >= nmin) %>%
      pull(LR) %>%
      unique()
    ret = df %>%
      mutate(LR = paste0(ligand.complex, "→", receptor.complex)) %>%
      filter(LR %in% supported_interactions)
    ret = ret %>%
      group_by(source, target, ligand.complex, receptor.complex) %>%
      mutate(mean_aggregate_rank = exp(mean(log(aggregate_rank)))) %>% # geometric mean of aggregate ranks
      ungroup() %>%
      arrange(mean_aggregate_rank)
    return(ret)
}

visium_results = readr::read_csv("data/LRI_results_melanomas_with_annos.csv.gz")
visium_results = visium_results %>%
  select(mechanism, count_signif) %>%
  filter(count_signif >= 2) %>%
  mutate(interaction_liana_format = gsub(pattern = "—", replacement = "→", mechanism),
         color = case_when(
           interaction_liana_format == "LGALS1→PTPRC" ~ "#004488",
           count_signif == 4 ~ "#AC4343",
           count_signif == 3 ~ "#E1783E",
           .default = "#EFBF44"
         ))
highlight_colors = setNames(visium_results$color, nm = visium_results$interaction_liana_format)

# ref_collapsing_file = "assets/single_cell_reference/cell_type_state_matching.tsv"
# celltype_collapsing = read_tsv(ref_collapsing_file)
# sc_ref_markers = readRDS("data/sc_markers_Stubenvoll_Schmidt_2025.rds")
# random_seed = 1234567

min2_melnc = do.call(rbind, lapply(seq_along(liana_aggs), function(i){
    agg = liana_aggs[[i]] %>%
      filter(aggregate_rank<=0.01 &
               source == "Mel_nc-like" &
               target %in% c("NK", "Treg", "Tcyt_activated", "Tdp", "Teff_mem"))
    agg$sample=names(liana_aggs)[i]
    agg
  })) %>% filter_by_supporting_samples_and_reorder(nmin = 2)

min2_meltrans = do.call(rbind, lapply(seq_along(liana_aggs), function(i){
    agg = liana_aggs[[i]] %>%
      filter(aggregate_rank<=0.01 &
               source == "Mel_trans" &
               target %in% c("NK", "Treg", "Tcyt_activated", "Tdp", "Teff_mem"))
    agg$sample=names(liana_aggs)[i]
    agg
  })) %>% filter_by_supporting_samples_and_reorder(nmin = 2)

min2_meltransmelan = do.call(rbind, lapply(seq_along(liana_aggs), function(i){
    agg = liana_aggs[[i]] %>%
      filter(aggregate_rank<=0.01 &
               source == "Mel_trans_melan" &
               target %in% c("NK", "Treg", "Tcyt_activated", "Tdp", "Teff_mem"))
    agg$sample=names(liana_aggs)[i]
    agg
  })) %>% filter_by_supporting_samples_and_reorder(nmin = 2)

min3_melnc = do.call(rbind, lapply(seq_along(liana_aggs), function(i){
    agg = liana_aggs[[i]] %>%
      filter(aggregate_rank<=0.01 &
               source == "Mel_nc-like" &
               target %in% c("NK", "Treg", "Tcyt_activated", "Tdp", "Teff_mem"))
    agg$sample=names(liana_aggs)[i]
    agg
  })) %>% filter_by_supporting_samples_and_reorder(nmin = 3)

min3_meltrans = do.call(rbind, lapply(seq_along(liana_aggs), function(i){
    agg = liana_aggs[[i]] %>%
      filter(aggregate_rank<=0.01 &
               source == "Mel_trans" &
               target %in% c("NK", "Treg", "Tcyt_activated", "Tdp", "Teff_mem"))
    agg$sample=names(liana_aggs)[i]
    agg
  })) %>% filter_by_supporting_samples_and_reorder(nmin = 3)

min3_meltransmelan = do.call(rbind, lapply(seq_along(liana_aggs), function(i){
    agg = liana_aggs[[i]] %>%
      filter(aggregate_rank<=0.01 &
               source == "Mel_trans_melan" &
               target %in% c("NK", "Treg", "Tcyt_activated", "Tdp", "Teff_mem"))
    agg$sample=names(liana_aggs)[i]
    agg
  })) %>% filter_by_supporting_samples_and_reorder(nmin = 3)

# LGALS1-PTPRC connection
lgals1_ptprc_dat = do.call(rbind, lapply(seq_along(liana_aggs), function(i){
    agg = liana_aggs[[i]] %>%
        filter(aggregate_rank<=0.01 & ligand.complex=="LGALS1" & receptor.complex=="PTPRC")
    agg$sample=names(liana_aggs)[i]
    agg
}))
# keep connections with at least 2 samples only
lgals1_ptprc_dat = lgals1_ptprc_dat %>%
  group_by(source, target) %>%
  distinct() %>%
  mutate(n = n()) %>%
  summarize(aggregate_rank = mean(aggregate_rank),
            n = first(n),
            samples = paste(sample, collapse=", "),
            .groups = "drop")

set.seed(1234)
celltypes_connection_plot(lgals1_ptprc_dat %>% filter(n>=2),
                          celltype_colors_table = cellcolors, layout = "nicely",
                          legend.font.size = 5, text.size = 2.4, 
                          vertex.size = 8, vertex_connections_cap = 3, 
                          edge_size_range = c(0.6,1.4), 
                          arrows = arrow(type = "closed", angle = 25, length = unit(6, "pt")))

celltypes_connection_plot(lgals1_ptprc_dat %>% filter(n>=2),
                          celltype_colors_table = cellcolors, 
                          layout = function(g)convert(g, to_undirected) %>%
                            ggraph::create_layout("backbone") %>% select(x,y), 
                          legend.font.size = 5, text.size = 2.4, 
                          vertex.size = 8, vertex_connections_cap = 3, 
                          edge_size_range = c(0.6,1.4), 
                          arrows = arrow(type = "closed", angle = 25, length = unit(6, "pt")))

celltypes_connection_plot(lgals1_ptprc_dat %>% filter(n>=3),
                          celltype_colors_table = cellcolors, 
                          layout = function(g)convert(g, to_undirected) %>%
                            ggraph::create_layout("backbone") %>% select(x,y),
                          legend.font.size = 5, text.size = 2.4, 
                          vertex.size = 8, vertex_connections_cap = 3, 
                          edge_size_range = c(0.6,1.4), 
                          arrows = arrow(type = "closed", angle = 25, length = unit(6, "pt")))


# for computing the aspect ratios
fixed_asp_pct = 0.13
variable_asp_pct = 0.021

asp.nc.2 = fixed_asp_pct + length(unique(min2_melnc$LR))*variable_asp_pct
asp.meltrans.2 = fixed_asp_pct + length(unique(min2_meltrans$LR))*variable_asp_pct
asp.melan.2 = fixed_asp_pct + length(unique(min2_meltransmelan$LR))*variable_asp_pct
asp.nc.3 = fixed_asp_pct + length(unique(min3_melnc$LR))*variable_asp_pct # 0.61 per 21 good
asp.meltrans.3 = fixed_asp_pct + length(unique(min3_meltrans$LR))*variable_asp_pct
asp.melan.3 = fixed_asp_pct + length(unique(min3_meltransmelan$LR))*variable_asp_pct
```

### Combined Tables Melanoma -> T/NK cells, min 2 samples support

```{r hd-liana-tables-min2-Mel_nc-like, fig.width=4., fig.asp=asp.nc.2, fig.show = 'hold', echo=T, eval=T}
bysample_dotplot(min2_melnc, 
  cellcolors = cellcolors, highlight = highlight_colors,
  Sample_order = names(liana_aggs), wrap_target_xaxis_sample = T,
  legend.key.width = unit(4, "pt"), legend.key.height = unit(1, "null"))
```

```{r hd-liana-tables-min2-Mel_trans, fig.width=4., fig.asp=asp.meltrans.2, fig.show = 'hold', echo=T, eval=T}
bysample_dotplot(min2_meltrans,
  cellcolors = cellcolors, highlight = highlight_colors,
  Sample_order = names(liana_aggs), wrap_target_xaxis_sample = T,
  legend.key.width = unit(4, "pt"), legend.key.height = unit(1, "null"))
```

```{r hd-liana-tables-min2-Mel_trans_melan, fig.width=4., fig.asp=asp.melan.2, fig.show = 'hold', echo=T, eval=T}
bysample_dotplot(min2_meltransmelan,
  cellcolors = cellcolors, highlight = highlight_colors,
  Sample_order = names(liana_aggs), wrap_target_xaxis_sample = T,
  legend.key.width = unit(4, "pt"), legend.key.height = unit(1, "null"))
```

### Combined Tables Melanoma -> T/NK cells, min 3 samples support

```{r hd-liana-tables-min3-Mel_nc-like, fig.width=3.54, fig.asp=asp.nc.3, fig.show = 'hold', echo=T, eval=T}
bysample_dotplot(min3_melnc, 
  cellcolors = cellcolors, highlight = highlight_colors,
  Sample_order = names(liana_aggs), wrap_target_xaxis_sample = T,
  legend.key.width = unit(4, "pt"), legend.key.height = unit(1, "null"))
```

```{r hd-liana-tables-min3-Mel_trans, fig.width=3.54, fig.asp=asp.meltrans.3, fig.show = 'hold', echo=T, eval=T}
bysample_dotplot(min3_meltrans,
  cellcolors = cellcolors, highlight = highlight_colors,
  Sample_order = names(liana_aggs), wrap_target_xaxis_sample = T,
  legend.key.width = unit(4, "pt"), legend.key.height = unit(1, "null"))
```

```{r hd-liana-tables-min3-Mel_trans_melan, fig.width=3.54, fig.asp=asp.melan.3, fig.show = 'hold', echo=T, eval=T}
bysample_dotplot(min3_meltransmelan,
  cellcolors = cellcolors, highlight = highlight_colors,
  Sample_order = names(liana_aggs), wrap_target_xaxis_sample = T,
  legend.key.width = unit(4, "pt"), legend.key.height = unit(1, "null"))
```

## NICHES

### Sample HD_PM3 for use in the main figures

```{r hd-niches-main, fig.width=16., fig.asp=0.7*0.88, fig.show = 'hold', echo=T, eval=T}
  s = "HD_PM3"

  curr_obj = hd_objs_processed[[s]]
  df = curr_obj$NICHES_output$CellToNeighborhood@meta.data %>%
    mutate(object_id = as.numeric(SendingCell)) %>%
    left_join(curr_obj$obs_for_alluvials) %>%
    mutate(x = x_tr, y = y_tr, 
           filt = celltype_broad %in% c("Mel", "T", "NK"),
           celltype_spider = setNames(celltype_spiderplots_collapsing[as.character(SendingType)], nm=NULL))
  
  df_edges = curr_obj$NICHES_output$CellToCellSpatial@meta.data %>%
    mutate(edge_id = colnames(curr_obj$NICHES_output$CellToCellSpatial),
           value = curr_obj$NICHES_output$CellToCellSpatial["LGALS1—PTPRC",]@assays$CellToCellSpatial@layers$data)
  df_edges[c("sending_id", "receiving_id")] = str_split_fixed(df_edges$edge_id, pattern = "—", n = 2)
  df_edges = df_edges %>%
    left_join(df %>% transmute(sending_id = SendingCell, x.Sending = x, y.Sending = y)) %>%
    left_join(df %>% transmute(receiving_id = SendingCell, x.Receiving = x, y.Receiving = y))
 
  avg_edge_values = df_edges %>% 
    group_by(sending_id) %>%
    summarize(n = n(), valuesum = sum(value)) %>%
    inner_join(df_edges %>% 
    group_by(receiving_id) %>%
    summarize(n = n(), valuesum = sum(value)), 
    by = join_by(sending_id == receiving_id)) %>%
    transmute(id = sending_id,
              value = (valuesum.x + valuesum.y)/(n.x + n.y -1))

  df = df %>% left_join(avg_edge_values, join_by(SendingCell == id))
  df_edges_pos = df_edges %>% filter(value>0)

# why are these values so weird? I choose nice 0.9-1 µm edge length 
# frames, but in the wrong coordinate system, and just translated those to 
# pixel coordinates +- rounding to nearest integer. Now they no longer look nice 
  
inset_xmin = 8711
inset_xmax = 9611
inset_ymin = 7351
inset_ymax = 8251

inset2_xmin = 4989
inset2_xmax = 5989
inset2_ymin = 6408
inset2_ymax = 7408

edgeval_lims = c(0., quantile(df$value[df$value>0], p = c(0.995)))
edgeval_lims = ceiling(10*edgeval_lims)/10

p_maxi = ggplot(df, aes(x = x, y = y)) +
  mytheme() +
  theme(axis.text = element_blank(), axis.title = element_blank(),
        axis.ticks = element_blank(), legend.box.spacing = unit(8, "pt"), legend.text = element_text(size = 12),
        legend.position = "inside", legend.position.inside = c(0.5, 0.02), legend.background = element_rect(fill=alpha("white", 0.8)),
        legend.direction = "horizontal", legend.justification = c(0.5, 0.), legend.box = "horizontal", 
        legend.justification.inside = "bottom", legend.title.position = "top", legend.title = element_text(face = "bold", size = 14), legend.justification.bottom = "left") +
  geom_node_voronoi(mapping = aes(fill=bordering), normalize = F, max.radius = 50, color = NA, alpha = 0.2) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region", guide = guide_legend(nrow = 4, order = 2)) +
  geom_point(aes(alpha = value, color = value), size = 3, stroke = 0.) +
  continuous_scale("alpha", name = waiver(), palette = rescale_saturating(range = c(0., 0.95), rel_saturating = 0.45),
                   limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish) +
  guides(alpha = guide_bins("Cell Average Interaction Value", order = 3,
                   theme = theme(legend.key.width = rel(1.3),
                                 legend.key = element_rect(fill = pattern, color = NA))),
         color = guide_bins("Cell Average Interaction Value", show.limits = T, order = 3,
                   override.aes = list(size = 6))) +
  new_scale_color() +
  geom_point(mapping = aes(x,y,color = celltype_spider), size = 0.08) +
  scale_color_manual(breaks = names(cellcolors), values = cellcolors, name = "Celltype",
                     guide = guide_legend(override.aes = list(size = 4), nrow = 8, order = 1)) +
  coord_equal(expand = F, xlim = range(df$x), ylim = range(df$y)) +
  annotate("segment", x = c(inset_xmin, inset_xmax), y = c(inset_ymin, inset_ymin), yend = c(inset_ymax, inset_ymax)) + # frame for inset
  annotate("segment", y = c(inset_ymin, inset_ymax), x = c(inset_xmin, inset_xmin), xend = c(inset_xmax, inset_xmax)) +
  annotate("segment", x = c(inset2_xmin, inset2_xmax), y = c(inset2_ymin, inset2_ymin), 
           yend = c(inset2_ymax, inset2_ymax), colour = "#1122AF") + # frame for inset
  annotate("segment", y = c(inset2_ymin, inset2_ymax), x = c(inset2_xmin, inset2_xmin), 
           xend = c(inset2_xmax, inset2_xmax), colour = "#1122AF") +
  plot_scalebars(max(df$x)-1200, max(df$y)-300, length = 1000, label = "1mm",
                 background_height_adjust = 70, tickheight_major = 30, background_expansion = 16,
                 tickheight_minor = 12, ticks_minor = c(0.2, 0.4, 0.6, 0.8))
  
p_mini_1 = ggplot(df_edges_pos, aes(x = x.Sending, y = y.Sending,
                           xend = x.Receiving, yend = y.Receiving)) +
  geom_node_voronoi(data = df, mapping = aes(fill=bordering),
                    normalize = F, max.radius = 25, color = NA, alpha = 0.2, show.legend = F) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region") +
  new_scale_fill() +
  geom_point(data = df, mapping = aes(x,y, alpha = value, color = value),
             inherit.aes = F, size = 3., show.legend = F) +
  scale_alpha_continuous(range = c(0., 1.), limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish, ) +
  new_scale_color() +
  geom_segment(mapping = aes(color = value), linewidth = 0.8) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = c(0., NA), name = "NICHES\nInteraction\nValue\n(Edges)") +
  new_scale_color() +
  geom_point(data = df, mapping = aes(x,y,color = celltype_spider),
             inherit.aes = F, size = 0.8, show.legend = F) +
  scale_color_manual(values = cellcolors) +
  mytheme() +
  coord_equal(xlim = c(inset_xmin, inset_xmax), ylim = c(inset_ymin, inset_ymax), expand = F) +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = "inside", legend.position.inside = c(.99, .99), legend.justification.inside = c(1, 1),
        legend.background = element_rect(fill=alpha("white", 0.8)), legend.key.height = unit(2, "lines"),
        plot.margin = margin(0,0,0,0, unit = "pt"), panel.background = element_blank(), legend.text = element_text(size = 12), legend.margin = margin(2,2,4,2, "pt"),
        panel.border = element_rect(fill = NA, color = "black", linewidth = 2), legend.title = element_text(face = "bold", size = 14)) +
  plot_scalebars(inset_xmax-220, inset_ymin+20)


p_mini_2 = ggplot(df_edges_pos, aes(x = x.Sending, y = y.Sending,
                           xend = x.Receiving, yend = y.Receiving)) +
  geom_node_voronoi(data = df, mapping = aes(fill=bordering),
                    normalize = F, max.radius = 25, color = NA, alpha = 0.2, show.legend = F) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region") +
  new_scale_fill() +
  geom_point(data = df, mapping = aes(x,y, alpha = value, color = value),
             inherit.aes = F, size = 3., show.legend = F) +
  scale_alpha_continuous(range = c(0., 1.), limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish, ) +
  new_scale_color() +
  geom_segment(mapping = aes(color = value), linewidth = 0.8) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = c(0., NA), name = "NICHES\nInteraction\nValue\n(Edges)") +
  new_scale_color() +
  geom_point(data = df, mapping = aes(x,y,color = celltype_spider),
             inherit.aes = F, size = 0.8, show.legend = F) +
  scale_color_manual(values = cellcolors) +
  mytheme() +
  coord_equal(xlim = c(inset2_xmin, inset2_xmax), ylim = c(inset2_ymin, inset2_ymax), expand = F) +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = "inside", legend.position.inside = c(.99, .99), legend.justification.inside = c(1, 1),
        legend.background = element_rect(fill=alpha("white", 0.8)), legend.key.height = unit(2, "lines"),
        plot.margin = margin(0,0,0,0, unit = "pt"), panel.background = element_blank(), legend.text = element_text(size = 12), legend.margin = margin(2,2,4,2, "pt"),
        panel.border = element_rect(fill = NA, color = "#1122AF", linewidth = 2), legend.title = element_text(face = "bold", size = 14)) +
  plot_scalebars(inset2_xmax-220, inset2_ymin+20, background_expansion = 5)

wrap_plots(a = p_maxi, b = p_mini_1, c = p_mini_2,
           design = "
           aab
           aac
           ")
```

### HD_PM1

```{r hd-niches-1, fig.width=16., fig.asp=0.7, fig.show = 'hold', echo=T, eval=T}
  s = "HD_PM1"

  curr_obj = hd_objs_processed[[s]]
  df = curr_obj$NICHES_output$CellToNeighborhood@meta.data %>%
    mutate(object_id = as.numeric(SendingCell)) %>%
    left_join(curr_obj$obs_for_alluvials) %>%
    mutate(x = x_tr, y = y_tr, 
           filt = celltype_broad %in% c("Mel", "T", "NK"),
           celltype_spider = setNames(celltype_spiderplots_collapsing[as.character(SendingType)], nm=NULL))
  
  df_edges = curr_obj$NICHES_output$CellToCellSpatial@meta.data %>%
    mutate(edge_id = colnames(curr_obj$NICHES_output$CellToCellSpatial),
           value = curr_obj$NICHES_output$CellToCellSpatial["LGALS1—PTPRC",]@assays$CellToCellSpatial@layers$data)
  df_edges[c("sending_id", "receiving_id")] = str_split_fixed(df_edges$edge_id, pattern = "—", n = 2)
  df_edges = df_edges %>%
    left_join(df %>% transmute(sending_id = SendingCell, x.Sending = x, y.Sending = y)) %>%
    left_join(df %>% transmute(receiving_id = SendingCell, x.Receiving = x, y.Receiving = y))
 
  avg_edge_values = df_edges %>% 
    group_by(sending_id) %>%
    summarize(n = n(), valuesum = sum(value)) %>%
    inner_join(df_edges %>% 
    group_by(receiving_id) %>%
    summarize(n = n(), valuesum = sum(value)), 
    by = join_by(sending_id == receiving_id)) %>%
    transmute(id = sending_id,
              value = (valuesum.x + valuesum.y)/(n.x + n.y -1))

  df = df %>% left_join(avg_edge_values, join_by(SendingCell == id))
  df_edges_pos = df_edges %>% filter(value>0)

inset_xmin = 3880
inset_xmax = 4980
inset_ymin = 4550
inset_ymax = 5650

inset2_xmin = 9905
inset2_xmax = 10805
inset2_ymin = 5445
inset2_ymax = 6345

edgeval_lims = c(0., quantile(df$value[df$value>0], p = c(0.995)))
edgeval_lims = ceiling(10*edgeval_lims)/10

p_maxi = ggplot(df, aes(x = x, y = y)) +
  mytheme() +
  theme(axis.text = element_blank(), axis.title = element_blank(),
        axis.ticks = element_blank(), legend.box.spacing = unit(8, "pt"), legend.text = element_text(size = 12),
        legend.position = "bottom", legend.direction = "horizontal", legend.box.margin = margin(l = 2, unit = "pt"),
        legend.title.position = "top", legend.title = element_text(face = "bold", size = 14), legend.justification.bottom = "left") +
  geom_node_voronoi(mapping = aes(fill=bordering), normalize = F, max.radius = 50, color = NA, alpha = 0.2) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region", guide = guide_legend(nrow = 2, order = 2)) +
  geom_point(aes(alpha = value, color = value), size = 3, stroke = 0.) +
  continuous_scale("alpha", name = waiver(), palette = rescale_saturating(range = c(0., 0.95), rel_saturating = 0.45),
                   limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish) +
  guides(alpha = guide_bins("Cell Average Interaction Value", order = 3,
                   theme = theme(legend.key.width = rel(1.3),
                                 legend.key = element_rect(fill = pattern, color = NA))),
         color = guide_bins("Cell Average Interaction Value", show.limits = T, order = 3,
                   override.aes = list(size = 6))) +
  new_scale_color() +
  geom_point(mapping = aes(x,y,color = celltype_spider), size = 0.08) +
  scale_color_manual(breaks = names(cellcolors), values = cellcolors, name = "Celltype",
                     guide = guide_legend(override.aes = list(size = 4), nrow = 3, order = 1)) +
  coord_equal(expand = F, xlim = range(df$x), ylim = range(df$y)) +
  annotate("segment", x = c(inset_xmin, inset_xmax), y = c(inset_ymin, inset_ymin), yend = c(inset_ymax, inset_ymax)) + # frame for inset
  annotate("segment", y = c(inset_ymin, inset_ymax), x = c(inset_xmin, inset_xmin), xend = c(inset_xmax, inset_xmax)) +
  annotate("segment", x = c(inset2_xmin, inset2_xmax), y = c(inset2_ymin, inset2_ymin), 
           yend = c(inset2_ymax, inset2_ymax), colour = "#1122AF") + # frame for inset
  annotate("segment", y = c(inset2_ymin, inset2_ymax), x = c(inset2_xmin, inset2_xmin), 
           xend = c(inset2_xmax, inset2_xmax), colour = "#1122AF") +
  plot_scalebars(max(df$x)-1200, max(df$y)-300, length = 1000, label = "1mm",
                 background_height_adjust = 70, tickheight_major = 30, background_expansion = 16,
                 tickheight_minor = 12, ticks_minor = c(0.2, 0.4, 0.6, 0.8)) 
  
p_mini_1 = ggplot(df_edges_pos, aes(x = x.Sending, y = y.Sending,
                           xend = x.Receiving, yend = y.Receiving)) +
  geom_node_voronoi(data = df, mapping = aes(fill=bordering),
                    normalize = F, max.radius = 25, color = NA, alpha = 0.2, show.legend = F) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region") +
  new_scale_fill() +
  geom_point(data = df, mapping = aes(x,y, alpha = value, color = value),
             inherit.aes = F, size = 3., show.legend = F) +
  scale_alpha_continuous(range = c(0., 1.), limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish, ) +
  new_scale_color() +
  geom_segment(mapping = aes(color = value), linewidth = 0.8) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = c(0., NA), name = "NICHES\nInteraction\nValue\n(Edges)") +
  new_scale_color() +
  geom_point(data = df, mapping = aes(x,y,color = celltype_spider),
             inherit.aes = F, size = 0.8, show.legend = F) +
  scale_color_manual(values = cellcolors) +
  mytheme() +
  coord_equal(xlim = c(inset_xmin, inset_xmax), ylim = c(inset_ymin, inset_ymax), expand = F) +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = "inside", legend.position.inside = c(.99, .99), legend.justification.inside = c(1, 1),
        legend.background = element_rect(fill=alpha("white", 0.8)), legend.key.height = unit(2, "lines"), legend.margin = margin(2,2,4,2, "pt"),
        plot.margin = margin(0,0,0,0, unit = "pt"), panel.background = element_blank(), legend.text = element_text(size = 12),
        panel.border = element_rect(fill = NA, color = "black", linewidth = 2), legend.title = element_text(face = "bold", size = 14)) +
  plot_scalebars(inset_xmax-210, inset_ymin+20)


p_mini_2 = ggplot(df_edges_pos, aes(x = x.Sending, y = y.Sending,
                           xend = x.Receiving, yend = y.Receiving)) +
  geom_node_voronoi(data = df, mapping = aes(fill=bordering),
                    normalize = F, max.radius = 25, color = NA, alpha = 0.2, show.legend = F) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region") +
  new_scale_fill() +
  geom_point(data = df, mapping = aes(x,y, alpha = value, color = value),
             inherit.aes = F, size = 3., show.legend = F) +
  scale_alpha_continuous(range = c(0., 1.), limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish, ) +
  new_scale_color() +
  geom_segment(mapping = aes(color = value), linewidth = 0.8) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = c(0., NA), name = "NICHES\nInteraction\nValue\n(Edges)") +
  new_scale_color() +
  geom_point(data = df, mapping = aes(x,y,color = celltype_spider),
             inherit.aes = F, size = 0.8, show.legend = F) +
  scale_color_manual(values = cellcolors) +
  mytheme() +
  coord_equal(xlim = c(inset2_xmin, inset2_xmax), ylim = c(inset2_ymin, inset2_ymax), expand = F) +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = "inside", legend.position.inside = c(.99, .99), legend.justification.inside = c(1, 1), legend.margin = margin(2,2,4,2, "pt"),
        legend.background = element_rect(fill=alpha("white", 0.8)), legend.key.height = unit(2, "lines"),
        plot.margin = margin(0,0,0,0, unit = "pt"), panel.background = element_blank(), legend.text = element_text(size = 12),
        panel.border = element_rect(fill = NA, color = "#1122AF", linewidth = 2), legend.title = element_text(face = "bold", size = 14)) +
  plot_scalebars(inset2_xmax-210, inset2_ymin+20)

wrap_plots(a = p_maxi, b = p_mini_1, c = p_mini_2,
           design = "
           aab
           aac
           ")
```

### HD_PM2

```{r hd-niches-2, fig.width=16., fig.asp=0.7, fig.show = 'hold', echo=T, eval=T}
  s = "HD_PM2"

  curr_obj = hd_objs_processed[[s]]
  df = curr_obj$NICHES_output$CellToNeighborhood@meta.data %>%
    mutate(object_id = as.numeric(SendingCell)) %>%
    left_join(curr_obj$obs_for_alluvials) %>%
    mutate(x = x_tr, y = y_tr, 
           filt = celltype_broad %in% c("Mel", "T", "NK"),
           celltype_spider = setNames(celltype_spiderplots_collapsing[as.character(SendingType)], nm=NULL))
  
  df_edges = curr_obj$NICHES_output$CellToCellSpatial@meta.data %>%
    mutate(edge_id = colnames(curr_obj$NICHES_output$CellToCellSpatial),
           value = curr_obj$NICHES_output$CellToCellSpatial["LGALS1—PTPRC",]@assays$CellToCellSpatial@layers$data)
  df_edges[c("sending_id", "receiving_id")] = str_split_fixed(df_edges$edge_id, pattern = "—", n = 2)
  df_edges = df_edges %>%
    left_join(df %>% transmute(sending_id = SendingCell, x.Sending = x, y.Sending = y)) %>%
    left_join(df %>% transmute(receiving_id = SendingCell, x.Receiving = x, y.Receiving = y))
 
  avg_edge_values = df_edges %>% 
    group_by(sending_id) %>%
    summarize(n = n(), valuesum = sum(value)) %>%
    inner_join(df_edges %>% 
    group_by(receiving_id) %>%
    summarize(n = n(), valuesum = sum(value)), 
    by = join_by(sending_id == receiving_id)) %>%
    transmute(id = sending_id,
              value = (valuesum.x + valuesum.y)/(n.x + n.y -1))

  df = df %>% left_join(avg_edge_values, join_by(SendingCell == id))
  df_edges_pos = df_edges %>% filter(value>0)

inset_xmin = 11324
inset_xmax = 12324
inset_ymin = 5952
inset_ymax = 6952

inset2_xmin = 8541
inset2_xmax = 9541
inset2_ymin = 6130
inset2_ymax = 7130

edgeval_lims = c(0., quantile(df$value[df$value>0], p = c(0.995)))
edgeval_lims = ceiling(10*edgeval_lims)/10

p_maxi = ggplot(df, aes(x = x, y = y)) +
  mytheme() +
  theme(axis.text = element_blank(), axis.title = element_blank(),
        axis.ticks = element_blank(), legend.box.spacing = unit(8, "pt"), legend.text = element_text(size = 12),
        legend.position = "bottom", legend.direction = "horizontal", legend.box.margin = margin(l = 2, unit = "pt"),
        legend.title.position = "top", legend.title = element_text(face = "bold", size = 14), legend.justification.bottom = "left") +
  geom_node_voronoi(mapping = aes(fill=bordering), normalize = F, max.radius = 50, color = NA, alpha = 0.2) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region", guide = guide_legend(nrow = 2, order = 2)) +
  geom_point(aes(alpha = value, color = value), size = 3, stroke = 0.) +
  continuous_scale("alpha", name = waiver(), palette = rescale_saturating(range = c(0., 0.9), rel_saturating = 0.4),
                   limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish) +
  guides(alpha = guide_bins("Cell Average Interaction Value", order = 3,
                   theme = theme(legend.key.width = rel(1.3),
                                 legend.key = element_rect(fill = pattern, color = NA))),
         color = guide_bins("Cell Average Interaction Value", show.limits = T, order = 3,
                   override.aes = list(size = 6))) +
  new_scale_color() +
  geom_point(mapping = aes(x,y,color = celltype_spider), size = 0.08) +
  scale_color_manual(breaks = names(cellcolors), values = cellcolors, name = "Celltype",
                     guide = guide_legend(override.aes = list(size = 4), nrow = 3, order = 1)) +
  coord_equal(expand = F, xlim = range(df$x), ylim = range(df$y)) +
  annotate("segment", x = c(inset_xmin, inset_xmax), y = c(inset_ymin, inset_ymin), yend = c(inset_ymax, inset_ymax)) + # frame for inset
  annotate("segment", y = c(inset_ymin, inset_ymax), x = c(inset_xmin, inset_xmin), xend = c(inset_xmax, inset_xmax)) +
  annotate("segment", x = c(inset2_xmin, inset2_xmax), y = c(inset2_ymin, inset2_ymin), 
           yend = c(inset2_ymax, inset2_ymax), colour = "#1122AF") + # frame for inset
  annotate("segment", y = c(inset2_ymin, inset2_ymax), x = c(inset2_xmin, inset2_xmin), 
           xend = c(inset2_xmax, inset2_xmax), colour = "#1122AF") +
  plot_scalebars(max(df$x)-1200, max(df$y)-300, length = 1000, label = "1mm",
                 background_height_adjust = 70, tickheight_major = 30, background_expansion = 16,
                 tickheight_minor = 12, ticks_minor = c(0.2, 0.4, 0.6, 0.8)) 
  
p_mini_1 = ggplot(df_edges_pos, aes(x = x.Sending, y = y.Sending,
                           xend = x.Receiving, yend = y.Receiving)) +
  geom_node_voronoi(data = df, mapping = aes(fill=bordering),
                    normalize = F, max.radius = 25, color = NA, alpha = 0.2, show.legend = F) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region") +
  new_scale_fill() +
  geom_point(data = df, mapping = aes(x,y, alpha = value, color = value),
             inherit.aes = F, size = 3., show.legend = F) +
  scale_alpha_continuous(range = c(0., 1.), limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish, ) +
  new_scale_color() +
  geom_segment(mapping = aes(color = value), linewidth = 0.8) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = c(0., NA), name = "NICHES\nInteraction\nValue\n(Edges)") +
  new_scale_color() +
  geom_point(data = df, mapping = aes(x,y,color = celltype_spider),
             inherit.aes = F, size = 0.8, show.legend = F) +
  scale_color_manual(values = cellcolors) +
  mytheme() +
  coord_equal(xlim = c(inset_xmin, inset_xmax), ylim = c(inset_ymin, inset_ymax), expand = F) +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = "inside", legend.position.inside = c(.99, .99), legend.justification.inside = c(1, 1),
        legend.background = element_rect(fill=alpha("white", 0.8)), legend.key.height = unit(2, "lines"), legend.margin = margin(2,2,4,2, "pt"),
        plot.margin = margin(0,0,0,0, unit = "pt"), panel.background = element_blank(), legend.text = element_text(size = 12),
        panel.border = element_rect(fill = NA, color = "black", linewidth = 2), legend.title = element_text(face = "bold", size = 14)) +
  plot_scalebars(inset_xmax-210, inset_ymin+20)


p_mini_2 = ggplot(df_edges_pos, aes(x = x.Sending, y = y.Sending,
                           xend = x.Receiving, yend = y.Receiving)) +
  geom_node_voronoi(data = df, mapping = aes(fill=bordering),
                    normalize = F, max.radius = 25, color = NA, alpha = 0.2, show.legend = F) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region") +
  new_scale_fill() +
  geom_point(data = df, mapping = aes(x,y, alpha = value, color = value),
             inherit.aes = F, size = 3., show.legend = F) +
  scale_alpha_continuous(range = c(0., 1.), limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish, ) +
  new_scale_color() +
  geom_segment(mapping = aes(color = value), linewidth = 0.8) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = c(0., NA), name = "NICHES\nInteraction\nValue\n(Edges)") +
  new_scale_color() +
  geom_point(data = df, mapping = aes(x,y,color = celltype_spider),
             inherit.aes = F, size = 0.8, show.legend = F) +
  scale_color_manual(values = cellcolors) +
  mytheme() +
  coord_equal(xlim = c(inset2_xmin, inset2_xmax), ylim = c(inset2_ymin, inset2_ymax), expand = F) +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = "inside", legend.position.inside = c(.99, .99), legend.justification.inside = c(1, 1),
        legend.background = element_rect(fill=alpha("white", 0.8)), legend.key.height = unit(2, "lines"), legend.margin = margin(2,2,4,2, "pt"),
        plot.margin = margin(0,0,0,0, unit = "pt"), panel.background = element_blank(), legend.text = element_text(size = 12),
        panel.border = element_rect(fill = NA, color = "#1122AF", linewidth = 2), legend.title = element_text(face = "bold", size = 14)) +
  plot_scalebars(inset2_xmax-210, inset2_ymin+20)

wrap_plots(a = p_maxi, b = p_mini_1, c = p_mini_2,
           design = "
           aab
           aac
           ")
```

### HD_PM4

```{r hd-niches-3, fig.width=16., fig.asp=0.7, fig.show = 'hold', echo=T, eval=T}
  s = "HD_PM4"

  curr_obj = hd_objs_processed[[s]]
  df = curr_obj$NICHES_output$CellToNeighborhood@meta.data %>%
    mutate(object_id = as.numeric(SendingCell)) %>%
    left_join(curr_obj$obs_for_alluvials) %>%
    mutate(x = x_tr, y = y_tr, 
           filt = celltype_broad %in% c("Mel", "T", "NK"),
           celltype_spider = setNames(celltype_spiderplots_collapsing[as.character(SendingType)], nm=NULL))
  
  df_edges = curr_obj$NICHES_output$CellToCellSpatial@meta.data %>%
    mutate(edge_id = colnames(curr_obj$NICHES_output$CellToCellSpatial),
           value = curr_obj$NICHES_output$CellToCellSpatial["LGALS1—PTPRC",]@assays$CellToCellSpatial@layers$data)
  df_edges[c("sending_id", "receiving_id")] = str_split_fixed(df_edges$edge_id, pattern = "—", n = 2)
  df_edges = df_edges %>%
    left_join(df %>% transmute(sending_id = SendingCell, x.Sending = x, y.Sending = y)) %>%
    left_join(df %>% transmute(receiving_id = SendingCell, x.Receiving = x, y.Receiving = y))
 
  avg_edge_values = df_edges %>% 
    group_by(sending_id) %>%
    summarize(n = n(), valuesum = sum(value)) %>%
    inner_join(df_edges %>% 
    group_by(receiving_id) %>%
    summarize(n = n(), valuesum = sum(value)), 
    by = join_by(sending_id == receiving_id)) %>%
    transmute(id = sending_id,
              value = (valuesum.x + valuesum.y)/(n.x + n.y -1))

  df = df %>% left_join(avg_edge_values, join_by(SendingCell == id))
  df_edges_pos = df_edges %>% filter(value>0)

inset_xmin = 7008
inset_xmax = 8018
inset_ymin = 5128
inset_ymax = 6138

inset2_xmin = 4823
inset2_xmax = 5823
inset2_ymin = 2682
inset2_ymax = 3682

edgeval_lims = c(0., quantile(df$value[df$value>0], p = c(0.995)))
edgeval_lims = ceiling(10*edgeval_lims)/10

p_maxi = ggplot(df, aes(x = x, y = y)) +
  mytheme() +
  theme(axis.text = element_blank(), axis.title = element_blank(),
        axis.ticks = element_blank(), legend.box.spacing = unit(8, "pt"), legend.text = element_text(size = 12),
        legend.position = "bottom", legend.direction = "horizontal", legend.box.margin = margin(l = 2, unit = "pt"),
        legend.title.position = "top", legend.title = element_text(face = "bold", size = 14), legend.justification.bottom = "left") +
  geom_node_voronoi(mapping = aes(fill=bordering), normalize = F, max.radius = 50, color = NA, alpha = 0.2) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region", guide = guide_legend(nrow = 2, order = 2)) +
  geom_point(aes(alpha = value, color = value), size = 3, stroke = 0.) +
  continuous_scale("alpha", name = waiver(), palette = rescale_saturating(range = c(0., 0.95), rel_saturating = 0.45),
                   limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish) +
  guides(alpha = guide_bins("Cell Average Interaction Value", order = 3,
                   theme = theme(legend.key.width = rel(1.3),
                                 legend.key = element_rect(fill = pattern, color = NA))),
         color = guide_bins("Cell Average Interaction Value", show.limits = T, order = 3,
                   override.aes = list(size = 6))) +
  new_scale_color() +
  geom_point(mapping = aes(x,y,color = celltype_spider), size = 0.08) +
  scale_color_manual(breaks = names(cellcolors), values = cellcolors, name = "Celltype",
                     guide = guide_legend(override.aes = list(size = 4), nrow = 3, order = 1)) +
  coord_equal(expand = F, xlim = range(df$x), ylim = range(df$y)) +
  annotate("segment", x = c(inset_xmin, inset_xmax), y = c(inset_ymin, inset_ymin), yend = c(inset_ymax, inset_ymax)) + # frame for inset
  annotate("segment", y = c(inset_ymin, inset_ymax), x = c(inset_xmin, inset_xmin), xend = c(inset_xmax, inset_xmax)) +
  annotate("segment", x = c(inset2_xmin, inset2_xmax), y = c(inset2_ymin, inset2_ymin), 
           yend = c(inset2_ymax, inset2_ymax), colour = "#1122AF") + # frame for inset
  annotate("segment", y = c(inset2_ymin, inset2_ymax), x = c(inset2_xmin, inset2_xmin), 
           xend = c(inset2_xmax, inset2_xmax), colour = "#1122AF") +
  plot_scalebars(max(df$x)-1200, max(df$y)-300, length = 1000, label = "1mm",
                 background_height_adjust = 70, tickheight_major = 30, background_expansion = 16,
                 tickheight_minor = 12, ticks_minor = c(0.2, 0.4, 0.6, 0.8)) 
  
p_mini_1 = ggplot(df_edges_pos, aes(x = x.Sending, y = y.Sending,
                           xend = x.Receiving, yend = y.Receiving)) +
  geom_node_voronoi(data = df, mapping = aes(fill=bordering),
                    normalize = F, max.radius = 25, color = NA, alpha = 0.2, show.legend = F) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region") +
  new_scale_fill() +
  geom_point(data = df, mapping = aes(x,y, alpha = value, color = value),
             inherit.aes = F, size = 3., show.legend = F) +
  scale_alpha_continuous(range = c(0., 1.), limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish, ) +
  new_scale_color() +
  geom_segment(mapping = aes(color = value), linewidth = 0.8) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = c(0., NA), name = "NICHES\nInteraction\nValue\n(Edges)") +
  new_scale_color() +
  geom_point(data = df, mapping = aes(x,y,color = celltype_spider),
             inherit.aes = F, size = 0.8, show.legend = F) +
  scale_color_manual(values = cellcolors) +
  mytheme() +
  coord_equal(xlim = c(inset_xmin, inset_xmax), ylim = c(inset_ymin, inset_ymax), expand = F) +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = "inside", legend.position.inside = c(.99, .99), legend.justification.inside = c(1, 1),
        legend.background = element_rect(fill=alpha("white", 0.8)), legend.key.height = unit(2, "lines"), legend.margin = margin(2,2,4,2, "pt"),
        plot.margin = margin(0,0,0,0, unit = "pt"), panel.background = element_blank(), legend.text = element_text(size = 12),
        panel.border = element_rect(fill = NA, color = "black", linewidth = 2), legend.title = element_text(face = "bold", size = 14)) +
  plot_scalebars(inset_xmax-210, inset_ymin+20)


p_mini_2 = ggplot(df_edges_pos, aes(x = x.Sending, y = y.Sending,
                           xend = x.Receiving, yend = y.Receiving)) +
  geom_node_voronoi(data = df, mapping = aes(fill=bordering),
                    normalize = F, max.radius = 25, color = NA, alpha = 0.2, show.legend = F) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region") +
  new_scale_fill() +
  geom_point(data = df, mapping = aes(x,y, alpha = value, color = value),
             inherit.aes = F, size = 3., show.legend = F) +
  scale_alpha_continuous(range = c(0., 1.), limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish, ) +
  new_scale_color() +
  geom_segment(mapping = aes(color = value), linewidth = 0.8) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = c(0., NA), name = "NICHES\nInteraction\nValue\n(Edges)") +
  new_scale_color() +
  geom_point(data = df, mapping = aes(x,y,color = celltype_spider),
             inherit.aes = F, size = 0.8, show.legend = F) +
  scale_color_manual(values = cellcolors) +
  mytheme() +
  coord_equal(xlim = c(inset2_xmin, inset2_xmax), ylim = c(inset2_ymin, inset2_ymax), expand = F) +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = "inside", legend.position.inside = c(.99, .99), legend.justification.inside = c(1, 1),
        legend.background = element_rect(fill=alpha("white", 0.8)), legend.key.height = unit(2, "lines"), legend.margin = margin(2,2,4,2, "pt"),
        plot.margin = margin(0,0,0,0, unit = "pt"), panel.background = element_blank(), legend.text = element_text(size = 12),
        panel.border = element_rect(fill = NA, color = "#1122AF", linewidth = 2), legend.title = element_text(face = "bold", size = 14)) +
  plot_scalebars(inset2_xmax-210, inset2_ymin+20)

wrap_plots(a = p_maxi, b = p_mini_1, c = p_mini_2,
           design = "
           aab
           aac
           ")
```

### HD_MM1

```{r hd-niches-4, fig.width=16., fig.asp=0.7, fig.show = 'hold', echo=T, eval=T}
  s = "HD_MM1"

  curr_obj = hd_objs_processed[[s]]
  df = curr_obj$NICHES_output$CellToNeighborhood@meta.data %>%
    mutate(object_id = as.numeric(SendingCell)) %>%
    left_join(curr_obj$obs_for_alluvials) %>%
    mutate(x = x_tr, y = y_tr, 
           filt = celltype_broad %in% c("Mel", "T", "NK"),
           celltype_spider = setNames(celltype_spiderplots_collapsing[as.character(SendingType)], nm=NULL))
  
  df_edges = curr_obj$NICHES_output$CellToCellSpatial@meta.data %>%
    mutate(edge_id = colnames(curr_obj$NICHES_output$CellToCellSpatial),
           value = curr_obj$NICHES_output$CellToCellSpatial["LGALS1—PTPRC",]@assays$CellToCellSpatial@layers$data)
  df_edges[c("sending_id", "receiving_id")] = str_split_fixed(df_edges$edge_id, pattern = "—", n = 2)
  df_edges = df_edges %>%
    left_join(df %>% transmute(sending_id = SendingCell, x.Sending = x, y.Sending = y)) %>%
    left_join(df %>% transmute(receiving_id = SendingCell, x.Receiving = x, y.Receiving = y))
 
  avg_edge_values = df_edges %>% 
    group_by(sending_id) %>%
    summarize(n = n(), valuesum = sum(value)) %>%
    inner_join(df_edges %>% 
    group_by(receiving_id) %>%
    summarize(n = n(), valuesum = sum(value)), 
    by = join_by(sending_id == receiving_id)) %>%
    transmute(id = sending_id,
              value = (valuesum.x + valuesum.y)/(n.x + n.y -1))

  df = df %>% left_join(avg_edge_values, join_by(SendingCell == id))
  df_edges_pos = df_edges %>% filter(value>0)

inset_xmin = -1989
inset_xmax = -1039
inset_ymin = -6801
inset_ymax = -5851

inset2_xmin = -4639
inset2_xmax = -3689
inset2_ymin = -6337
inset2_ymax = -5387

edgeval_lims = c(0., quantile(df$value[df$value>0], p = c(0.995)))
edgeval_lims = ceiling(10*edgeval_lims)/10

p_maxi = ggplot(df, aes(x = x, y = y)) +
  mytheme() +
  theme(axis.text = element_blank(), axis.title = element_blank(),
        axis.ticks = element_blank(), legend.box.spacing = unit(8, "pt"), legend.text = element_text(size = 12),
        legend.position = "bottom", legend.direction = "horizontal", legend.box.margin = margin(l = 2, unit = "pt"),
        legend.title.position = "top", legend.title = element_text(face = "bold", size = 14), legend.justification.bottom = "left") +
  geom_node_voronoi(mapping = aes(fill=bordering), normalize = F, max.radius = 50, color = NA, alpha = 0.2) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region", guide = guide_legend(nrow = 2, order = 2)) +
  geom_point(aes(alpha = value, color = value), size = 3, stroke = 0.) +
  continuous_scale("alpha", name = waiver(), palette = rescale_saturating(range = c(0., 0.95), rel_saturating = 0.45),
                   limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish) +
  guides(alpha = guide_bins("Cell Average Interaction Value", order = 3,
                   theme = theme(legend.key.width = rel(1.3),
                                 legend.key = element_rect(fill = pattern, color = NA))),
         color = guide_bins("Cell Average Interaction Value", show.limits = T, order = 3,
                   override.aes = list(size = 6))) +
  new_scale_color() +
  geom_point(mapping = aes(x,y,color = celltype_spider), size = 0.08) +
  scale_color_manual(breaks = names(cellcolors), values = cellcolors, name = "Celltype",
                     guide = guide_legend(override.aes = list(size = 4), nrow = 3, order = 1)) +
  coord_equal(expand = F, xlim = range(df$x), ylim = range(df$y)) +
  annotate("segment", x = c(inset_xmin, inset_xmax), y = c(inset_ymin, inset_ymin), yend = c(inset_ymax, inset_ymax)) + # frame for inset
  annotate("segment", y = c(inset_ymin, inset_ymax), x = c(inset_xmin, inset_xmin), xend = c(inset_xmax, inset_xmax)) +
  annotate("segment", x = c(inset2_xmin, inset2_xmax), y = c(inset2_ymin, inset2_ymin), 
           yend = c(inset2_ymax, inset2_ymax), colour = "#1122AF") + # frame for inset
  annotate("segment", y = c(inset2_ymin, inset2_ymax), x = c(inset2_xmin, inset2_xmin), 
           xend = c(inset2_xmax, inset2_xmax), colour = "#1122AF") +
  plot_scalebars(max(df$x)-1200, max(df$y)-300, length = 1000, label = "1mm",
                 background_height_adjust = 70, tickheight_major = 30, background_expansion = 16,
                 tickheight_minor = 12, ticks_minor = c(0.2, 0.4, 0.6, 0.8)) 
  
p_mini_1 = ggplot(df_edges_pos, aes(x = x.Sending, y = y.Sending,
                           xend = x.Receiving, yend = y.Receiving)) +
  geom_node_voronoi(data = df, mapping = aes(fill=bordering),
                    normalize = F, max.radius = 25, color = NA, alpha = 0.2, show.legend = F) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region") +
  new_scale_fill() +
  geom_point(data = df, mapping = aes(x,y, alpha = value, color = value),
             inherit.aes = F, size = 3., show.legend = F) +
  scale_alpha_continuous(range = c(0., 1.), limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish, ) +
  new_scale_color() +
  geom_segment(mapping = aes(color = value), linewidth = 0.8) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = c(0., NA), name = "NICHES\nInteraction\nValue\n(Edges)") +
  new_scale_color() +
  geom_point(data = df, mapping = aes(x,y,color = celltype_spider),
             inherit.aes = F, size = 0.8, show.legend = F) +
  scale_color_manual(values = cellcolors) +
  mytheme() +
  coord_equal(xlim = c(inset_xmin, inset_xmax), ylim = c(inset_ymin, inset_ymax), expand = F) +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = "inside", legend.position.inside = c(.99, .99), legend.justification.inside = c(1, 1),
        legend.background = element_rect(fill=alpha("white", 0.8)), legend.key.height = unit(2, "lines"), legend.margin = margin(2,2,4,2, "pt"),
        plot.margin = margin(0,0,0,0, unit = "pt"), panel.background = element_blank(), legend.text = element_text(size = 12),
        panel.border = element_rect(fill = NA, color = "black", linewidth = 2), legend.title = element_text(face = "bold", size = 14)) +
  plot_scalebars(inset_xmax-210, inset_ymin+20)


p_mini_2 = ggplot(df_edges_pos, aes(x = x.Sending, y = y.Sending,
                           xend = x.Receiving, yend = y.Receiving)) +
  geom_node_voronoi(data = df, mapping = aes(fill=bordering),
                    normalize = F, max.radius = 25, color = NA, alpha = 0.2, show.legend = F) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region") +
  new_scale_fill() +
  geom_point(data = df, mapping = aes(x,y, alpha = value, color = value),
             inherit.aes = F, size = 3., show.legend = F) +
  scale_alpha_continuous(range = c(0., 1.), limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish, ) +
  new_scale_color() +
  geom_segment(mapping = aes(color = value), linewidth = 0.8) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = c(0., NA), name = "NICHES\nInteraction\nValue\n(Edges)") +
  new_scale_color() +
  geom_point(data = df, mapping = aes(x,y,color = celltype_spider),
             inherit.aes = F, size = 0.8, show.legend = F) +
  scale_color_manual(values = cellcolors) +
  mytheme() +
  coord_equal(xlim = c(inset2_xmin, inset2_xmax), ylim = c(inset2_ymin, inset2_ymax), expand = F) +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = "inside", legend.position.inside = c(.99, .99), legend.justification.inside = c(1, 1),
        legend.background = element_rect(fill=alpha("white", 0.8)), legend.key.height = unit(2, "lines"), legend.margin = margin(2,2,4,2, "pt"),
        plot.margin = margin(0,0,0,0, unit = "pt"), panel.background = element_blank(), legend.text = element_text(size = 12),
        panel.border = element_rect(fill = NA, color = "#1122AF", linewidth = 2), legend.title = element_text(face = "bold", size = 14)) +
  plot_scalebars(inset2_xmax-210, inset2_ymin+20)

wrap_plots(a = p_maxi, b = p_mini_1, c = p_mini_2,
           design = "
           aab
           aac
           ")
```

### HD_MM2

```{r hd-niches-5, fig.width=16., fig.asp=0.7, fig.show = 'hold', echo=T, eval=T}
  s = "HD_MM2"

  curr_obj = hd_objs_processed[[s]]
  df = curr_obj$NICHES_output$CellToNeighborhood@meta.data %>%
    mutate(object_id = as.numeric(SendingCell)) %>%
    left_join(curr_obj$obs_for_alluvials) %>%
    mutate(x = x_tr, y = y_tr, 
           filt = celltype_broad %in% c("Mel", "T", "NK"),
           celltype_spider = setNames(celltype_spiderplots_collapsing[as.character(SendingType)], nm=NULL))
  
  df_edges = curr_obj$NICHES_output$CellToCellSpatial@meta.data %>%
    mutate(edge_id = colnames(curr_obj$NICHES_output$CellToCellSpatial),
           value = curr_obj$NICHES_output$CellToCellSpatial["LGALS1—PTPRC",]@assays$CellToCellSpatial@layers$data)
  df_edges[c("sending_id", "receiving_id")] = str_split_fixed(df_edges$edge_id, pattern = "—", n = 2)
  df_edges = df_edges %>%
    left_join(df %>% transmute(sending_id = SendingCell, x.Sending = x, y.Sending = y)) %>%
    left_join(df %>% transmute(receiving_id = SendingCell, x.Receiving = x, y.Receiving = y))
 
  avg_edge_values = df_edges %>% 
    group_by(sending_id) %>%
    summarize(n = n(), valuesum = sum(value)) %>%
    inner_join(df_edges %>% 
    group_by(receiving_id) %>%
    summarize(n = n(), valuesum = sum(value)), 
    by = join_by(sending_id == receiving_id)) %>%
    transmute(id = sending_id,
              value = (valuesum.x + valuesum.y)/(n.x + n.y -1))

  df = df %>% left_join(avg_edge_values, join_by(SendingCell == id))
  df_edges_pos = df_edges %>% filter(value>0)

inset_xmin = 8224
inset_xmax = 9174
inset_ymin = 7368
inset_ymax = 8318

inset2_xmin = 6957
inset2_xmax = 7907
inset2_ymin = 9568
inset2_ymax = 10518

edgeval_lims = c(0., quantile(df$value[df$value>0], p = c(0.995)))
edgeval_lims = ceiling(10*edgeval_lims)/10

p_maxi = ggplot(df, aes(x = x, y = y)) +
  mytheme() +
  theme(axis.text = element_blank(), axis.title = element_blank(),
        axis.ticks = element_blank(), legend.box.spacing = unit(8, "pt"), legend.text = element_text(size = 12),
        legend.position = "bottom", legend.direction = "horizontal", legend.box.margin = margin(l = 2, unit = "pt"),
        legend.title.position = "top", legend.title = element_text(face = "bold", size = 14), legend.justification.bottom = "left") +
  geom_node_voronoi(mapping = aes(fill=bordering), normalize = F, max.radius = 50, color = NA, alpha = 0.2) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region", guide = guide_legend(nrow = 2, order = 2)) +
  geom_point(aes(alpha = value, color = value), size = 3, stroke = 0.) +
  continuous_scale("alpha", name = waiver(), palette = rescale_saturating(range = c(0., 0.95), rel_saturating = 0.45),
                   limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish) +
  guides(alpha = guide_bins("Cell Average Interaction Value", order = 3,
                   theme = theme(legend.key.width = rel(1.3),
                                 legend.key = element_rect(fill = pattern, color = NA))),
         color = guide_bins("Cell Average Interaction Value", show.limits = T, order = 3,
                   override.aes = list(size = 6))) +
  new_scale_color() +
  geom_point(mapping = aes(x,y,color = celltype_spider), size = 0.08) +
  scale_color_manual(breaks = names(cellcolors), values = cellcolors, name = "Celltype",
                     guide = guide_legend(override.aes = list(size = 4), nrow = 3, order = 1)) +
  coord_equal(expand = F, xlim = range(df$x), ylim = range(df$y)) +
  annotate("segment", x = c(inset_xmin, inset_xmax), y = c(inset_ymin, inset_ymin), yend = c(inset_ymax, inset_ymax)) + # frame for inset
  annotate("segment", y = c(inset_ymin, inset_ymax), x = c(inset_xmin, inset_xmin), xend = c(inset_xmax, inset_xmax)) +
  annotate("segment", x = c(inset2_xmin, inset2_xmax), y = c(inset2_ymin, inset2_ymin), 
           yend = c(inset2_ymax, inset2_ymax), colour = "#1122AF") + # frame for inset
  annotate("segment", y = c(inset2_ymin, inset2_ymax), x = c(inset2_xmin, inset2_xmin), 
           xend = c(inset2_xmax, inset2_xmax), colour = "#1122AF") +
  plot_scalebars(max(df$x)-1200, max(df$y)-300, length = 1000, label = "1mm",
                 background_height_adjust = 70, tickheight_major = 30, background_expansion = 16,
                 tickheight_minor = 12, ticks_minor = c(0.2, 0.4, 0.6, 0.8)) 
  
p_mini_1 = ggplot(df_edges_pos, aes(x = x.Sending, y = y.Sending,
                           xend = x.Receiving, yend = y.Receiving)) +
  geom_node_voronoi(data = df, mapping = aes(fill=bordering),
                    normalize = F, max.radius = 25, color = NA, alpha = 0.2, show.legend = F) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region") +
  new_scale_fill() +
  geom_point(data = df, mapping = aes(x,y, alpha = value, color = value),
             inherit.aes = F, size = 3., show.legend = F) +
  scale_alpha_continuous(range = c(0., 1.), limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish, ) +
  new_scale_color() +
  geom_segment(mapping = aes(color = value), linewidth = 0.8) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = c(0., NA), name = "NICHES\nInteraction\nValue\n(Edges)") +
  new_scale_color() +
  geom_point(data = df, mapping = aes(x,y,color = celltype_spider),
             inherit.aes = F, size = 0.8, show.legend = F) +
  scale_color_manual(values = cellcolors) +
  mytheme() +
  coord_equal(xlim = c(inset_xmin, inset_xmax), ylim = c(inset_ymin, inset_ymax), expand = F) +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = "inside", legend.position.inside = c(.99, .99), legend.justification.inside = c(1, 1),
        legend.background = element_rect(fill=alpha("white", 0.8)), legend.key.height = unit(2, "lines"), legend.margin = margin(2,2,4,2, "pt"),
        plot.margin = margin(0,0,0,0, unit = "pt"), panel.background = element_blank(), legend.text = element_text(size = 12),
        panel.border = element_rect(fill = NA, color = "black", linewidth = 2), legend.title = element_text(face = "bold", size = 14)) +
  plot_scalebars(inset_xmax-210, inset_ymin+20)


p_mini_2 = ggplot(df_edges_pos, aes(x = x.Sending, y = y.Sending,
                           xend = x.Receiving, yend = y.Receiving)) +
  geom_node_voronoi(data = df, mapping = aes(fill=bordering),
                    normalize = F, max.radius = 25, color = NA, alpha = 0.2, show.legend = F) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region") +
  new_scale_fill() +
  geom_point(data = df, mapping = aes(x,y, alpha = value, color = value),
             inherit.aes = F, size = 3., show.legend = F) +
  scale_alpha_continuous(range = c(0., 1.), limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish, ) +
  new_scale_color() +
  geom_segment(mapping = aes(color = value), linewidth = 0.8) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = c(0., NA), name = "NICHES\nInteraction\nValue\n(Edges)") +
  new_scale_color() +
  geom_point(data = df, mapping = aes(x,y,color = celltype_spider),
             inherit.aes = F, size = 0.8, show.legend = F) +
  scale_color_manual(values = cellcolors) +
  mytheme() +
  coord_equal(xlim = c(inset2_xmin, inset2_xmax), ylim = c(inset2_ymin, inset2_ymax), expand = F) +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = "inside", legend.position.inside = c(.99, .99), legend.justification.inside = c(1, 1),
        legend.background = element_rect(fill=alpha("white", 0.8)), legend.key.height = unit(2, "lines"), legend.margin = margin(2,2,4,2, "pt"),
        plot.margin = margin(0,0,0,0, unit = "pt"), panel.background = element_blank(), legend.text = element_text(size = 12),
        panel.border = element_rect(fill = NA, color = "#1122AF", linewidth = 2), legend.title = element_text(face = "bold", size = 14)) +
  plot_scalebars(inset2_xmax-210, inset2_ymin+20)

wrap_plots(a = p_maxi, b = p_mini_1, c = p_mini_2,
           design = "
           aab
           aac
           ")
```

### HD_MM3

```{r hd-niches-6, fig.width=16., fig.asp=0.7, fig.show = 'hold', echo=T, eval=T}
  s = "HD_MM3"

  curr_obj = hd_objs_processed[[s]]
  df = curr_obj$NICHES_output$CellToNeighborhood@meta.data %>%
    mutate(object_id = as.numeric(SendingCell)) %>%
    left_join(curr_obj$obs_for_alluvials) %>%
    mutate(x = x_tr, y = y_tr, 
           filt = celltype_broad %in% c("Mel", "T", "NK"),
           celltype_spider = setNames(celltype_spiderplots_collapsing[as.character(SendingType)], nm=NULL))
  
  df_edges = curr_obj$NICHES_output$CellToCellSpatial@meta.data %>%
    mutate(edge_id = colnames(curr_obj$NICHES_output$CellToCellSpatial),
           value = curr_obj$NICHES_output$CellToCellSpatial["LGALS1—PTPRC",]@assays$CellToCellSpatial@layers$data)
  df_edges[c("sending_id", "receiving_id")] = str_split_fixed(df_edges$edge_id, pattern = "—", n = 2)
  df_edges = df_edges %>%
    left_join(df %>% transmute(sending_id = SendingCell, x.Sending = x, y.Sending = y)) %>%
    left_join(df %>% transmute(receiving_id = SendingCell, x.Receiving = x, y.Receiving = y))
 
  avg_edge_values = df_edges %>% 
    group_by(sending_id) %>%
    summarize(n = n(), valuesum = sum(value)) %>%
    inner_join(df_edges %>% 
    group_by(receiving_id) %>%
    summarize(n = n(), valuesum = sum(value)), 
    by = join_by(sending_id == receiving_id)) %>%
    transmute(id = sending_id,
              value = (valuesum.x + valuesum.y)/(n.x + n.y -1))

  df = df %>% left_join(avg_edge_values, join_by(SendingCell == id))
  df_edges_pos = df_edges %>% filter(value>0)

inset_xmin = -4764
inset_xmax = -3814
inset_ymin = -3249
inset_ymax = -2299

inset2_xmin = -7594
inset2_xmax = -6657
inset2_ymin = -6471
inset2_ymax = -5534

edgeval_lims = c(0., quantile(df$value[df$value>0], p = c(0.995)))
edgeval_lims = ceiling(10*edgeval_lims)/10

p_maxi = ggplot(df, aes(x = x, y = y)) +
  mytheme() +
  theme(axis.text = element_blank(), axis.title = element_blank(),
        axis.ticks = element_blank(), legend.box.spacing = unit(8, "pt"), legend.text = element_text(size = 12),
        legend.position = "bottom", legend.direction = "horizontal", legend.box.margin = margin(l = 2, unit = "pt"),
        legend.title.position = "top", legend.title = element_text(face = "bold", size = 14), legend.justification.bottom = "left") +
  geom_node_voronoi(mapping = aes(fill=bordering), normalize = F, max.radius = 50, color = NA, alpha = 0.2) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region", guide = guide_legend(nrow = 2, order = 2)) +
  geom_point(aes(alpha = value, color = value), size = 3, stroke = 0.) +
  continuous_scale("alpha", name = waiver(), palette = rescale_saturating(range = c(0., 0.95), rel_saturating = 0.45),
                   limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish) +
  guides(alpha = guide_bins("Cell Average Interaction Value", order = 3,
                   theme = theme(legend.key.width = rel(1.3),
                                 legend.key = element_rect(fill = pattern, color = NA))),
         color = guide_bins("Cell Average Interaction Value", show.limits = T, order = 3,
                   override.aes = list(size = 6))) +
  new_scale_color() +
  geom_point(mapping = aes(x,y,color = celltype_spider), size = 0.08) +
  scale_color_manual(breaks = names(cellcolors), values = cellcolors, name = "Celltype",
                     guide = guide_legend(override.aes = list(size = 4), nrow = 3, order = 1)) +
  coord_equal(expand = F, xlim = range(df$x), ylim = range(df$y)) +
  annotate("segment", x = c(inset_xmin, inset_xmax), y = c(inset_ymin, inset_ymin), yend = c(inset_ymax, inset_ymax)) + # frame for inset
  annotate("segment", y = c(inset_ymin, inset_ymax), x = c(inset_xmin, inset_xmin), xend = c(inset_xmax, inset_xmax)) +
  annotate("segment", x = c(inset2_xmin, inset2_xmax), y = c(inset2_ymin, inset2_ymin), 
           yend = c(inset2_ymax, inset2_ymax), colour = "#1122AF") + # frame for inset
  annotate("segment", y = c(inset2_ymin, inset2_ymax), x = c(inset2_xmin, inset2_xmin), 
           xend = c(inset2_xmax, inset2_xmax), colour = "#1122AF") +
  plot_scalebars(max(df$x)-1200, max(df$y)-300, length = 1000, label = "1mm",
                 background_height_adjust = 70, tickheight_major = 30, background_expansion = 16,
                 tickheight_minor = 12, ticks_minor = c(0.2, 0.4, 0.6, 0.8)) 
  
p_mini_1 = ggplot(df_edges_pos, aes(x = x.Sending, y = y.Sending,
                           xend = x.Receiving, yend = y.Receiving)) +
  geom_node_voronoi(data = df, mapping = aes(fill=bordering),
                    normalize = F, max.radius = 25, color = NA, alpha = 0.2, show.legend = F) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region") +
  new_scale_fill() +
  geom_point(data = df, mapping = aes(x,y, alpha = value, color = value),
             inherit.aes = F, size = 3., show.legend = F) +
  scale_alpha_continuous(range = c(0., 1.), limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish, ) +
  new_scale_color() +
  geom_segment(mapping = aes(color = value), linewidth = 0.8) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = c(0., NA), name = "NICHES\nInteraction\nValue\n(Edges)") +
  new_scale_color() +
  geom_point(data = df, mapping = aes(x,y,color = celltype_spider),
             inherit.aes = F, size = 0.8, show.legend = F) +
  scale_color_manual(values = cellcolors) +
  mytheme() +
  coord_equal(xlim = c(inset_xmin, inset_xmax), ylim = c(inset_ymin, inset_ymax), expand = F) +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = "inside", legend.position.inside = c(.8, .99), legend.justification.inside = c(1, 1),
        legend.background = element_rect(fill=alpha("white", 0.8)), legend.key.height = unit(2, "lines"), legend.margin = margin(2,2,4,2, "pt"),
        plot.margin = margin(0,0,0,0, unit = "pt"), panel.background = element_blank(), legend.text = element_text(size = 12),
        panel.border = element_rect(fill = NA, color = "black", linewidth = 2), legend.title = element_text(face = "bold", size = 14)) +
  plot_scalebars(inset_xmax-210, inset_ymin+20)


p_mini_2 = ggplot(df_edges_pos, aes(x = x.Sending, y = y.Sending,
                           xend = x.Receiving, yend = y.Receiving)) +
  geom_node_voronoi(data = df, mapping = aes(fill=bordering),
                    normalize = F, max.radius = 25, color = NA, alpha = 0.2, show.legend = F) +
  scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
                               "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
                    name = "Region") +
  new_scale_fill() +
  geom_point(data = df, mapping = aes(x,y, alpha = value, color = value),
             inherit.aes = F, size = 3., show.legend = F) +
  scale_alpha_continuous(range = c(0., 1.), limits = edgeval_lims, oob = scales::squish) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish, ) +
  new_scale_color() +
  geom_segment(mapping = aes(color = value), linewidth = 0.8) +
  scale_color_distiller(direction = 1, palette = "PuBuGn", limits = c(0., NA), name = "NICHES\nInteraction\nValue\n(Edges)") +
  new_scale_color() +
  geom_point(data = df, mapping = aes(x,y,color = celltype_spider),
             inherit.aes = F, size = 0.8, show.legend = F) +
  scale_color_manual(values = cellcolors) +
  mytheme() +
  coord_equal(xlim = c(inset2_xmin, inset2_xmax), ylim = c(inset2_ymin, inset2_ymax), expand = F) +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = "inside", legend.position.inside = c(.80, .99), legend.justification.inside = c(1, 1),
        legend.background = element_rect(fill=alpha("white", 0.8)), legend.key.height = unit(2, "lines"), legend.margin = margin(2,2,4,2, "pt"),
        plot.margin = margin(0,0,0,0, unit = "pt"), panel.background = element_blank(), legend.text = element_text(size = 12),
        panel.border = element_rect(fill = NA, color = "#1122AF", linewidth = 2), legend.title = element_text(face = "bold", size = 14)) +
  plot_scalebars(inset2_xmax-210, inset2_ymin+20)

wrap_plots(a = p_maxi, b = p_mini_1, c = p_mini_2,
           design = "
           aab
           aac
           ")
```

## celltype connection plots for other mechanisms

### LGALS3—CD6

```{r celltype-conns-supp-1, fig.width=2*8., fig.asp=0.7, fig.show = 'hold', echo=T, eval=T}
dat = do.call(rbind, lapply(seq_along(liana_aggs), function(i){
    agg = liana_aggs[[i]] %>%
        filter(aggregate_rank<=0.01 & ligand.complex=="LGALS3" & receptor.complex=="CD6")
    agg$sample=names(liana_aggs)[i]
    agg
})) %>%
  group_by(source, target) %>%
  distinct() %>%
  mutate(n = n()) %>%
  summarize(aggregate_rank = mean(aggregate_rank),
            n = first(n),
            samples = paste(sample, collapse=", "),
            .groups = "drop")

celltypes_connection_plot(dat %>% filter(n>=2),
                          celltype_colors_table = cellcolors, layout = function(g)convert(g, to_undirected) %>% ggraph::create_layout("backbone") %>% select(x,y))
```

### ANXA1—CXCR4

```{r celltype-conns-supp-2, fig.width=2*8., fig.asp=0.7, fig.show = 'hold', echo=T, eval=T}
dat = do.call(rbind, lapply(seq_along(liana_aggs), function(i){
    agg = liana_aggs[[i]] %>%
        filter(aggregate_rank<=0.01 & ligand.complex=="ANXA1" & receptor.complex=="CXCR4")
    agg$sample=names(liana_aggs)[i]
    agg
})) %>%
  group_by(source, target) %>%
  distinct() %>%
  mutate(n = n()) %>%
  summarize(aggregate_rank = mean(aggregate_rank),
            n = first(n),
            samples = paste(sample, collapse=", "),
            .groups = "drop")

set.seed(1234)
celltypes_connection_plot(dat %>% filter(n>=2),
                          celltype_colors_table = cellcolors, layout = "nicely")


celltypes_connection_plot(dat %>% filter(n>=3),
                          celltype_colors_table = cellcolors, layout = function(g)convert(g, to_undirected) %>% ggraph::create_layout("backbone") %>% select(x,y))
```

### CD59—CD2

```{r celltype-conns-supp-3, fig.width=2*8., fig.asp=0.7, fig.show = 'hold', echo=T, eval=T}
dat = do.call(rbind, lapply(seq_along(liana_aggs), function(i){
    agg = liana_aggs[[i]] %>%
        filter(aggregate_rank<=0.01 & ligand.complex=="CD59" & receptor.complex=="CD2")
    agg$sample=names(liana_aggs)[i]
    agg
})) %>%
  group_by(source, target) %>%
  distinct() %>%
  mutate(n = n()) %>%
  summarize(aggregate_rank = mean(aggregate_rank),
            n = first(n),
            samples = paste(sample, collapse=", "),
            .groups = "drop")

set.seed(12345)
celltypes_connection_plot(dat %>% filter(n>=2),
                          celltype_colors_table = cellcolors, layout = "nicely")


celltypes_connection_plot(dat %>% filter(n>=3),
                          celltype_colors_table = cellcolors, layout = function(g)convert(g, to_undirected) %>% ggraph::create_layout("backbone") %>% select(x,y))
```

### VIM—CD44

```{r celltype-conns-supp-4, fig.width=2*8., fig.asp=1, fig.show = 'hold', echo=T, eval=T}
dat = do.call(rbind, lapply(seq_along(liana_aggs), function(i){
    agg = liana_aggs[[i]] %>%
        filter(aggregate_rank<=0.01 & ligand.complex=="VIM" & receptor.complex=="CD44")
    agg$sample=names(liana_aggs)[i]
    agg
})) %>%
  group_by(source, target) %>%
  distinct() %>%
  mutate(n = n()) %>%
  summarize(aggregate_rank = mean(aggregate_rank),
            n = first(n),
            samples = paste(sample, collapse=", "),
            .groups = "drop")

celltypes_connection_plot(dat %>% filter(n>=3),
                          celltype_colors_table = cellcolors, layout = "circle")


celltypes_connection_plot(dat %>% filter(n>=4),
                          celltype_colors_table = cellcolors, layout = function(g)convert(convert(g, to_undirected), to_simple) %>% ggraph::create_layout("backbone") %>% select(x,y))
```

### MIF—CD74

```{r celltype-conns-supp-5, fig.width=2*8., fig.asp=1, fig.show = 'hold', echo=T, eval=T}
dat = do.call(rbind, lapply(seq_along(liana_aggs), function(i){
    agg = liana_aggs[[i]] %>%
        filter(aggregate_rank<=0.01 & ligand.complex=="MIF" & receptor.complex=="CD74")
    agg$sample=names(liana_aggs)[i]
    agg
})) %>%
  group_by(source, target) %>%
  distinct() %>%
  mutate(n = n()) %>%
  summarize(aggregate_rank = mean(aggregate_rank),
            n = first(n),
            samples = paste(sample, collapse=", "),
            .groups = "drop")

celltypes_connection_plot(dat %>% filter(n>=3),
                          celltype_colors_table = cellcolors, layout = "circle")
set.seed(123456)
celltypes_connection_plot(dat %>% filter(n>=4),
                          celltype_colors_table = cellcolors, layout = "nicely")
```

### ITGB2—ITGAL

```{r celltype-conns-supp-6, fig.width=2*8., fig.asp=1, fig.show = 'hold', echo=T, eval=T}
dat = do.call(rbind, lapply(seq_along(liana_aggs), function(i){
    agg = liana_aggs[[i]] %>%
        filter(aggregate_rank<=0.01 & ligand.complex=="ITGB2" & receptor.complex=="ITGAL")
    agg$sample=names(liana_aggs)[i]
    agg
})) %>%
  group_by(source, target) %>%
  distinct() %>%
  mutate(n = n()) %>%
  summarize(aggregate_rank = mean(aggregate_rank),
            n = first(n),
            samples = paste(sample, collapse=", "),
            .groups = "drop")

celltypes_connection_plot(dat %>% filter(n>=1),
                          celltype_colors_table = cellcolors, layout = "circle")
set.seed(12345)
celltypes_connection_plot(dat %>% filter(n>=2),
                          celltype_colors_table = cellcolors, layout = "nicely")
```


```{dummy, eval = F}
### `r I(names(hd_objs_processed)[1])`

# ```{r hd-niches-1, fig.width=16., fig.asp=0.7, fig.show = 'hold', echo=T, eval=T}
#   s = 1 # 983-1 removed due to missing consent
# 
#   curr_obj = hd_objs_processed[[s]]
#   df = curr_obj$NICHES_output$CellToNeighborhood@meta.data %>%
#     mutate(object_id = as.numeric(SendingCell)) %>%
#     left_join(curr_obj$obs_for_alluvials) %>%
#     mutate(x = x_tr, y = y_tr, 
#            filt = celltype_broad %in% c("Mel", "T", "NK"),
#            celltype_spider = setNames(celltype_spiderplots_collapsing[as.character(SendingType)], nm=NULL))
#   
#   df_edges = curr_obj$NICHES_output$CellToCellSpatial@meta.data %>%
#     mutate(edge_id = colnames(curr_obj$NICHES_output$CellToCellSpatial),
#            value = curr_obj$NICHES_output$CellToCellSpatial["LGALS1—PTPRC",]@assays$CellToCellSpatial@layers$data)
#   df_edges[c("sending_id", "receiving_id")] = str_split_fixed(df_edges$edge_id, pattern = "—", n = 2)
#   df_edges = df_edges %>%
#     left_join(df %>% transmute(sending_id = SendingCell, x.Sending = x, y.Sending = y)) %>%
#     left_join(df %>% transmute(receiving_id = SendingCell, x.Receiving = x, y.Receiving = y))
#  
#   avg_edge_values = df_edges %>% 
#     group_by(sending_id) %>%
#     summarize(n = n(), valuesum = sum(value)) %>%
#     inner_join(df_edges %>% 
#     group_by(receiving_id) %>%
#     summarize(n = n(), valuesum = sum(value)), 
#     by = join_by(sending_id == receiving_id)) %>%
#     transmute(id = sending_id,
#               value = (valuesum.x + valuesum.y)/(n.x + n.y -1))
# 
#   df = df %>% left_join(avg_edge_values, join_by(SendingCell == id))
#   df_edges_pos = df_edges %>% filter(value>0)
#     
# inset_xmin = 950
# inset_xmax = 1400
# inset_ymin = 400
# inset_ymax = 850
# 
# inset2_xmin = 1850
# inset2_xmax = 2400
# inset2_ymin = 770
# inset2_ymax = 1320
# 
# edgeval_lims = c(0., quantile(df$value[df$value>0], p = c(0.995)))
# edgeval_lims = ceiling(10*edgeval_lims)/10
# 
# p_maxi = ggplot(df, aes(x = x, y = y)) +
#   mytheme() +
#   theme(axis.text = element_blank(), axis.title = element_blank(),
#         axis.ticks = element_blank(), legend.box.spacing = unit(8, "pt"), legend.text = element_text(size = 12),
#         legend.position = "bottom", legend.direction = "horizontal", legend.box.margin = margin(l = 2, unit = "pt"),
#         legend.title.position = "top", legend.title = element_text(face = "bold", size = 14), legend.justification.bottom = "left") +
#   geom_node_voronoi(mapping = aes(fill=bordering), normalize = F, max.radius = 50, color = NA, alpha = 0.2) +
#   scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
#                                "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
#                     name = "Region", guide = guide_legend(nrow = 2, order = 2)) +
#   geom_point(aes(alpha = value, color = value), size = 3, stroke = 0.) +
#   continuous_scale("alpha", name = waiver(), palette = rescale_saturating(range = c(0., 0.95), rel_saturating = 0.45),
#                    limits = edgeval_lims, oob = scales::squish) +
#   scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish) +
#   guides(alpha = guide_bins("Cell Average Interaction Value", 
#                    theme = theme(legend.key.width = rel(1.3),
#                                  legend.key = element_rect(fill = pattern, color = NA))),
#          color = guide_bins("Cell Average Interaction Value", show.limits = T, order = 3,
#                    override.aes = list(size = 6))) +
#   new_scale_color() +
#   geom_point(mapping = aes(x,y,color = celltype_spider), size = 0.08) +
#   scale_color_manual(breaks = names(cellcolors), values = cellcolors, name = "Celltype",
#                      guide = guide_legend(override.aes = list(size = 4), nrow = 3, order = 1)) +
#   coord_equal(expand = F, xlim = range(df$x), ylim = range(df$y)) +
#   annotate("segment", x = c(inset_xmin, inset_xmax), y = c(inset_ymin, inset_ymin), yend = c(inset_ymax, inset_ymax)) + # frame for inset
#   annotate("segment", y = c(inset_ymin, inset_ymax), x = c(inset_xmin, inset_xmin), xend = c(inset_xmax, inset_xmax)) +
#   annotate("segment", x = c(inset2_xmin, inset2_xmax), y = c(inset2_ymin, inset2_ymin), 
#            yend = c(inset2_ymax, inset2_ymax), colour = "#1122AF") + # frame for inset
#   annotate("segment", y = c(inset2_ymin, inset2_ymax), x = c(inset2_xmin, inset2_xmin), 
#            xend = c(inset2_xmax, inset2_xmax), colour = "#1122AF") +
#   plot_scalebars(max(df$x)-600, max(df$y)-150, length = 1000, label = "1mm",
#                  background_height_adjust = 70, tickheight_major = 30, background_expansion = 16,
#                  tickheight_minor = 6, ticks_minor = c(0.2, 0.4, 0.6, 0.8)) 
#   
# p_mini_1 = ggplot(df_edges_pos, aes(x = x.Sending, y = y.Sending,
#                            xend = x.Receiving, yend = y.Receiving)) +
#   geom_node_voronoi(data = df, mapping = aes(fill=bordering),
#                     normalize = F, max.radius = 25, color = NA, alpha = 0.2, show.legend = F) +
#   scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
#                                "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
#                     name = "Region") +
#   new_scale_fill() +
#   geom_point(data = df, mapping = aes(x,y, alpha = value, color = value),
#              inherit.aes = F, size = 3., show.legend = F) +
#   scale_alpha_continuous(range = c(0., 1.), limits = edgeval_lims, oob = scales::squish) +
#   scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish) +
#   new_scale_color() +
#   geom_segment(mapping = aes(color = value), linewidth = 0.8) +
#   scale_color_distiller(direction = 1, palette = "PuBuGn", limits = c(0., NA), name = "NICHES\nInteraction\nValue\n(Edges)") +
#   new_scale_color() +
#   geom_point(data = df, mapping = aes(x,y,color = celltype_spider),
#              inherit.aes = F, size = 0.8, show.legend = F) +
#   scale_color_manual(values = cellcolors) +
#   mytheme() +
#   coord_equal(xlim = c(inset_xmin, inset_xmax), ylim = c(inset_ymin, inset_ymax), expand = F) +
#   theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
#         legend.position = "inside", legend.position.inside = c(.99, .99), legend.justification.inside = c(1, 1),
#         legend.background = element_rect(fill=alpha("white", 0.8)), legend.key.height = unit(2, "lines"),
#         plot.margin = margin(0,0,0,0, unit = "pt"), panel.background = element_blank(), legend.text = element_text(size = 12),
#         panel.border = element_rect(fill = NA, color = "black", linewidth = 2), legend.title = element_text(face = "bold", size = 14)) +
#   plot_scalebars(inset_xmax-210, inset_ymin+20)
# 
# 
# p_mini_2 = ggplot(df_edges_pos, aes(x = x.Sending, y = y.Sending,
#                            xend = x.Receiving, yend = y.Receiving)) +
#   geom_node_voronoi(data = df, mapping = aes(fill=bordering),
#                     normalize = F, max.radius = 25, color = NA, alpha = 0.2, show.legend = F) +
#   scale_fill_manual(values = c("Background" = NA, "T/NK neighborhood" = "darkgreen",
#                                "Tumor neighborhood" = "darkred", "Tumor-Immune border" = "#DDDD22"),
#                     name = "Region") +
#   new_scale_fill() +
#   geom_point(data = df, mapping = aes(x,y, alpha = value, color = value),
#              inherit.aes = F, size = 3., show.legend = F) +
#   scale_alpha_continuous(range = c(0., 1.), limits = edgeval_lims, oob = scales::squish) +
#   scale_color_distiller(direction = 1, palette = "PuBuGn", limits = edgeval_lims, oob = scales::squish, ) +
#   new_scale_color() +
#   geom_segment(mapping = aes(color = value), linewidth = 0.8) +
#   scale_color_distiller(direction = 1, palette = "PuBuGn", limits = c(0., NA), name = "NICHES\nInteraction\nValue\n(Edges)") +
#   new_scale_color() +
#   geom_point(data = df, mapping = aes(x,y,color = celltype_spider),
#              inherit.aes = F, size = 0.8, show.legend = F) +
#   scale_color_manual(values = cellcolors) +
#   mytheme() +
#   coord_equal(xlim = c(inset2_xmin, inset2_xmax), ylim = c(inset2_ymin, inset2_ymax), expand = F) +
#   theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
#         legend.position = "inside", legend.position.inside = c(.99, .99), legend.justification.inside = c(1, 1),
#         legend.background = element_rect(fill=alpha("white", 0.8)), legend.key.height = unit(2, "lines"),
#         plot.margin = margin(0,0,0,0, unit = "pt"), panel.background = element_blank(), legend.text = element_text(size = 12),
#         panel.border = element_rect(fill = NA, color = "#1122AF", linewidth = 2), legend.title = element_text(face = "bold", size = 14)) +
#   plot_scalebars(inset2_xmax-110, inset2_ymin+10, background_expansion = 4, background_height_adjust = 11)
# 
# wrap_plots(a = p_maxi, b = p_mini_1, c = p_mini_2,
#            design = "
#            aab
#            aac
#            ")
# ```
```
