---
title: "UCCL Melanoma 10X Visium Project - Part 2"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
workflowr:
  suppress_report: TRUE
editor_options: 
  chunk_output_type: console
---

Author: Florian Große, Michael Rade (Last updated: `r Sys.Date()`)

```{r libs-and-functions, echo = F, message = FALSE, warning=FALSE}
library("yaml")
library("readr")
library("kableExtra")
```

# Available samples

```{r good-samples, results='asis', echo = FALSE}
manifest = yaml.load_file("./assets/manifest.yaml")
all_samples = setdiff(manifest$samples, manifest$rejected_samples)
# optional: custom ordering for nicer list
# all_samples = c(sort(all_samples[startsWith(all_samples, "P")]), sort(all_samples[!startsWith(all_samples, "P")]))

has_rejected_samples = length(manifest$rejected_samples)>0 
has_external_samples = length(manifest$external_samples)>0 

for (sample in all_samples){
  sample_full_name = manifest[[sample]]$full_name
  cat("* [Sample: ", sample_full_name, "](./", sample, "/index.html)", "\n", sep = "")
}
cat("\n\n")
```

```{r rejected-samples, results='asis', echo = FALSE, eval=has_rejected_samples}
cat("\n# Rejected samples with only basic analysis steps present\n\n")
rejected_samples = manifest$rejected_samples
for (sample in rejected_samples){
  sample_full_name = manifest[[sample]]$full_name
  cat("* [Sample: ", sample_full_name, "](./", sample, "/index.html)", "\n", sep = "")
}
cat("\n\n")
```

```{r external-samples, results='asis', echo = FALSE, eval=has_external_samples}
cat("\n# Samples imported from external, previously published datasets\n\n")
external_samples = manifest$external_samples
for (sample in external_samples){
  sample_full_name = manifest[[sample]]$full_name
  cat("* [Sample: ", sample_full_name, "](./", sample, "/index.html)", "\n", sep = "")
}
cat("\n\n")
```


# Analyses Spanning Multiple Samples

* [Aggregation of LRI Analysis](10_global_02_aggregated_LRI.html)
* [Basic stats for this papers samples](11_global_03_basic_spot_stats.html)
* [TCIA Survival analyses](14_global_04_TCGA_survival.html)

# Collected Plots for the Paper

* [Visium non-HD and Survival](15_paper_plots_merged.html)
* [Visium HD Plots](16_HD_plots.html)

# Notes

```{r loadnotes, results='hide', echo = FALSE, eval=file.exists("./assets/global_notes.csv")}
# removed, since they contained internal identifiers and potentially sensitive data
gn = read_csv("./assets/global_notes.csv", show_col_types = F)
```

```{r shownotes, results='markup', echo = FALSE, eval=file.exists("./assets/global_notes.csv")}
kable_styling(kbl(gn), full_width = T)
```


The In-/Output paths for all files used/produced in this analysis are stored in "manifest.yaml" (assets/manifest.yaml). 
Additionally, all major sample-specific settings are in the manifest as well.

HTML sites were compiled using [workflowr](https://github.com/jdblischak/workflowr).

# Labbook structure

```
2025-UCCL-Melanom-Spatial-Comparisons/
├── analysis/               <---------- R Markdown files
├── assets/                 <---------- Manifest and misc. files that are synchronized
│   ├── annotations/        <---------- Histological annotations for each sample from the Loupe browser
│   ├── manifest.yaml
│   └── ...
├── data                    <---------- Caches for various downloaded datasets etc., not synchronized
├── code/                   <---------- R scripts to process the data
├── public/                 <---------- Unused, see README
├── .gitlab-ci.yml
├── _workflowr.yml
├── README.md
├── renv/                   <---------= Dependency version control
├── renv.lock               <--------/ 
├── .gitignore
├── .gitattributes
├── 2025-UCCL-Melanom-Spatial-Comparisons.Rproj
├── .Rprofile
└── Interactive_analysis.R  <---------- Read this if you want to redo the analysis
```
