---
title: "UCCL Melanoma 10X Visium Project - Part 2"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
workflowr:
  suppress_report: TRUE
editor_options: 
  chunk_output_type: console
---

```{r sample-title, results='asis', echo = FALSE}
library(yaml)
manifest = yaml.load_file("./assets/manifest.yaml")
current_sample = scan(file.path(manifest$workdata_parent, ".tmp/active_sample"), what = "character", quiet = T)
sample_full_name = manifest[[current_sample]]$full_name
cat("# Sample:", sample_full_name, "\n")

```

Author: Florian Große, Michael Rade (Last updated: `r Sys.Date()`)

## Data analysis

* [Quality control: `r I(current_sample)`](01_qc.html)
* [QC Spot Selection: `r I(current_sample)`](03_qc_select.html)
* [Expert Annotations: `r I(current_sample)`](04_annotations.html)
* [Deconvolved Celltypes: `r I(current_sample)`](05_deconv_results.html)
* [Module Scores: `r I(current_sample)`](06_module_scores.html)
* [Cluster Markers and ORA: `r I(current_sample)`](07_clustering.html)
* [Collected Plots: `r I(current_sample)`](08_paper_plots.html)
```{r only-non-rejected, results='asis', echo = FALSE}
if (!(current_sample %in% manifest$rejected_samples)){
  cat(paste0(
"* [Ligand-Receptor Analysis: ",current_sample,"](09_LRI.html)

"))
}


```



```{r notes, results='asis', echo = FALSE}
note = manifest[[current_sample]]$note
if (!(length(note) == 0 || note == "")){
  cat("# Notes \n\n")
  cat(note)
}
```


